---
title: "Demonstrating Brute Force Fit"
author: "Bill Prisbrey"
date: "2025-10-21"
output:
  html_document:
    keep_md: true
---


```{r include=FALSE}

# Author's notes:

# I want a handful of flowcharts showing the breakdown of the individual journey,
# with the counts or percentages of the different "regimes"

# Then I want a handful of individual journey plots demonstrating the different regimes

# I should have my "boundary" graphic at the bottom as well



```

  
```{r include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.height=7, fig.width=10)

# adjust graphic parameters

oldPar <- par(cex.main = 3,
    cex.lab = 3,
    cex.axis = 2,
    mar = c(5.1,4.1,4.1,2.1), # default is c(5.1,4.1,4.1,2.1)
    mfrow = c(1,1)
    )

```


```{r}

##########
## LOAD ##
##########

journeyPopulation <- readRDS(here::here("Data", "journeyPopulation.rds") )
piEMPLIDs <- readRDS(here::here("Data", "piEMPLIDs.rds") )

ex_diff <- readRDS(here::here("Data", "exclusiveJourney_timeDiff.rds") )
cj_diff <- readRDS(here::here("Data", "concurrentJourney_timeDiff.rds") )
cjEmplids <- readRDS(here::here("Data", "Journey Activity from Brute Force Fit Approximation for Concurrent Journey Employees.rds"))
univBound <- readRDS(here::here("Data", "Brute Force Fit Approximation for Concurrent Journey Employees.rds") )


source(here::here("Functions", "Turnover Functions.R"))
source(here::here("Functions", "Brute Force Functions.R"))


library(lubridate)
library(RColorBrewer)

```

```{r}

###############
## FUNCTIONS ##
###############

plotJourney <- function(data, plotMap){
  
  # where data is the journeyData for a single EMPLID
  # where plotMap is the actionFrame or actionReasonFrame with color, shape, and size specified
  
  # merge plotMap if necessary
  
  if(!all(c("shape_color", "shape_shape", "shape_size") %in% colnames(data))) {
    timeLine <- merge(data, plotMap, by = c("ACTION", "ACTION_REASON") , all.x = TRUE)
  } else {timeLine <- data}
  
  # create jitter
  yPos <- ave(as.numeric(timeLine$EFFDT), timeLine$EFFDT, FUN = function(dates) {
    n <- length(dates)
    if (n == 1) {
      return(1)  # single point: no jitter
    } else {
      # Evenly spaced jitter around y = 1
      jitter_values <- seq(0.9, 1.1, length.out = n)
      return(jitter_values)
    }
  })
  
  # draw the plot
  
  plot(y = yPos + timeLine$EMPL_RCD,
       x = as.Date(timeLine$EFFDT),
       ylab = "Concurrent jobs (EMPL_RCD)",
       xlab = "Date [EFFDT]",
       pch = timeLine[,"shape_shape"],
       col = timeLine[,"shape_color"],
       cex = timeLine[,"shape_size"]
  )
  
  # legend
  # Legend might need to be an entire separate graphic
  # legend("topright",
  #       title = "Boundary type (size)",
  #       legend = c("primary, university", "primary" , "break", "leave" ),
  #       pch = c(1, 1, 1,1),
  #       cex = c(2, 0.75, 1,1)
  #        )
  
  
  
}



plotMetrics_univ <- function(data,
                        plotList = "all",
                        featureMap = NA,
                        cumulative_plot_params = list(),
                        cumulative_points_params = list(),
                        cumulative_legend_params = list(),
                        cumulative_mtext_params = list(),
                        cumulative_rect_args = list(),
                        cumulative_grid_args = list(),
                        delta_plot_params = list(),
                        delta_points_params = list(),
                        delta_mtext_params = list(),
                        delta_upper_rect_args = list(),
                        delta_lower_rect_args = list(),
                        delta_grid_args = list(),
                        delta_legend_params = list(),
                        metric_plot_params = list(),
                        hire_points_params = list(),
                        term_points_params = list(),
                        metric_legend_params = list(),
                        metric_legend2_params = list(),
                        metric_mtext_params = list(),
                        metric_rect_args = list(),
                        metric_grid_args = list(),
                        title_mtext_params = list()
){
  
  # This accepts the output of "deltaHeadCount" or "calculateMetrics" 
  # and creates several graphics of different metrics over time.
  
  # NOTES FOR USING:
  # If you don't want to show both start and stop lines, then set the argument "type='n'"
  #  in the appropriate params list (but it would still show up in the legend.  Hmm.)
  # If you don't want to show a particular legend, then set the argument "plot = FALSE"
  #  in the appropriate params list.
  
  # If you don't want to show the headcount or cumulative graphic, you may need to
  # set oma = c(0,0,2,0) in metric_plot_params or delta_plot_params to provide space
  # for the title.
  
  # Values for plotList are c("all", "headcount", "cumulative", "rate", "count", "delta.count", "delta.rate")
  
  # Don't forget you can change the name displayed in the legends, which is especially useful
  # if you are entering a single data frame that doesn't have a name:
  # cumulative_legend_params = list(legend = "Frankfurt")
  
  # Here are some example calls:
  # plotMetrics(complexMetrics[c("unassigned")], 
  #             plotList = c("count", "delta.count"), 
  #             delta_legend_params = list(plot = FALSE), 
  #             term_points_params = list(type = "b", lty = 2), 
  #             featureMap = complexClusterColors, 
  #             metric_plot_params = list(oma=c(0,0,2,0)), 
  #             metric_legend_params = list(x="bottom", pch = c(NA,1),  lty = c(3,3), pt.cex = 1.2))
  # plotMetrics(complexMetrics[c(2,4)], 
  #              plotList = c("headcount", "rate", "delta.rate"), 
  #              delta_legend_params = list(plot = FALSE), 
  #              term_points_params = list(type = "b", lty = 2), 
  #              featureMap = complexClusterColors, 
  #              metric_legend_params = list(x="top", pch = c(NA,1),  lty = c(3,3), pt.cex = 1.2))
  
  #  plotMetrics(collegeMetrics[c("Med", "Engr", "Hunt","Science")], 
  #                             plotList = c("rate", "delta.rate"), 
  #                             metric_plot_params = list(oma = c(0,0,2,0)),
  #                             delta_legend_params = list(plot = FALSE), 
  #                             term_points_params = list(type = "b", lty = 2), 
  #                             featureMap = NA, 
  #                             metric_legend_params = list(x="top", pch = c(NA,1),  lty = c(3,3), pt.cex = 1.2))
  
  #  plotMetrics(collegeMetrics[2:11], 
  #                             plotList = c("count", "delta.count"), 
  #                             metric_plot_params = list(oma = c(0,0,2,0)),
  #                             delta_legend_params = list(plot = FALSE), 
  #                             term_points_params = list(type = "b", lty = 2),
  #                             hire_points_params = list(type = "n"),
  #                             featureMap = NA, 
  #                             metric_legend_params = list(x="top", pch = c(NA,1),  lty = c(3,3), pt.cex = 1.2), 
  #                             metric_legend2_params = list(x="topleft"))
  
  # POSSIBLE ADJUSTMENTS:
  # I could explicitly label points like starting date and concluding headcount values 
  #    with text on the graphic
  
  ################################
  ## MANAGE INCOMING PARAMETERS ##
  ################################
  
  incoming.par <- par(fg="red") # dummy adjustment for ease of coding because par() likes adjustments
  on.exit(layout(matrix(1,1,1)))
  on.exit(par(incoming.par), add = TRUE)
  
  
  ###############################
  ## MANAGE INCOMING ARGUMENTS ##
  ###############################
  
  if(is.data.frame(data)) {data <- list(data)} # convert to a data frame
  
  if(is.null(names(data))) {  # assign default names
    
    names(data) <- paste("data", 1:length(data), sep = " ")
    
  }
  
  if(length(data) > 10 && all(is.na(featureMap))  ) { 
    stop("Add featureMap or reduce list length to <=10")
  }  
  
  if(all(is.na(featureMap))) {
    featureMap <- c("sienna",
                    "forestgreen",
                    "deepskyblue",
                    "goldenrod",
                    "firebrick",
                    "darkslategray",
                    "chartreuse",
                    "slateblue",
                    "darkkhaki",
                    "coral")
    
    if(all(!is.na(names(data)))){
      names(featureMap) <- names(data)
    }
    
  }
  
  # Filter out unknown metrics
  plotList <- plotList[plotList %in% c("all", "headcount", "cumulative", "rate", "count", "delta.count", "delta.rate")]
  if(length(plotList) == 0) { plotList <- "all"}
  
  if("all" %in% plotList){ plotList <- c("cumulative", "rate", "delta.rate")  }
  
  if(all(c("rate","count") %in% plotList )) { 
    stop("Select 'rate' or 'count' in plotList argument.")
  }
  
  if(all(c("delta.rate","delta.count") %in% plotList )) { 
    stop("Select 'delta.rate' or 'delta.count' in plotList argument.")
  }
  
  if(!any(c("all", "headcount", "cumulative", "rate", "count", "delta.count", "delta.rate") %in% plotList )) { 
    stop("Select some of c('all', 'headcount', 'cumulative', 'rate', 'count', 'delta.count', 'delta.rate') in plotList argument.")
  }
  
  ############
  ## LAYOUT ##
  ############
  
  if(any(grepl("rate", tolower(colnames(data[[1]])))) # has the necessary columns
     & (
       any(c("headcount", "cumulative") %in% plotList) &&
       any(c("rate", "count") %in% plotList) &&
       any(c("delta.rate", "delta.count") %in% plotList)
     )
  ) {
    
    layout(matrix(1:3, nrow = 3), heights = c(0.456, 0.281, 0.263))
    
  } else if( 
    length(plotList) > 1  # if there's more than one plot
  ) {
    
    layout(matrix(1:2, nrow = 2), heights = c(0.618, 0.382))
    
  } else {
    
    layout(matrix(1:1, nrow = 1), heights = c(1))
    
  }
  
  
  #####################
  ## CUMULATIVE PLOT ##
  #####################
  
  if( any(c("cumulative", "headcount") %in% plotList)  ){
    
    
    if(any(c("rate", "count", "delta.rate","delta.count") %in% plotList)){ # include the x axis if there's no metrics or delta plot
      
      xAxt <- "n"
      default_cumulative_plot_params <- list(
        oma = c(0,0,2,0),
        mar = c(0,6,0,1),
        bg = "ivory",
        fg = "grey30")
      
    } else { 
      
      xAxt <- "s" 
      default_cumulative_plot_params <- list(
        oma = c(0,0,2,0),
        mar = c(4,6,0,1),
        bg = "ivory",
        fg = "grey30")
    } 
    
    cumulative_plot_params <- modifyList(default_cumulative_plot_params,
                                         cumulative_plot_params)
    
    do.call(par, cumulative_plot_params)
    
    # set yLim
    yLim <- range(sapply(data, function(x){range(x[,"delta.cum"])}))
    
    plot(y = data[[1]][,"delta.cum"],
         x = data[[1]][,"periodEnd"],
         ylim = yLim,
         type = "n",
         xlab = "",
         xaxt = xAxt, #"n",
         las = 1,
         ylab = ""
    )
    
    # Rectangle (plot color) (Establish after plot is drawn)
    
    default_cumulative_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray95", 
      border = NA
    )
    
    cumulative_rect_args <- modifyList(default_cumulative_rect_args, cumulative_rect_args)
    
    # Grid (Establish after the plot is drawn)
    
    default_cumulative_grid_args <- list(col = "gray100", lwd = 2, lty = "dotted")
    cumulative_grid_args <- modifyList(default_cumulative_grid_args, cumulative_grid_args)
    
    # Draw rectangle and grid
    do.call("rect", cumulative_rect_args)
    
    # experimenting with the grid
    grid_y <- axTicks(2)
    grid_x <- pretty(data[[1]][,"periodEnd"], n = 5)
    
    do.call(abline, c(list(h=grid_y), cumulative_grid_args))
    do.call(abline, c(list(v=grid_x), cumulative_grid_args))
    
    # draw the lines using Map
    invisible(Map(function(df, col) {
      do.call(points, modifyList(
        list(
          y = df[,"delta.cum"],
          x = df[,"periodEnd"],
          type = "l",
          col =col
          #col = ifelse(length(data) == 1, "sienna",col)
        ),
        cumulative_points_params
      ))
    }, data, featureMap[names(data)]))
    
    # margin text
    
    if(any(grepl("rate", tolower(colnames(data[[1]]))))){ 
      
      default_cumulative_mtext_params <- list(
        side = 2,
        line = 4,
        text = "head count")
      
    } else {
      
      default_cumulative_mtext_params <- list(
        side = 2,
        line = 4,
        text = "cumulative")
      
    }
    
    cumulative_mtext_params <- modifyList(default_cumulative_mtext_params, 
                                          cumulative_mtext_params)
    
    do.call(mtext, cumulative_mtext_params)
    
    # legend
    
    default_cumulative_legend_params <- list(
      x = "topleft",
      legend = names(data),
      col = featureMap[names(data)],
      pch = 15, 
      pt.cex = 2
    )
    
    cumulative_legend_params <- modifyList(default_cumulative_legend_params, cumulative_legend_params)
    
    do.call(legend, cumulative_legend_params)
    
    
  }
  
  
  
  ##################
  ## METRICS PLOT ##
  ##################
  
  if(any(grepl("rate", tolower(colnames(data[[1]])))) &
     any(c("rate", "count") %in% plotList)
  ) {
    
    if("rate" %in% plotList && !("count" %in% plotList) ) {
      
      hireVal <- 100*data[[1]][,"startRate"]
      termVal <- 100*data[[1]][,"stopRate"]
      
      
      yLim <- range(unlist(
        lapply(data, function(df) {
          apply(df[, c("startRate", "stopRate")], 1, range, na.rm = TRUE)
        })
      ), na.rm = TRUE) * 100
      
      y_label <- "rates (%)"
      legendText <- c("start", "stop")
      
    }
    
    if("count" %in% plotList && !("rate" %in% plotList) ) {
      hireVal <- data[[1]][,"start"]
      termVal <- data[[1]][,"stop"] 
      
      yLim <- range(unlist(
        lapply(data, function(df) {
          apply(df[, c("start", "stop")], 1, range, na.rm = TRUE)
        })
      ), na.rm = TRUE)    
      
      y_label <- "count"
      legendText <- c("start", "stop")
    }
    
    if(any(c("delta.rate","delta.count") %in% plotList)){ # include the x axis if there's no delta plot
      
      xAxt <- "n"
      default_metric_plot_params <- list(
        # oma = c(0,0,2,0),
        mar = c(0,6,0,1),
        bg = "ivory",
        fg = "grey30")
      
    } else { 
      
      xAxt <- "s" 
      default_metric_plot_params <- list(
        # oma = c(0,0,2,0),
        mar = c(4,6,0,1),
        bg = "ivory",
        fg = "grey30")
    } 
    
    
    
    metric_plot_params <- modifyList(default_metric_plot_params,
                                     metric_plot_params)
    
    do.call(par, metric_plot_params)
    
    # empty plot
    
    plot(y= hireVal, #100*data[,"startRate"],
         x= data[[1]][,"periodEnd"],
         ylim = yLim,
         type = "n",
         xlab = "",
         xaxt = xAxt, # "n",
         # col = "seagreen3",
         las = 2,
         ylab = "" # "delta"
    )
    
    # Rectangle (plot color) (Establish after plot is drawn)
    
    default_metric_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray95", 
      border = NA
    )
    
    metric_rect_args <- modifyList(default_metric_rect_args, metric_rect_args)
    
    # Grid (Establish after the plot is drawn)
    
    default_metric_grid_args <- list(col = "gray100", lwd = 2, lty = "dotted")
    metric_grid_args <- modifyList(default_metric_grid_args, metric_grid_args)
    
    # Draw rectangle
    do.call("rect", metric_rect_args)
    
    # Draw the grid
    grid_y <- axTicks(2)
    grid_x <- pretty(data[[1]][,"periodEnd"], n = 5)
    
    do.call(abline, c(list(h=grid_y), metric_grid_args))
    do.call(abline, c(list(v=grid_x), metric_grid_args))
    
    # Draw the points  
    
    if("rate" %in% plotList && !("count" %in% plotList)  ) {
      
      # draw the hire points using Map
      invisible(Map(function(df, col) {
        do.call(points, modifyList(
          list(
            y = 100*df[,"startRate"],
            x = df[,"periodEnd"],
            type = "l",
            lty =3,
            col = ifelse(length(data) == 1, "darkcyan",col)
          ),
          hire_points_params
        ))
      }, data, featureMap[names(data)]))  
      
      # draw the termination points using Map
      invisible(Map(function(df, col) {
        do.call(points, modifyList(
          list(
            y = 100*df[,"stopRate"],
            x = df[,"periodEnd"],
            type = "l",
            col = ifelse(length(data) == 1, "coral",col)
          ),
          term_points_params
        ))
      }, data, featureMap[names(data)]))  
      
    }
    
    if("count" %in% plotList && !("rate" %in% plotList) ) {
      
      # draw the hire lines using Map
      invisible(Map(function(df, col) {
        do.call(points, modifyList(
          list(
            y = df[,"start"],
            x = df[,"periodEnd"],
            type = "l",
            lty = 3,
            col = ifelse(length(data) == 1, "darkcyan",col)
          ),
          hire_points_params
        ))
      }, data, featureMap[names(data)]))
      
      # draw the termination lines using Map
      invisible(Map(function(df, col) {
        do.call(points, modifyList(
          list(
            y = df[,"stop"],
            x = df[,"periodEnd"],
            type = "l",
            col = ifelse(length(data) == 1, "coral",col)
          ),
          term_points_params
        ))
      }, data, featureMap[names(data)]))
      
    }
    
    
    # Margin text
    
    default_metric_mtext_params <- list(side = 2,
                                        line = 4,
                                        text = y_label 
    )
    metric_mtext_params <- modifyList(default_metric_mtext_params, 
                                      metric_mtext_params)
    
    do.call(mtext, metric_mtext_params)
    
    # legend
    
    if (length(data) == 1) {
      default_metric_legend_params <- list(
        x = "topleft",
        legend = legendText, 
        col = c("darkcyan","coral"),
        lty = c(3,1),
        lwd = c(1.5,3)
      )
    } else {
      
      default_metric_legend_params <- list(
        x = "topleft",
        legend = legendText, 
        col = c("gray50","gray50"),
        lty = c(3,1),
        lwd = c(1.5,3)
      ) 
      
    }
    
    metric_legend_params <- modifyList(default_metric_legend_params, metric_legend_params)
    
    do.call(legend, metric_legend_params)
    
    
    # Display legend of colors per featuremap on the right
    
    if(length(data) == 1 ) {
      
      default_metric_legend2_params <- list(
        x = "topright",
        legend = names(data),
        #  col = featureMap[names(data)],
        #  pch = 15, 
        #  pt.cex = 2,
        plot = FALSE
      )
      
    } else {
      
      default_metric_legend2_params <- list(
        x = "topright",
        legend = names(data), 
        col = featureMap[names(data)],
        pch = 15, 
        pt.cex = 2
      ) 
    }
    
    
    metric_legend2_params <- modifyList(default_metric_legend2_params, metric_legend2_params)
    
    do.call(legend, metric_legend2_params)
    
    
  }
  
  
  ################
  ## DELTA PLOT ##
  ################
  
  if("delta.count" %in% plotList && !("delta.rate" %in% plotList) ) {
    yVal <- data[[1]][,"delta"]; 
    y_label <- "delta count"
    yLim <- range(sapply(data, function(x){range(x[,"delta"])}))
    
  }
  
  if("delta.rate" %in% plotList && !("delta.count" %in% plotList) ) {
    
    yVal <- 100*data[[1]][,"deltaRate"]; 
    y_label <- "delta (%)"
    yLim <- 100*range(sapply(data, function(x){range(x[,"deltaRate"])}))
    
  }
  
  if(any(c("delta.count","delta.rate") %in% plotList)) {
    
    default_delta_plot_params <- list(mar = c(4,6,0,1),
                                      #bg = "ivory",
                                      fg = "grey30")
    
    delta_plot_params <- modifyList(default_delta_plot_params,
                                    delta_plot_params)
    
    do.call(par, delta_plot_params)
    
    # Empty plot
    
    plot(y= yVal, 
         x= data[[1]][,"periodEnd"],
         type = "n",
         ylim = yLim,
         xlab = "",
         las = 2,
         ylab = "" 
    )
    
    # Rectangle (plot color) (Establish after plot is drawn)
    
    default_delta_upper_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = 0, #par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = adjustcolor("palegreen3", alpha.f = 0.1), 
      border = NA
    )
    
    delta_upper_rect_args <- modifyList(default_delta_upper_rect_args, delta_upper_rect_args)
    
    default_delta_lower_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = 0, #par("usr")[4],
      col = adjustcolor("pink", alpha.f = 0.5), 
      border = NA
    )
    
    delta_lower_rect_args <- modifyList(default_delta_lower_rect_args, delta_lower_rect_args)
    
    
    # Grid (Establish after the plot is drawn)
    default_delta_grid_args <- list(col = "lightgray", lwd = 1, lty = "dotted")
    delta_grid_args <- modifyList(default_delta_grid_args, delta_grid_args)
    
    # Draw rectangle and grid
    do.call("rect", delta_upper_rect_args)
    do.call("rect", delta_lower_rect_args)
    
    # Making grid work
    grid_y <- axTicks(2)
    grid_x <- pretty(data[[1]][,"periodEnd"], n = 5)
    
    do.call(abline, c(list(h=grid_y), delta_grid_args))
    do.call(abline, c(list(v=grid_x), delta_grid_args))
    
    
    # draw the delta lines using Map
    
    if("delta.count" %in% plotList && !("delta.rate" %in% plotList) ) {
      
      invisible(Map(function(df, col) {
        do.call(points, modifyList(
          list(
            y = df[,"delta"],
            x = df[,"periodEnd"],
            type = "l",
            col = col
          ),
          delta_points_params
        ))
      }, data, featureMap[names(data)]))
      
    }
    
    
    if("delta.rate" %in% plotList && !("delta.count" %in% plotList) ) {
      
      invisible(Map(function(df, col) {
        do.call(points, modifyList(
          list(
            y = 100*df[,"deltaRate"],
            x = df[,"periodEnd"],
            type = "l",
            col = col
          ),
          delta_points_params
        ))
      }, data, featureMap[names(data)]))
      
    }
    
    # Margin text
    
    default_delta_mtext_params <- list(side = 2,
                                       line = 4,
                                       text = y_label
    )
    delta_mtext_params <- modifyList(default_delta_mtext_params, 
                                     delta_mtext_params)
    
    do.call(mtext, delta_mtext_params)
    
  }
  
  
  default_delta_legend_params <- list(
    x = "topleft",
    legend = names(data),
    col = featureMap[names(data)],
    pch = 15, 
    pt.cex = 2
  )
  
  delta_legend_params <- modifyList(default_delta_legend_params, delta_legend_params)
  
  do.call(legend, delta_legend_params)
  
  ################
  ## OUTER TEXT ##
  ################
  
  titleText <- "Head count metrics"
  
  default_title_mtext_params <- list(text = titleText, 
                                     side = 3,
                                     line = 0.3,
                                     font =2, 
                                     cex = 1.3, 
                                     outer = TRUE)
  
  
  title_mtext_params <- modifyList(default_title_mtext_params, title_mtext_params)
  
  do.call(mtext, title_mtext_params)
  
  
  
  
  
}


calculateMetrics_univ <- function(initial_count=NA, 
                             initial_date=NA,
                             calendar = "day",
                             minDate = ymd(paste(year(today()), "01","01", sep = "-")),
                             maxDate = today(),
                             data
){
  
  # This calculates various HR metrics: the counts and rates of people entering, stoping, and the net change
  # (called "delta" here.)
  
  # It uses deltaHeadCount_univ.  Because deltaHeadCount_univ always starts at a value of zero, this function
  # accepts or creates an initial value to result in an accurate current head count.
  
  # It accepts either an initial_count, or an initial_date.
  
  # If neither is provided, then it finds the earliest date in the data set as the initial_date.
  
  # If the initial_count is provided, then it uses that. If it is not provided, then a value for 
  # initial_count is calculated using initial_date in deltaHeadCount_univ.  It uses the initial_date as the value for the 
  # minDate argument and the maximum date before the start of the desired period as 
  # the maxDate argument.  The final "delta.cum" result is used as the initial_count.
  
  # Then, either accepting the provided initial_count or calculating it, it uses deltaHeadCount_univ 
  # to calculate the counts of people entering, stoping, and the net change (called "delta" here.)
  
  # What makes this an interesting wrapper to deltaHeadCount_univ is that it will count the daily change,
  # and then aggregate that to a mean value over the desired period.  This mean is the denominator
  # for the rate calculations.
  
  #  calculateTurnover (or calculateMetrics) and deltaHeadCount_univ will provide identical results if the minDate used
  #  for deltaHeadCount_univ is the same as the initial_date used in calculateTurnover, and the other arguments
  #  are identical.
  
  
  # calculate initial_count
  
  if(is.na(initial_count) & is.na(initial_date)){
    
    # set initial date to the earliest date in the data set
    initial_date <-  data[,"EFFDT"] |>
      min() |>
      ymd()
    
  }
  
  
  if(is.na(initial_count)){
    
    # Establish maximum date for the period that concludes before the first period of the minDate
    
    
    if(calendar == "day"){initial_max <- minDate-1 }
    
    if(calendar == "week"){
      initial_max <- minDate
      while(isoweek(initial_max) == isoweek(minDate)) {initial_max = initial_max -1} 
    }
    
    if(calendar == "month"){
      initial_max <- minDate
      while(month(initial_max) == month(minDate)) {initial_max = initial_max -1} 
    }
    
    if(calendar == "quarter"){
      initial_max <- minDate
      while(quarter(initial_max) == quarter(minDate)) {initial_max = initial_max -1} 
    }
    
    if(calendar == "year"){
      initial_max <- minDate
      while(year(initial_max) == year(minDate)) {initial_max = initial_max -1} 
    }
    
    # calculate a value for the initial count
    
    intermediate <- deltaHeadCount_univ(
      minDate = initial_date,
      maxDate = initial_max,
      calendar = "day",
      data = data
    ) 
    
    initial_count <-  intermediate |>
      (\(x){tail(x[,"delta.cum"],1) })()
    
  }
  
  # Calculate the foundation deltaHeadCount_univ
  
  foundation <- deltaHeadCount_univ(minDate = minDate,
                               maxDate = maxDate,
                               calendar = calendar,
                               initial_count = initial_count,
                               data = data)
  
  
  # Calculate mean headcount by first repeating deltaHeadCount_univ by day
  
  meanHeadCount <- deltaHeadCount_univ(minDate = minDate,
                                  maxDate = maxDate,
                                  calendar = "day",
                                  initial_count = initial_count,
                                  data = data)  
  
  # aggregate the mean headcount per period
  
  if(calendar == "day") {
    
    # for consistency in flow; these should essentially do nothing
    periodHeadCountMean <- aggregate(delta.cum ~ EFFDT, data = meanHeadCount, mean)
  }
  
  if(calendar == "week") {
    periodHeadCountMean <- aggregate(delta.cum ~ paste(year(EFFDT), sprintf("%02d", isoweek(EFFDT)), sep = "-W"), data = meanHeadCount, mean)
  }
  
  if(calendar == "month") {
    periodHeadCountMean <- aggregate(delta.cum ~ format(EFFDT, "%Y-%m"), data = meanHeadCount, mean)
  }
  
  if(calendar == "quarter") {
    periodHeadCountMean <- aggregate(delta.cum ~ paste(year(EFFDT), quarter(EFFDT), sep = "-Q"), data = meanHeadCount, mean)
  }
  
  if(calendar == "year") {
    periodHeadCountMean <- aggregate(delta.cum ~ year(EFFDT), data = meanHeadCount, mean)
  }
  
  
  # fix names
  names(periodHeadCountMean)[grepl("EFFDT", names(periodHeadCountMean))] <- "adjDate"
  names(periodHeadCountMean)[names(periodHeadCountMean) == "delta.cum"] <- "headcount_mean"
  
  # merge mean headcount back into the foundation
  foundation <- merge(foundation, periodHeadCountMean, by = "adjDate", sort = FALSE)
  
  # calculate metrics
  foundation$startRate <- foundation$start/foundation$headcount_mean
  foundation$stopRate <- foundation$stop/foundation$headcount_mean
  foundation$deltaRate <- foundation$delta/foundation$headcount_mean
  
  
  return(foundation)  
  
}

deltaHeadCount_univ <- function(minDate, 
                           maxDate,
                           calendar = "day",
                           initial_count = 0,
                           data) {
  
  # where data is journeyData or retData
  # where minDate and maxDate are self-explanatory and work best as class "Date", or 
  # as ymd("2021-5-31")
  # where calendar is one of "day", "week", "month", "quarter", and "year".
  # where initial_count is mostly for use with calculateTurnover,
  # and represents the headcount on minDate.
  # if initial_count is zero, then the delta function will always start at zero
  
  # This is not adapted to work in fiscal years.
  
  # Working with dates has some curious behavior.
  # For one, when I have a period other than "day" then I expand
  # the aggregation period to "contain" the date range.
  # For example, if one of my dates is June 7th and the period is "month",
  # then I expand to use all of June.  The adjusted date becomes June 1st if
  # it's the minDate, and June 30th if it's the maxDate.
  
  # For two, I use "isoweek".  If I specify the calendar as "week", then I 
  # bound the minDate-maxDate range with the isoweek that contains the minDate 
  # and the maxDate.
  
  # This has the strange behavior that the isoweek doesn't align with months,
  # or quarters, or years.  This makes direct comparison between weeks and 
  # everything else impossible, as the adjusted start dates and end dates
  # will be different.
  
  # I initially tried using "isoyear", but it got really tricky if my minDate was 2019-12-31, as the isoyear for this
  # was actually 2020.  I tried dealing with this in several ways that were
  # various levels of silly and undesirable.  So I abandoned using isoyears.
  
  # Add a dummy column for aggregation
  
  data[,"one"] <- 1 # because there is one PI per row
  
  if(calendar == "day") {
    
    # create data frame with one row per calendar period
    hrDates <- data.frame(EFFDT = seq(from = minDate, to = maxDate, by = calendar)) 
    
    # maintain consistency with the other time periods
    hrDates$adjDate <- hrDates$EFFDT
    
    # Aggregate hire and termination actions
    theActions <- aggregate(one ~ univ_boundary+EFFDT, data = data, sum)
    
    # prepare for merge
    # theActions$EFFDT <- as.Date(theActions$EFFDT)
    # names(theActions)[names(theActions) == "EFFDT"] <- "adjDate"
    theActions$adjDate <- as.Date(theActions$EFFDT)
    
  }    
  
  if(calendar == "week") {
    
    # discover the extreme dates of the isoweek
    weekMin <- minDate
    while(isoweek(weekMin) == isoweek(minDate) ) {weekMin <- weekMin - 1 }
    weekMin <- weekMin+1
    
    weekMax <- maxDate
    while(isoweek(weekMax) == isoweek(maxDate) ) {weekMax <- weekMax + 1 }
    weekMax <- weekMax-1
    
    # create data frame with one row per calendar period
    hrDates <- data.frame(EFFDT = seq(from = weekMin, to = weekMax, by = calendar)) 
    
    # convert to ISO standard format
    hrDates$adjDate <- paste(isoyear(hrDates$EFFDT), sprintf("%02d", isoweek(hrDates$EFFDT)), sep = "-W")
    
    # aggregate
    theActions <- aggregate(one ~ univ_boundary+paste(isoyear(EFFDT), sprintf("%02d", isoweek(EFFDT)), sep = "-W"), data = data, sum)
    names(theActions)[!names(theActions) %in% c("univ_boundary", "one")] <- "adjDate"
    
  }
  
  if(calendar == "month") {
    
    monthMin <- ymd(paste(year(minDate), month(minDate),"01", sep = "-") )
    monthMax <- ymd(paste(year(maxDate), month(maxDate),"01", sep = "-") )
    
    # create data frame with one row per calendar period
    hrDates <- data.frame(EFFDT = seq(from = monthMin, to = monthMax, by = calendar)) 
    
    # convert to ISO standard format
    hrDates$adjDate <- format(hrDates$EFFDT, "%Y-%m") 
    
    
    # aggregate
    theActions <- aggregate(one ~ univ_boundary + format(EFFDT, "%Y-%m"), data = data, sum)
    names(theActions)[!names(theActions) %in% c("univ_boundary", "one")] <- "adjDate"
  }
  
  if(calendar == "quarter"){
    
    # discover the extreme dates of the quarter
    quarterMin <- minDate
    while(quarter(quarterMin) == quarter(minDate) ) {quarterMin <- quarterMin - 1 }
    quarterMin <- quarterMin+1
    
    quarterMax <- maxDate
    while(quarter(quarterMax) == quarter(maxDate) ) {quarterMax <- quarterMax + 1 }
    quarterMax <- quarterMax-1    
    
    # create data frame with one row per calendar period
    hrDates <- data.frame(EFFDT = seq(from = quarterMin, to = quarterMax, by = calendar)) 
    
    # convert to quarter format
    hrDates$adjDate <- paste(year(hrDates$EFFDT), quarter(hrDates$EFFDT), sep = "-Q")
    
    # aggregate
    theActions <- aggregate(one ~ univ_boundary+paste(year(EFFDT), quarter(EFFDT), sep = "-Q"), data = data, sum)
    names(theActions)[!names(theActions) %in% c("univ_boundary", "one")] <- "adjDate"
    
  }
  
  if(calendar == "year") {
    
    yearMin <- ymd(paste(year(minDate), "01","01", sep = "-") )
    yearMax <- ymd(paste(year(maxDate), "12","31", sep = "-") )
    
    # create data frame with one row per calendar period
    hrDates <- data.frame(EFFDT = seq(from = yearMin, to = yearMax, by = calendar)) 
    
    # convert to desired format
    hrDates$adjDate <- year(hrDates$EFFDT) 
    
    # aggregate
    theActions <- aggregate(one ~ univ_boundary+year(EFFDT), data = data, sum)
    names(theActions)[!names(theActions) %in% c("univ_boundary", "one")] <- "adjDate"
  }
  
  # merge 
  
  hrDates <- merge(hrDates, theActions[theActions$univ_boundary == "stop",-which(colnames(theActions) %in% c("univ_boundary", "EFFDT"))], by = "adjDate", all.x = TRUE, sort=FALSE)
  names(hrDates)[names(hrDates) == "one"] <- "stop"
  hrDates <- merge(hrDates, theActions[theActions$univ_boundary == "start",-which(colnames(theActions) %in% c("univ_boundary", "EFFDT"))], by = "adjDate", all.x = TRUE, sort=FALSE)
  names(hrDates)[names(hrDates) == "one"] <- "start"
  
  # restore chronological order 
  hrDates <- hrDates[order(hrDates$adjDate),] # uh oh.  I need to keep the EFFDT,  I guess
  
  # convert NA values to zero
  hrDates[is.na(hrDates)] <- 0
  
  # calculate delta
  hrDates$delta <- hrDates$start - hrDates$stop
  
  # calculate delta cumulative
  hrDates$delta.cum <- cumsum(hrDates$delta) + initial_count
  
  # add a "periodEnd" column for clarity in plotting
  if(calendar == "week"){ 
    hrDates$periodEnd <- ceiling_date(hrDates$EFFDT, unit = calendar, week_start = 1) - days(1)
  } else {
    hrDates$periodEnd <- ceiling_date(hrDates$EFFDT, unit = calendar) - days(1)
  }
  
  
  return(hrDates)
  
  
  ###############
  ## OLD LOGIC ##
  ###############
  
  # old logic for backwards compatibility:
  
  if(!("EFFDT" %in% colnames(data)) & all(c("HIRE_DT","TERMINATION_DT") %in% colnames(data))) {
    
    # Add a dummy column for aggregation
    
    data[,"one"] <- 1 # because there is one PI per row
    
    if(calendar == "day") {
      
      # create data frame with one row per calendar period
      hrDates <- data.frame(actionDate = seq(from = minDate, to = maxDate, by = calendar)) 
      
      # duplicate column name to parallel other calendar periods
      hrDates$adjDate <- hrDates$actionDate
      
      # Aggregate hire and termination actions
      hireActions <- aggregate(one ~ HIRE_DT, data = data, sum)
      names(hireActions) <- c("adjDate","one")
      termActions <- aggregate(one ~ TERMINATION_DT, data = data, sum)
      names(termActions) <- c("adjDate","one")
      
      # prepare for merge
      hireActions$adjDate <- as.Date(hireActions$adjDate)
      termActions$adjDate <- as.Date(termActions$adjDate)
    }
    
    if(calendar == "week") {
      
      # discover the extreme dates of the isoweek
      weekMin <- minDate
      while(isoweek(weekMin) == isoweek(minDate) ) {weekMin <- weekMin - 1 }
      weekMin <- weekMin+1
      
      weekMax <- maxDate
      while(isoweek(weekMax) == isoweek(maxDate) ) {weekMax <- weekMax + 1 }
      weekMax <- weekMax-1
      
      # create data frame with one row per calendar period
      hrDates <- data.frame(actionDate = seq(from = weekMin, to = weekMax, by = calendar)) 
      
      # convert to ISO standard format
      hrDates$adjDate <- paste(isoyear(hrDates$actionDate), sprintf("%02d", isoweek(hrDates$actionDate)), sep = "-W")
      # hrDates$adjDate <- paste(year(hrDates$actionDate), isoweek(hrDates$actionDate), sep = "-W")
      
      # aggregate
      hireActions <- aggregate(one ~ paste(isoyear(HIRE_DT), sprintf("%02d", isoweek(HIRE_DT)), sep = "-W"), data = data, sum)
      names(hireActions) <- c("adjDate","one")
      termActions <- aggregate(one ~ paste(isoyear(TERMINATION_DT), sprintf("%02d", isoweek(TERMINATION_DT)), sep = "-W"), data = data, sum)
      names(termActions) <- c("adjDate","one")
      
    }
    
    if(calendar == "month") {
      
      monthMin <- ymd(paste(year(minDate), month(minDate),"01", sep = "-") )
      monthMax <- ymd(paste(year(maxDate), month(maxDate),"01", sep = "-") )
      
      # create data frame with one row per calendar period
      hrDates <- data.frame(actionDate = seq(from = monthMin, to = monthMax, by = calendar)) 
      
      # convert to ISO standard format
      hrDates$adjDate <- format(hrDates$actionDate, "%Y-%m") 
      
      # aggregate
      hireActions <- aggregate(one ~ format(HIRE_DT, "%Y-%m"), data = data, sum)
      names(hireActions) <- c("adjDate","one")
      termActions <- aggregate(one ~ format(TERMINATION_DT, "%Y-%m"), data = data, sum)
      names(termActions) <- c("adjDate","one")
      
    }
    
    if(calendar == "quarter"){
      
      # discover the extreme dates of the quarter
      quarterMin <- minDate
      while(quarter(quarterMin) == quarter(minDate) ) {quarterMin <- quarterMin - 1 }
      quarterMin <- quarterMin+1
      
      quarterMax <- maxDate
      while(quarter(quarterMax) == quarter(maxDate) ) {quarterMax <- quarterMax + 1 }
      quarterMax <- quarterMax-1    
      
      # create data frame with one row per calendar period
      hrDates <- data.frame(actionDate = seq(from = quarterMin, to = quarterMax, by = calendar)) 
      
      # convert to ISO standard format
      hrDates$adjDate <- paste(year(hrDates$actionDate), quarter(hrDates$actionDate), sep = "-Q")
      
      # aggregate
      hireActions <- aggregate(one ~ paste(year(HIRE_DT), quarter(HIRE_DT), sep = "-Q"), data = data, sum)
      names(hireActions) <- c("adjDate","one")
      termActions <- aggregate(one ~ paste(year(TERMINATION_DT), quarter(TERMINATION_DT), sep = "-Q"), data = data, sum)
      names(termActions) <- c("adjDate","one")
      
    }
    
    if(calendar == "year") {
      
      yearMin <- ymd(paste(year(minDate), "01","01", sep = "-") )
      yearMax <- ymd(paste(year(maxDate), "12","31", sep = "-") )
      
      # create data frame with one row per calendar period
      hrDates <- data.frame(actionDate = seq(from = yearMin, to = yearMax, by = calendar)) 
      
      # convert to ISO standard format
      hrDates$adjDate <- year(hrDates$actionDate) 
      
      # aggregate
      hireActions <- aggregate(one ~ year(HIRE_DT), data = data, sum)
      names(hireActions) <- c("adjDate","one")
      termActions <- aggregate(one ~ year(TERMINATION_DT), data = data, sum)
      names(termActions) <- c("adjDate","one")
      
    }
    
    # merge
    hrDates <- merge(hrDates, hireActions, by = "adjDate", all.x = TRUE, sort=FALSE)
    names(hrDates)[names(hrDates) == "one"] <- "hireCount"
    hrDates <- merge(hrDates, termActions, by = "adjDate", all.x = TRUE, sort= FALSE)
    names(hrDates)[names(hrDates) == "one"] <- "termCount"
    
    # restore chronological order 
    # (although I've fixed it be setting sort=FALSE in merge, I'll double-correct)
    hrDates <- hrDates[order(hrDates$actionDate),]
    
    # Calculate delta
    
    # convert NA to zero
    hrDates[is.na(hrDates)] <- 0
    
    # calculate delta
    hrDates$delta <- hrDates$hireCount - hrDates$termCount
    
    # calculate delta cumulative
    hrDates$delta.cum <- cumsum(hrDates$delta) + initial_count
    
    # add a "periodEnd" column for clarity in plotting
    if(calendar == "week"){ 
      hrDates$periodEnd <- ceiling_date(hrDates$actionDate, unit = calendar, week_start = 1) - days(1)
    } else {
      hrDates$periodEnd <- ceiling_date(hrDates$actionDate, unit = calendar) - days(1)
    }
    
    return(hrDates)
    
  }
  
}

```

```{r}

##########
## PREP ##
##########

# flowchart conditionals
journeyPopulation$wb <- journeyPopulation$WORKBREAK == "wb"
journeyPopulation$lv <- journeyPopulation$LEAVE == "leave"

journeyPopulation$PI <- journeyPopulation$EMPLID %in% piEMPLIDs
journeyPopulation$singleJob <- journeyPopulation$HCJ == "not_hcj" & journeyPopulation$SINGLERCD == "single_rcd"



```


**PURPOSE:**


**OBJECTIVES:**

### Population breakdown by journey characteristics 

```{r}

################
## FLOWHCHART ##
################

library(flowchart)

# Showing split between concurrent and non-concurrent jobs

journeyPopulation |>
  as_fc(label = "Workforce population", text_fs = 12) |>
  fc_filter(PI, label = "Principal investigators", show_exc = TRUE) |>
  #fc_split(PI, label = c("Not PI", "PI")) |>
  fc_split(singleJob, label = c("Concurrent jobs", "No concurrent jobs"), bg_fill = "ivory", text_fs = 12) |> 
  fc_draw()


journeyPopulation |>
  (\(x){x[x$PI & !x$singleJob,]})() |>
  as_fc(label = "Concurrent jobs", text_fs = 12) |>
  fc_split(wb, label = c("No work break", "Work break"), bg_fill = "mintcream", text_fs = 12) |>
  fc_split(lv, label = c("No leave", "leave"), bg_fill = "plum1", text_fs = 12) |>
  fc_draw()


journeyPopulation |>
  (\(x){x[x$PI & x$singleJob,]})() |>
  as_fc(label = "Exclusive jobs", text_fs = 12) |>
  fc_split(wb, label = c("No work break", "Work break"), bg_fill = "mintcream", text_fs = 12) |>
  fc_split(lv, label = c("No leave", "leave"), bg_fill = "plum1", text_fs = 12) |>
  fc_draw()

```

### Individual journeys 

#### **EXCLUSIVE JOURNEY**

```{r}



exc_wb_lv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           journeyPopulation$singleJob &
                                           journeyPopulation$wb &
                                           journeyPopulation$lv
                                           ]

exc_nwb_nlv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           journeyPopulation$singleJob &
                                           !journeyPopulation$wb &
                                           !journeyPopulation$lv
                                           ]

exc_wb_nlv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           journeyPopulation$singleJob &
                                           journeyPopulation$wb &
                                           !journeyPopulation$lv
                                           ]

exc_nwb_lv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           journeyPopulation$singleJob &
                                           !journeyPopulation$wb &
                                           journeyPopulation$lv
                                           ]

exclusivePopulation <- list(exc_wb_lv, exc_nwb_nlv, exc_wb_nlv, exc_nwb_lv)

set.seed(99)
excSample <- sapply(exclusivePopulation, function(x){sample(x,1)} )

# Plot it

par(mfrow = c(5,1), mar = c(0,2,2,1))

plotJourneyKey()

for(pi in excSample) {
  
ex_diff[[paste(pi,"0", sep = ".")]] |>
  assignBoundaries() |>
  plotJourney()  
  
}







ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  assignBoundaries() |>
  plotJourney()



```

#### **CONCURRENT JOURNEY**

***No work break, no leave***

```{r}

conc_nwb_nlv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           !journeyPopulation$singleJob &
                                           !journeyPopulation$wb &
                                           !journeyPopulation$lv
                                           ]


# sample(conc_nwb_nlv,1)
selectPI <- "00954937"

par(mfrow= c(2,1))

cjEmplids[[selectPI]] |>
  assignBoundaries() |>
  plotJourney()

cjEmplids[[selectPI]] |>
  forceFitPlot(fitLine = FALSE)

cjEmplids[[selectPI]] |>
  forceFitPlot(fitLine = TRUE)

cjEmplids[[selectPI]] |>
  assignBoundaries() |>
  plotJourney()

addVerticals(univBound[grep(selectPI, row.names(univBound)),])

```



***Work break, no leave***


```{r}

conc_wb_nlv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           !journeyPopulation$singleJob &
                                           journeyPopulation$wb &
                                           !journeyPopulation$lv
                                           ]


# sample(conc_wb_nlv,1)
selectPI <- "00292990"

par(mfrow= c(2,1))

cjEmplids[[selectPI]] |>
  assignBoundaries() |>
  plotJourney()

cjEmplids[[selectPI]] |>
  forceFitPlot(fitLine = FALSE)

cjEmplids[[selectPI]] |>
  forceFitPlot(fitLine = TRUE)

cjEmplids[[selectPI]] |>
  assignBoundaries() |>
  plotJourney()

addVerticals(univBound[grep(selectPI, row.names(univBound)),])

```


***No work break, leave***


```{r}

conc_nwb_lv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           !journeyPopulation$singleJob &
                                           !journeyPopulation$wb &
                                           journeyPopulation$lv
                                           ]


# sample(conc_nwb_lv,1)
selectPI <- "00244272"

par(mfrow= c(2,1))

cjEmplids[[selectPI]] |>
  assignBoundaries() |>
  plotJourney()

cjEmplids[[selectPI]] |>
  forceFitPlot(fitLine = FALSE)

cjEmplids[[selectPI]] |>
  forceFitPlot(fitLine = TRUE)

cjEmplids[[selectPI]] |>
  assignBoundaries() |>
  plotJourney()

addVerticals(univBound[grep(selectPI, row.names(univBound)),])

```


***Work break and leave***  

```{r}

conc_wb_lv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           !journeyPopulation$singleJob &
                                           journeyPopulation$wb &
                                           journeyPopulation$lv
                                           ]


# sample(conc_wb_lv,1)
selectPI <- "00028851"

par(mfrow= c(2,1))

cjEmplids[[selectPI]] |>
  assignBoundaries() |>
  plotJourney()

cjEmplids[[selectPI]] |>
  forceFitPlot(fitLine = FALSE)

cjEmplids[[selectPI]] |>
  forceFitPlot(fitLine = TRUE)

cjEmplids[[selectPI]] |>
  assignBoundaries() |>
  plotJourney()

addVerticals(univBound[grep(selectPI, row.names(univBound)),])

```



### Headcount Metrics 

#### **PI's WITH EXCLUSIVE JOBS**

```{r}


```


#### **PI's WITH CONCURRENT JOBS**

```{r}


univBound |>
  (\(x){ 
    calculateMetrics_univ(data = x)
  })() |> 
  plotMetrics_univ()


```

#### **ALL PI's**

```{r}


```

### Explain HeadCount

```{r}

explainHeadCount()

```  

'Exit' condition needs to be modified to accommodate the transfer to volunteer positions

*Per Brian Gelsinger, Teams message on 14 Oct 2025:*

I can't remember if we included the field UU_BEN_IND in the SQL we provided, but that would be my recommendation to use. A Benefit Indicator of "30 " means that job code is considered "Non-Employee". I'm happy to add that field to the SQL if you'd like, but below is a list of all the job codes in Ben Ind 30, so you could just try excluding those job codes as well. I've also provided the current headcount of employees in these job codes
 
Job Code	Job Title	UU Ben Ind	UU Ben Ind Descr	Count Distinct EE
0233	Field Instructor	30	Non-Employee	15
6000	Volunteer Staff	30	Non-Employee	21
6001	Volunteer Faculty	30	Non-Employee	34
6002	Adjunct Professor	30	Non-Employee	474
6003	Adjunct Associate Professor	30	Non-Employee	371
6004	Adjunct Assistant Professor	30	Non-Employee	748
6005	Adjunct Instructor	30	Non-Employee	588
6006	Shared Faculty (Unpaid)	30	Non-Employee	17
6100	Univ Asia Staff	30	Non-Employee	56
6102	Adjunct Professor (Ext)	30	Non-Employee	134
6103	Adjunct Assoc Professor (Ext)	30	Non-Employee	126
6104	Adjunct Asst Professor (Ext)	30	Non-Employee	784
6105	Adjunct Instructor (Ext)	30	Non-Employee	416
7993	Surviving Spouse	30	Non-Employee	0
9261	COBRA Beneficiary	30	Non-Employee	0
799999	Volunteer Staff	30	Non-Employee	0 

#### **EXCLUSIVE JOURNEY**  

***No work break, no leave***


```{r}

exc_nwb_nlv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           journeyPopulation$singleJob &
                                           !journeyPopulation$wb &
                                           !journeyPopulation$lv
                                           ]


# sample(exc_nwb_nlv,1)
selectPI <- "06025740"

par(mfrow= c(3,1))

plotJourneyKey()

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  assignBoundaries() |>
  plotJourney()

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  forceFitPlot(fitLine = FALSE)

par(mfrow= c(2,1))

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  forceFitPlot(fitLine = TRUE)

plot(x=0.5,
  type = "n",
     ylim = c(0,1),
     xlim = c(0,1),
  ylab = "",
  xlab = "",
     xaxt = "n",
     yaxt = "n")

text(0.5, 0.5, "University boundaries not calculated")


```



***Work break, no leave***


```{r}

exc_wb_nlv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           journeyPopulation$singleJob &
                                           journeyPopulation$wb &
                                           !journeyPopulation$lv
                                           ]


# sample(exc_wb_nlv,1)
selectPI <- "00703457"

par(mfrow= c(2,1))

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  assignBoundaries() |>
  plotJourney()

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  forceFitPlot(fitLine = FALSE)

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  forceFitPlot(fitLine = TRUE)

plot(x=0.5,
  type = "n",
     ylim = c(0,1),
     xlim = c(0,1),
  ylab = "",
  xlab = "",
     xaxt = "n",
     yaxt = "n")

text(0.5, 0.5, "University boundaries not calculated")


```


***No work break, leave***


```{r}

exc_nwb_lv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           journeyPopulation$singleJob &
                                           !journeyPopulation$wb &
                                           journeyPopulation$lv
                                           ]


# sample(exc_nwb_lv,1)
selectPI <- "06022652"

par(mfrow= c(2,1))

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  assignBoundaries() |>
  plotJourney()

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  forceFitPlot(fitLine = FALSE)

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  forceFitPlot(fitLine = TRUE)

plot(x=0.5,
  type = "n",
     ylim = c(0,1),
     xlim = c(0,1),
  ylab = "",
  xlab = "",
     xaxt = "n",
     yaxt = "n")

text(0.5, 0.5, "University boundaries not calculated")


```




***Work break and leave***  


```{r}

exc_wb_lv <- journeyPopulation$EMPLID[journeyPopulation$PI &
                                           journeyPopulation$singleJob &
                                           journeyPopulation$wb &
                                           journeyPopulation$lv
                                           ]


# sample(exc_wb_lv,1)
selectPI <- "06012575"

par(mfrow= c(2,1))

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  assignBoundaries() |>
  plotJourney()

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  forceFitPlot(fitLine = FALSE)

ex_diff[[paste(selectPI,"0", sep = ".")]] |>
  forceFitPlot(fitLine = TRUE)

plot(x=0.5,
  type = "n",
     ylim = c(0,1),
     xlim = c(0,1),
  ylab = "",
  xlab = "",
     xaxt = "n",
     yaxt = "n")

text(0.5, 0.5, "University boundaries not calculated")


```



