---
title: "Retention Exploratory Data Analysis"
author: "Bill Prisbrey"
date: "2025-04-25"
output: html_document
---

Thoughts:

- Re-name this as "exploratory data analysis" and summarize the new data view.  "Skimr" and a couple of histograms.
- I'll probably uncover some data inconsistencies.
- Use my new "activePI" function to calculate the number of active PI's every week.
- Create a dataframe that is week-by-week active PI's
- Use prepData to align publishing periods with hire or re-hire period
- Lack of an initial termination date could be problematic


**PURPOSE:**  The purpose of this document is to breifly calculate and compare principal investigator turnover to determine if this project is worth pursuing. It could be re-named "back of the envelope" or "quick-and-dirty."

**OBJECTIVES:**

  1.  Calculate and compare turnover    
      a.  By colleges     
      b.  By departments    
      c.  By PIs    
      d.  By percentiles    
  2.  Determine if there are differences and if they merit continuation of the project.         
  
  
```{r include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.height=7, fig.width=10)

# adjust graphic parameters

oldPar <- par(cex.main = 3,
    cex.lab = 3,
    cex.axis = 2,
    mar = c(5.1,4.1,4.1,2.1) # default is c(5.1,4.1,4.1,2.1)
    )

```



```{r}

##########
## LOAD ##
##########

# PREP SCRIPT

source(here::here("Prep scripts", "Prep Script from Win Rates by College.R"))

# PER PI

piClusters <- lapply(list.files(here::here("Robjects", "Clustering PIs"), full.names = TRUE), readRDS)

###########
## MERGE ## 
###########

piEmplid <- Reduce(function(x, y) {
  merged <- merge(x, y, by = "emplid", all = FALSE)
  merged <- merged[, !duplicated(sub("\\.x$|\\.y$", "", names(merged)))]  # Remove duplicate columns
  
  # Rename columns to remove ".x"
  names(merged) <- sub("\\.x$", "", names(merged))
  
  merged
}, piClusters)
# piEmplid <- Reduce(function(x, y) merge(x, y, by = "emplid", all = FALSE), piClusters)
# piEmplid <- do.call(merge, piClusters)



prepData <- merge(prepData, piEmplid[,c("emplid","complex_cluster","rate_cluster")], by.x =  "PROPOSAL_PI_EMPLID", by.y = "emplid", all.x = TRUE)

##########
## PREP ##
##########

piEmplid$combined_cluster <-  factor(paste(piEmplid$complex_cluster, piEmplid$rate_cluster, sep = ", "))

prepData$combined_cluster <- factor(paste(prepData$complex_cluster, prepData$rate_cluster, sep = ", "))

###############
## LIBRARIES ##
###############

library(viridis)
library(pheatmap)

# I'll probably want my per-cluster wins as well

# I am going to want to bring in the colors

filterThreeCount <- piEmplid$count.total > 3

piMap <- data.frame(college = row.names(piEmplid), abbrv = row.names(piEmplid), color = NA, pch = 19, cex = 0.7 )


```



```{r}
             
# Manage colors

#############
## COMPLEX ##
#############

complexClusterColors <- c("forestgreen", "deepskyblue", "goldenrod",  
                            "firebrick", "darkslategray", "chartreuse",  
                            "slateblue", "darkkhaki", "coral")

clusterComplexMapping <- piMap
clusterComplexMapping[,"pch"] <- rep(19, nrow(clusterComplexMapping))

clusterComplexMapping[,"color"] <- complexClusterColors [piEmplid$complex_cluster]

# clusterComplexMapping[,"color"] <- complexClusterColors[complexHCPC$data.clust$clust[match(clusterComplexMapping[,"college"], row.names(complexHCPC$data.clust))] ]

##############
## COMBINED ##
##############

combinedClusterColors <- c("forestgreen", "deepskyblue", "goldenrod",  
  "firebrick", "darkslategray", "chartreuse",  
  "slateblue", "darkkhaki", "coral",  
  "mediumorchid", "dodgerblue", "tomato",  
  "orchid", "darkseagreen", "sienna",  
  "royalblue", "indianred", "seagreen",  
  "peru", "cadetblue", "plum",  
  "midnightblue", "lawngreen", "darkorange",  
  "lightsteelblue")

clusterCombinedMapping <- piMap
clusterCombinedMapping[,"pch"] <- rep(19, nrow(clusterCombinedMapping))

clusterCombinedMapping[,"color"] <- combinedClusterColors [piEmplid$combined_cluster]




################
## PERCENTILE ##
################

percentileColors <- viridis(10,direction = -1)[c(1:7,10)]
percentileMapping <- piMap

percentileMapping[,"pch"] <- rep(19, nrow(percentileMapping))

percentileMapping[,"color"] <- percentileColors[as.numeric(cut(piEmplid$win.sum_percentile, breaks = seq(0.2,1,by=0.1)))]

```


```{r}

# Calculate turnover


```


