---
title: "Workforce Journey Exploratory Data Analysis"
author: "Bill Prisbrey"
date: "2025-06-19"
output:
  html_document:
    keep_md: true
---

Notes to self:

Looks like I need to exclude ACTION_REASON == "HCJ" (Hire Concurrent Job) because somehow three records snuck in.

There's no "Too Many Entries" which is encouraging.  

I haven't attempted to see how many of the "Too Many Exits" are due to pre-query hires.  It could be a large number of these.  However figuring that out, opposed to an error, could be tricky.

```{r settings, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE,
                      fig.height=5, fig.width=7)

# fig.height=6.75, fig.width=12) for ppt
# fig.height=7, fig.width=10) typical report


# adjust graphic parameters

oldPar <- par(cex.main = 3,
    cex.lab = 3,
    cex.axis = 2,
    mar = c(5.1,4.1,4.1,2.1) # default is c(5.1,4.1,4.1,2.1)
    )

# par(oldPar) #restore old parameters # restore old parameters after the plot

library(kableExtra)

```


```{r}

##########
## LOAD ##
##########

# PREP SCRIPT

source(here::here("Prep scripts","Adjusting prepData and loading things.R"))


# PER PI

piClusters <- lapply(list.files(here::here("Robjects", "Clustering PIs"), full.names = TRUE), readRDS)

###########
## MERGE ## 
###########

piEmplid <- Reduce(function(x, y) {
  merged <- merge(x, y, by = "emplid", all = FALSE)
  merged <- merged[, !duplicated(sub("\\.x$|\\.y$", "", names(merged)))]  # Remove duplicate columns
  
  # Rename columns to remove ".x"
  names(merged) <- sub("\\.x$", "", names(merged))
  
  merged
}, piClusters)
# piEmplid <- Reduce(function(x, y) merge(x, y, by = "emplid", all = FALSE), piClusters)
# piEmplid <- do.call(merge, piClusters)



prepData <- merge(prepData, piEmplid[,c("emplid","complex_cluster","rate_cluster")], by.x =  "PROPOSAL_PI_EMPLID", by.y = "emplid", all.x = TRUE)

##########
## PREP ##
##########

piEmplid$combined_cluster <-  factor(paste(piEmplid$complex_cluster, piEmplid$rate_cluster, sep = ", "))

prepData$combined_cluster <- factor(paste(prepData$complex_cluster, prepData$rate_cluster, sep = ", "))

###############
## LIBRARIES ##
###############

library(viridis)
library(pheatmap)

#########
## MAP ##
#########

piMap <- data.frame(college = row.names(piEmplid), abbrv = row.names(piEmplid), color = NA, pch = 19, cex = 0.7 )

# I need the college map of colors

fullEmplid <- calculateWinRates(data = cleanData, categoryColumn = "PROPOSAL_PI_EMPLID") |>
  (\(x){x[[1]]})()

fullEmplid$count.total <- apply(fullEmplid[,c("win.count","loss.count")],1,sum)

filterTwoCount <- fullEmplid$count.total >= 2
# filterThreeCount <- piEmplid$count.total > 3

##################################
## EMPL_RCD otherwise duplicate ##
##################################

# RCD_otherwiseDupe <- readRDS(here::here("Robjects", "RCD_otherwiseDupe from Workforced Journey EDA.R"))

```



```{r}

###########
## QUERY ##
###########

# Obtain age data

keyring::keyring_unlock(keyring = "BIPR", password = "Excelsior!")

library(DBI)
con.ds <- DBI::dbConnect(odbc::odbc(), 
                         Driver = "oracle", 
                         Host = "ocm-campus01.it.utah.edu", 
                         SVC = keyring::key_list(keyring = "BIPR")[1, 1],
                         UID = keyring::key_list(keyring = "BIPR")[1, 2],
                         PWD = keyring::key_get(keyring = "BIPR", 
                         service = keyring::key_list(keyring = "BIPR")[1,1]),
                         Port = 2080)

journeyQuery <- "select * from ds_hr.EMPL_AGE_RANGE_ACTION_MV_V WHERE EFFDT < TO_DATE('2025-06-30', 'YYYY-MM-DD') " # a view of that same query

journeyData <- dbGetQuery(con.ds, journeyQuery)


actionReasonQuery <- "
SELECT
  ACTION,
  ACTION_DESCR,
  ACTION_REASON,
  ACTION_REASON_DESCR,
  COUNT(*) AS count
FROM
  ds_hr.EMPL_AGE_RANGE_ACTION_MV_V
WHERE
  EFFDT <= DATE '2025-06-01'
GROUP BY
  ACTION,
  ACTION_DESCR,
  ACTION_REASON,
  ACTION_REASON_DESCR
ORDER BY
  count DESC
"

actionReasonFrame <- dbGetQuery(con.ds, actionReasonQuery)

DBI::dbDisconnect(con.ds)


```

```{r}

##########
## PREP ##
##########

# Create factors with levels in the proper order
journeyData$AGE_BAND <- factor(journeyData$AGE_BAND, levels = 
                                 c(
                                   "Under 20",
                                   "20s",
                                   "30s",
                                   "40s",
                                   "50s",
                                   "60s",
                                   "70s",
                                   "80s",
                                   "90 and Above"
                                 )
                                 )

# Remove a bad date
#> journeyData[journeyData$EFFDT > as.Date("2030-12-31"),]
#          EMPLID EMPL_RCD      EFFDT EFFSEQ ACTION     ACTION_DESCR ACTION_REASON
#1950047 06013186        0 2202-11-21      0    LOA Leave of Absence           EXT
#        ACTION_REASON_DESCR VOLUNTARY_FLAG AGE_BAND
# 1950047           Extension           <NA>      30s

journeyData <- journeyData[journeyData$EFFDT <= as.Date("2030-12-31"),]

```

```{r}

# Identify count of records per EMPLID in order to separate employees with multiple

recordsPerEMPLID <- aggregate(EMPL_RCD ~ EMPLID, data = journeyData, function(x){length(unique(x))})

dupeRCD <- recordsPerEMPLID |> (\(x){x[x$EMPL_RCD > 1, "EMPLID", drop = TRUE]})()  
singleRCD <- unique(journeyData$EMPLID)[!unique(journeyData$EMPLID) %in% dupeRCD]

# proportions(c(length(dupeRCD),length(singleRCD))) # 0.181 0.819

journeySingleFilter <- journeyData$EMPLID %in% singleRCD & journeyData$EMPL_RCD == 0

# NEEDS ATTENTION
journeyZeroFilter <- journeyData$EMPLID %in% singleRCD & journeyData$EMPL_RCD != 0

nonZeroEmplids <- unique(journeyData$EMPLID[journeyZeroFilter])

# I should figure out how to incorporate this logic into a query

journeyData <- journeyData[journeySingleFilter,]

```



```{r}

uniqueActions <- unique(journeyData[,c("ACTION", "ACTION_DESCR")])
actionCount <- table(journeyData[,c("ACTION")]) 
actionFrame <- merge(uniqueActions, data.frame(ACTION = names(actionCount), count = as.vector(actionCount)), by = "ACTION")


uniqueReasons <- unique(journeyData[, c("ACTION_REASON","ACTION_REASON_DESCR")])
reasonCount <- table(journeyData[, "ACTION_REASON"])
reasonFrame <- merge(uniqueReasons, data.frame(ACTION_REASON = names(reasonCount), count = as.vector(reasonCount)), by = "ACTION_REASON")


# Association between action and reason

actionReasonCombo <- table(journeyData[,c("ACTION", "ACTION_DESCR",  "ACTION_REASON", "ACTION_REASON_DESCR")]) |>
  as.data.frame() |>
  (\(x){ x[x$Freq>0,]})() |>
  (\(x){x[order(x$ACTION),]  })()

actionsPerReason <- aggregate(ACTION ~ ACTION_REASON, journeyData, function(x){length(unique(x))} ) |>
  (\(x){x[x$ACTION >1,]})()


reasonsPerAction <- aggregate(ACTION_REASON ~ ACTION + ACTION_DESCR, journeyData, function(x){length(unique(x))} )  


```


```{r}

# Action items per PI

eventsPerEmplid <- aggregate(ACTION ~ EMPLID, data = journeyData, length)
names(eventsPerEmplid) <- c("EMPLID", "eventCount")



```




```{r}

# Define boundary actions

actionReasonFrame$boundary <- NA
actionReasonFrame$boundary_type <- NA

# Primary entry or exit

primeEntryFilter <- actionReasonFrame$ACTION %in% c("HIR", "REH")
primeExitFilter <- actionReasonFrame$ACTION %in% c("TER", "RET", "RWP")

actionReasonFrame$boundary[primeEntryFilter] <- "entry"
actionReasonFrame$boundary[primeExitFilter] <- "exit"
actionReasonFrame$boundary_type[primeEntryFilter|primeExitFilter] <- "primary"

# Break entry or exit

breakEntryFilter <- actionReasonFrame$ACTION %in% c("RWB")
breakExitFilter <- actionReasonFrame$ACTION %in% c("SWB")

actionReasonFrame$boundary[breakEntryFilter] <- "entry"
actionReasonFrame$boundary[breakExitFilter] <- "exit"
actionReasonFrame$boundary_type[breakEntryFilter|breakExitFilter] <- "break"

# Leave entry or exit

leaveEntryFilter <- actionReasonFrame$ACTION %in% c("RFL")
leaveExitFilter <- actionReasonFrame$ACTION %in% c("LOA","LTO", "PLA") & !(actionReasonFrame$ACTION_REASON %in% c("EXT"))

actionReasonFrame$boundary[leaveEntryFilter] <- "entry"
actionReasonFrame$boundary[leaveExitFilter] <- "exit"
actionReasonFrame$boundary_type[leaveEntryFilter|leaveExitFilter] <- "leave"

# now I gotta merge this back to the journey data

journeyData <- merge(journeyData, unique(actionReasonFrame[,c("ACTION","boundary","boundary_type")]), by = "ACTION")


```

```{r}

# boundary consistency

# Commencing a series of aggregations and re-shapes

boundaryFrame <- as.data.frame(table(journeyData[,c("EMPLID","boundary_type","boundary")]))

boundaryList <- lapply(levels(boundaryFrame$boundary_type), function(x){ 
  
  aggregate(Freq~EMPLID + boundary, data = boundaryFrame[boundaryFrame$boundary_type == x,], sum) |>
    (\(y){
      reshape(y,
              idvar = "EMPLID",
              timevar = "boundary",
              v.names = "Freq",
              direction = "wide")
    })() |>
    (\(z){ 
      names(z) <- c("EMPLID","entry","exit");
      return(z)
      })()

  })
names(boundaryList) <- levels(boundaryFrame$boundary_type)

# internal consistency

tooManyExits <- lapply(boundaryList, function(x) { x[x[,"exit"] > x[,"entry"],] })


tooManyEntries <- lapply(boundaryList, function(x) {x[x[,"entry"] > x[,"exit"]+1, ] })



```


```{r}

# Actions per date

actionsPerDate_all <- aggregate(ACTION ~ EFFDT, journeyData, length)

actionsPerWeek_all <- aggregate(ACTION ~ week(EFFDT), actionsPerDate_all, sum)

# nice seasonality graphic

actionsPerDate_EMPLID <- aggregate(ACTION ~ EFFDT + EMPLID, journeyData, length)

# a hist(log()) of this one is kinda nice.

```

```{r eval = FALSE}

# investigating more than one EMPL_RCD
# NOT DONE when I've already filtered out the duplicate or multiples

# and notice my formula -- it's length, not length(unique(x))
# So this can still EMPL_RCD == 0 but it finds multiple actions in a day
# Looks like I made a mistake.

# ...hold on, I might be mistaken about that mistake.  This might do exactly what I wanted.

rcdPerDate_EMPLID <- aggregate(EMPL_RCD ~ EFFDT+EMPLID, data = journeyData, length)

rcdPerDate_EMPLID_Multiple <- rcdPerDate_EMPLID[rcdPerDate_EMPLID[,3]>1,]


# nrow(rcdPerDate_EMPLID_Multiple)/nrow(rcdPerDate_EMPLID) # 0.14

# rcdPerDate_EMPLID_unique <- aggregate(EMPL_RCD ~ EFFDT+EMPLID, data = journeyData, function(x) {length(unique(x))}  )

# skim(rcdPerDate_EMPLID_unique[,"EMPL_RCD"]) # 1

# that's what I thought I was doing. # No, I had it right the first time.

```


```{r eval=FALSE}

# Identifying duplicates other than EMPL_RCD (otherwise everything else is the same)
startTime <- Sys.time()
RCD_otherwiseDupe <- mapply(function(x,y){ 
  filterCondition <- journeyData$EFFDT == x & journeyData$EMPLID == y
  theDiff <- nrow(unique(journeyData[filterCondition,])) -
    nrow(unique(journeyData[filterCondition, -which(colnames(journeyData) %in% "EMPL_RCD")]))
  return(data.frame(EFFDT = x, EMPLID = y, diff = theDiff))
  
  },
  x = rcdPerDate_EMPLID_Multiple[,1],
  y = rcdPerDate_EMPLID_Multiple[,2],
  SIMPLIFY = FALSE) |>
  (\(x){do.call(rbind, x)})()
endTime <- Sys.time()

# this actually will take ten hours
# chatGPT says I can do this with "data.table" and it will take minutes.

# but I'll just run it overnight, why not?
# endTime - startTime # Time difference of 10.66296 hours
saveRDS(RCD_otherwiseDupe, here::here("Robjects", "RCD_otherwiseDupe from Workforced Journey EDA.R"))


```


```{r eval=FALSE}

# investigating more than one EFFSEQ

seqPerDate_EMPLID <- aggregate(EFFSEQ ~ EFFDT+EMPLID, data = journeyData, length)

seqPerDate_EMPLID_Multiple <- seqPerDate_EMPLID[seqPerDate_EMPLID[,3]>1,]


nrow(seqPerDate_EMPLID_Multiple)/nrow(seqPerDate_EMPLID) # 

# Identifying duplicates other than seqPerDate (otherwise everything else is the same)


```


```{r eval=FALSE}

startTime <- Sys.time()
seq_otherwiseDupe <- mapply(function(x,y){ 
  filterCondition <- journeyData$EFFDT == x & journeyData$EMPLID == y
  theDiff <- nrow(unique(journeyData[filterCondition,])) -
    nrow(unique(journeyData[filterCondition, -which(colnames(journeyData) %in% "EFFSEQ")]))
  return(data.frame(EFFDT = x, EMPLID = y, diff = theDiff))
  
  },
  x = rcdPerDate_EMPLID_Multiple[,1],
  y = rcdPerDate_EMPLID_Multiple[,2],
  SIMPLIFY = FALSE) |>
  (\(x){do.call(rbind, x)})()
endTime <- Sys.time()

```


```{r eval = FALSE}

# Sandbox for seeing duplicate rows for everything BUT EMPL_RCD

# let's see how many of these have duplicate rows except for the EMPL_RCD

otherwiseDupe <- apply(rcdPerDate_EMPLID_Multiple[1:10,1:2], 1,
                       function(x,y){print(data.frame(c(x,y)))})



# seems like this should work
mapply(
  function(x, y) {
    print(data.frame(col1 = x, col2 = y))
  },
  x = rcdPerDate_EMPLID_Multiple[1:10, 1],
  y = rcdPerDate_EMPLID_Multiple[1:10, 2],
  SIMPLIFY = FALSE
)

otherwise_RCD_Dupe <- mapply(function(x,y){ 
  filterCondition <- journeyData$EFFDT == x & journeyData$EMPLID == y
  theDiff <- nrow(unique(journeyData[filterCondition,])) -
    nrow(unique(journeyData[filterCondition, -which(colnames(journeyData) %in% "EMPL_RCD")]))
  return(data.frame(EFFDT = x, EMPLID = y, diff = theDiff))
  
  },
  x = rcdPerDate_EMPLID_Multiple[,1],
  y = rcdPerDate_EMPLID_Multiple[,2],
  SIMPLIFY = FALSE) |>
  (\(x){do.call(rbind, x)})()
  
# let's check that

View(journeyData[journeyData$EFFDT == "2022-10-16" & journeyData$EMPLID == "00000133",]) # looks like "EFFSEQ" is different for 2 rows, and third row is completely different

View(journeyData[journeyData$EFFDT == "2010-07-01" & journeyData$EMPLID == "00000395",]) # the onlyh thing that changes is EMPL_RCD

View(journeyData[journeyData$EFFDT == "2010-09-01" & journeyData$EMPLID == "00000395",]) # only change is EMPL_RCD

# Let's do that over all of them and see where we are

```


**PURPOSE:**  The purpose of this report is to describe the "workforce journey" data for employees that only have had one employee record value of zero, meaning no concurrent jobs or possible mis-use of the "EMPL_RCD" field.

It will highlight possible inconsistencies for further clarification with HR. 

**OBJECTIVES:**   

  1.  Describe the "Workforce Journey" data.   
  2.  Define "boundary" actions and check for internal consistency.   

**EXECUTIVE SUMMARY:**

The "workforce journey" data is examined. It is filtered according to the condition of no concurrent jobs, or more specifically, that the employee only ever had a EMPL_RCD value of zero.  Some `r nrow(journeyData) |> (\(x){x+1})() |> format(big.mark=",")` rows use `r length(unique(journeyData$ACTION))` distinct actions to describe the employment experience of `r length(unique(journeyData$EMPLID)) |> format(big.mark = ",")` employees since `r year(min(journeyData$EFFDT))`.  This filtered out the employment history of  `r (length(dupeRCD) + length(nonZeroEmplids)) |> format(big.mark=",")` employees with more than one employee record.

Of particular interest are the so-called "boundary actions" that define entry and exit into the workforce.  These include "primary" actions like hiring and retiring; "break" actions that describe a "short work break"; and "leave" actions for long term disability or a paid leave of absence.   

==

Old version; these were pretty much cleared up.  Fields were clarified and a new query is being written.

When these boundary actions are so simply defined, there are inconsistencies in the respective tallies.  These need to be clarified with HR.

Possible improvements and points for additional clarity include:    

  * Anchor the age bands to a specific date rather than the time of the query.   
  * Clarify the actions and reasons that define boundary entry and exits.   
  * Clarify the fields "EMPL_RCD" and "EFFSEQ".   
  * Clarify edge cases and complex cases (like the employment history of "00028073" and the records of “00029479” on EFFDT == “2011-07-16”).  
  * Review the definition of "voluntary" vs "involuntary" termination and discuss specific cases.   

**SUMMARY:**   

`r nrow(journeyData) |> (\(x){x+1})() |> format(big.mark=",")` rows covering dates between `r min(journeyData$EFFDT)` and `r max(journeyData$EFFDT)` are queried.

One row with a bad date ("2202-11-21") is filtered out.

All employees that had employee record values other than zero are filtered out.  This removes employees who had concurrent positions or erroneous data entry.  These `r(length(dupeRCD) + length(nonZeroEmplids)) |> format(big.mark=",")` employees will be treated separately.     

The remaining data describes `r length(unique(journeyData$EMPLID)) |> format(big.mark = ",")` employees and every employment "action" that defines their experience (such as transfers, pay raises, leave of absence, and eventual termination or retirement.)  `r length(unique(journeyData$ACTION))` distinct actions are further explained with `r length(unique(journeyData$ACTION_REASON))` unique "action reasons." Employees have a median count of `r median(eventsPerEmplid$eventCount)` of these "actions" in their employment history. (At an extreme, one employee (`r eventsPerEmplid$EMPLID[eventsPerEmplid$eventCount == max(eventsPerEmplid$eventCount)]`)  has `r max(eventsPerEmplid$eventCount)` actions.)    

Overwhelmingly, an action of "Data Change" ("DTA") is the most common, followed by "Pay Rate Change" ("PAY") and distantly "Termination" ("TER"), "Hire" ("HIR"), and "Transfer" ("XFR").  The "Termination" ("TER") action has the most "action reasons" (`r length(unique(journeyData$ACTION_REASON[journeyData$ACTION == "TER"]))`).

An age band of the employee is provided of the decade of their age at the time of the query.  Each employee has one un-changing age band. The most frequent band is the decade of the thirties.

Termination is almost exclusively flagged as "voluntary", with only `r proportions(table(journeyData$VOLUNTARY_FLAG))["Involuntary"] |> (\(x){round(100*x,1)})()`% of the flagged rows described as "involuntary."    

A strong seasonal spike in actions is seen around July 1st.  A consistent bi-weekly pattern in actions is also noted.    

Boundary actions are defined as actions that signify entering or leaving the workforce.  This includes "primary" entry actions of hire or rehire, and exit actions of terminate or retire; "break" actions of a "short work break" or "return from work break"; and "leave" actions of paid leave of absence, leave of absence, or long-term disability, and "return from leave."  Leave of absence actions that are explained with the action reason of "extension" are excluded as they do not signify a new boundary event.     


== Let's see 
These are tallied per employee and found to have inconsistent results, where an employee has "exited" the workforce without a prior "entry" action, for example.  This appears to  be due to:    
  (1) simplistic definitions of boundary actions (for example, "rehire" could be ignored if  "Hire Concurrent Job" is the supplied action reason)   
  (2) incomplete understanding of the use of the "EMPL_RCD" field causing duplication that needs to be managed       
  (3) complicated dates where multiple actions are entered that may be contradictory   
  
  

## (1) DESCRIBE THE WORKFORCE JOURNEY DATA

### Data summary

```{r}

skim(journeyData)

```

```{r}

incomingPar <- par(mar = c(13,4,3,1), bg = "ivory", fg = "gray10"  )

theBar <- barplot(actionFrame[order(actionFrame$count, decreasing = TRUE) ,"count"],
        col = "thistle",
        las =2)
axis(side = 1, at = theBar, labels = actionFrame[order(actionFrame$count, decreasing = TRUE) ,"ACTION"], las = 2, line = -3, tick = FALSE)

axis(side = 1, at = theBar, labels = actionFrame[order(actionFrame$count, decreasing = TRUE) ,"ACTION_DESCR"], las = 2, line = 1)

mtext(side = 3, font =2, cex = 1.3, "Count of actions", line = 1.319)

par(incomingPar)

```


```{r}

merge(actionFrame, reasonsPerAction[,c("ACTION","ACTION_REASON")], by = "ACTION") |>
  kbl(caption = "Actions and unique associated reasons",
      col.names = c("", "description", "count", "unique\nreasons"),
      row.names=FALSE,
      escape = FALSE) |>  
  kable_styling(full_width = FALSE) 


```

```{r}

reasonFrame |>
  kbl(caption = "Action reasons",
      col.names = c("", "description","count"),
      row.names=FALSE,
      escape = FALSE) |>  
  kable_styling(full_width = FALSE) |>
  scroll_box(height = "1000px")

```

```{r}

# Reason graphic
# turn this horizontal?

incomingPar <- par(mar = c(13,4,3,1), bg = "ivory", fg = "gray10"  )

plotFrame <- reasonFrame[order(reasonFrame$count, decreasing = TRUE) ,][1:20,]

theBar <- barplot(plotFrame[,"count"],
        col = "lemonchiffon",
        las =2)
axis(side = 1, at = theBar, labels = plotFrame[,"ACTION_REASON"], las = 2, line = -3, tick = FALSE)

axis(side = 1, at = theBar, labels = plotFrame[,"ACTION_REASON_DESCR"], las = 2, line = 1)

mtext(side = 3, font =2, cex = 1.3, "Top twenty reasons by count", line = 1.319)

par(incomingPar)



```

```{r}



actionReasonCombo |>
  kbl(caption = "Actions and reasons",
      col.names = c("", "action", "reason", "description",  "count"),
      row.names=FALSE,
      escape = FALSE) |>  
  kable_styling(full_width = FALSE) |>
  scroll_box(height = "1000px")


actionReasonCombo[actionReasonCombo$ACTION_REASON %in% actionsPerReason$ACTION_REASON,] |>
  (\(x){x[order(x$ACTION_REASON),]})() |>
  kbl(caption = "Reasons applied to multiple actions",
      col.names = c("", "action", "reason","description", "count"),
      row.names=FALSE,
      escape=FALSE
  ) |>  
  kable_styling(full_width = FALSE) |>
  scroll_box(height = "1000px")

actionReasonCombo[!(actionReasonCombo$ACTION_REASON %in% actionsPerReason$ACTION_REASON),] |>
#  (\(x){x[order(x$ACTION_REASON),]})() |>
  kbl(caption = "Reasons applied to a single action",
      col.names = c("", "action", "reason","description", "count"),
      row.names=FALSE,
      escape=FALSE
  ) |>  
  kable_styling(full_width = FALSE) |>
  scroll_box(height = "1000px")



```


```{r}

# Age distribution

journeyData$AGE_BAND |>
  table() |>
  barplot(col = viridis::viridis(n=length(levels(journeyData$AGE_BAND)), dir = -1),
  ylab = "record count",
  main = "Age distribution of action events")



```


```{r}

# Action items per PI

# eventsPerEmplid <- aggregate(ACTION ~ EMPLID, data = journeyData, length)
# names(eventsPerEmplid) <- c("EMPLID", "eventCount")

theHist <- hist(log(eventsPerEmplid$eventCount), plot = FALSE)


plot(theHist, col = viridis::mako(n = length(theHist$breaks)),
     main = "Events per EMPLID\n(log compression)",
     ylab = "count of EMPLID",
     xlab = "events per EMPLID",
     xaxt = "n"
     )

axis(side = 1, at = theHist$breaks, labels = round(exp(theHist$breaks),1)  )

```

```{r}

# Actions per day per EMPLID

#theHist <- hist(actionsPerDate_EMPLID$ACTION, breaks = c(0:5,35), plot=FALSE)

theHist <- hist(actionsPerDate_EMPLID$ACTION, plot = FALSE) #, breaks = c(0:5,35), plot=FALSE)

plot(theHist, col = viridis::mako(n = 3, direction =-1),
     main = "Events per EMPLID per date",
     ylab = "count of EMPLID",
     xlab = "events per EMPLID per date"
     )


```

```{r eval=FALSE}

# termination flag check

table(journeyData[,c("ACTION","VOLUNTARY_FLAG")], useNA = "always") # o.k.

# actions per employee

actionsPerEmplid <- aggregate(count ~ EMPLID + ACTION, 
          data = transform(journeyData, count = 1), 
          FUN = sum)

actionsPerEmplid[actionsPerEmplid$ACTION == "TER","count"] |> hist()

# can I align rehire and termination dates?
# can I see if someone was involuntarily terminated multiple times?

```


```{r}

# Termination flag

termFlag <- aggregate(rep(1, nrow(journeyData)) ~ VOLUNTARY_FLAG, data = journeyData, length )

(termFlag[,2]) |> 
  barplot(col = c("slateblue4", "saddlebrown"),
          names = termFlag[,1],
          main = "Termination action",
          ylab = "count of records"
          )


```


```{r fig.height = 7, fig.width = 5}

# Termination reasons

incomingPar <- par(mar = c(5,17,2,1), bg = "ivory", fg = "gray10"  )

plotFrame <- actionReasonFrame[actionReasonFrame$ACTION == "TER",][1:20,] 

theBar <- barplot(rev(plotFrame[,"COUNT"]),
        col = "mistyrose",
        horiz = TRUE,
        las =2)
axis(side = 2, at = theBar, labels = rev(plotFrame[,"ACTION_REASON"]), las = 2, line = -3, tick = FALSE)

axis(side = 2, at = theBar, labels = rev(plotFrame[,"ACTION_REASON_DESCR"]), las = 2, line = 1)

mtext(side = 3, font =2, cex = 1.3, "Top twenty termination reasons by count", line = .681)

par(incomingPar)



```




```{r}

# Actions over time

actionsInOrder <-  c("POS", "DTA", "PAY", "TER", "HIR", "SWB", "XFR", "RWB", "REH", "JRC", "PLA", "RFL", "LOA", "RET", "RWP", "LTO", "TWP")

actionsPerDate <- lapply(actionsInOrder, function(x){ aggregate(count ~ EFFDT,  data = transform(journeyData[journeyData$ACTION == x,], count =1), length)}   )
names(actionsPerDate) <- actionsInOrder # unique(journeyData$ACTION)

# Determine maximum per action
actionDateMax <- sapply(actionsPerDate, function(x) {max(x[,"count"],na.rm=TRUE )}) |> max() 

# Sort into clusters
actionDateCluster <- 
  actionsPerDate |> 
  sapply(function(x){mean(x[,"count"], na.rm = TRUE)}) |>
  dist() |>
  hclust() |>
  # plot() # looks like 3 is reasonable
  cutree(3)

# Establish decreasing order of mean count
meanCount <- actionsPerDate |> 
     sapply(function(x){mean(x[,"count"], na.rm = TRUE)}) |> sort(decreasing = TRUE)

# Use this to define "actionsInOrder"
# It would be better if I re-ordered it here as this is no longer
# flexible to add a new action

# Establish maximum per cluster

clusterMax <- sapply(1:3,
                     function(y) {
                       
                       sapply(actionsPerDate[actionDateCluster == y ], function(x) {max(x[,"count"], na.rm=TRUE)}  ) 
                     }
                     )

# Create color map per action

# create action colors

actionColorMap <- vector("numeric", length(actionDateCluster))

names(actionColorMap) <- names(actionDateCluster)

actionColorMap[actionDateCluster == 3] <- viridis::viridis(n = length(actionDateCluster[actionDateCluster == 3]))

actionColorMap[actionDateCluster == 2] <- viridis::turbo(n = length(actionDateCluster[actionDateCluster == 2]), begin = 0.3, end = 0.7)

actionColorMap[actionDateCluster == 1] <- viridis::mako(n = length(actionDateCluster[actionDateCluster == 1]))

# Function to plot a cluster at a time

plotActions <- function(cluster) {

incomingPar <-  par(mar = c(2,4,3,6))
on.exit(par(incomingPar))


plot(1,
     type = "n",
     ylim = log(c(1,max(clusterMax[[cluster]]))),
     xlim = range(as.Date(journeyData$EFFDT)),
     ylab = "",
     xlab = "",
     xaxt = "n"
     )

axis.Date(side = 1, at = pretty(as.Date(journeyData$EFFDT)), format = "%Y-%m")

invisible(
  lapply(
    names(actionDateCluster[actionDateCluster == cluster]),
    function(act){
    
    points(x = as.Date(actionsPerDate[[act]][,"EFFDT"]),
           y = log(actionsPerDate[[act]][,"count"]),
           col = actionColorMap[act],
           type = "l",
           lwd = 0.619
    )
      
    }
  )
)

mtext(side = 3, font=2, cex = 1.381, line = 0.619, text = paste0("Actions in Cluster ", cluster)  )

mtext(side = 2, font = 1, cex = 1, line = 2.5, text = "Count of actions (log)")

legend("topright",
       legend = names(actionDateCluster[actionDateCluster == cluster]),
       col = actionColorMap[actionDateCluster == cluster],
       pch = 19,
       pt.cex = 3,
       xpd = TRUE,
       inset = -0.1
       )



}



```

### Count of actions per date

```{r fig.height = 3}

# Plot the count of actions per date, one plot per cluster

plotActions(cluster=1)
plotActions(cluster=2)

```

```{r fig.height = 5}

plotActions(cluster=3)


```


```{r eval=FALSE}

# Dates sandbox

# One line per action over time

jGo <- aggregate(EFFDT ~ ACTION, data = journeyData, function(x){length(unique(x))})

skim(journeyData$EFFDT)

tBlo <- lapply(unique(journeyData$ACTION), function(x){ aggregate(count ~ EFFDT,  data = transform(journeyData[journeyData$ACTION == x,], count =1), length)}   )
names(tBlo) <- unique(journeyData$ACTION)

actionDateMax <- sapply(tBlo, function(x) {max(x[,"count"],na.rm=TRUE )}) |> max() 


# let's lump these

actionDateCluster <- 
  tBlo |> 
  sapply(function(x){mean(x[,"count"], na.rm = TRUE)}) |>
  dist() |>
  hclust() |>
  # plot() # looks like 3 is reasonable
  cutree(3)
  
names(tBlo)

clusterMax <- sapply(1:3,
                     function(y) {
                       
                       sapply(tBlo[actionDateCluster == y ], function(x) {max(x[,"count"], na.rm=TRUE)}  ) 
                     }
                     )

# well, let's plot it and see what we get
# create action colors

actionColorMap <- vector("numeric", length(actionDateCluster))

names(actionColorMap) <- names(actionDateCluster)

actionColorMap[actionDateCluster == 1] <- viridis::viridis(n = length(actionDateCluster[actionDateCluster == 1]))

actionColorMap[actionDateCluster == 2] <- viridis::turbo(n = length(actionDateCluster[actionDateCluster == 2]), begin = 0.3, end = 0.7)

actionColorMap[actionDateCluster == 3] <- viridis::mako(n = length(actionDateCluster[actionDateCluster == 3]))

plotActions <- function(cluster) {

incomingPar <-  par(mar = c(2,4,3,6))
on.exit(par(incomingPar))


plot(1,
     type = "n",
     ylim = c(0,max(clusterMax[[cluster]])),
     xlim = range(as.Date(journeyData$EFFDT)),
     ylab = "",
     xlab = "",
     xaxt = "n"
     )

axis.Date(side = 1, at = pretty(as.Date(journeyData$EFFDT)), format = "%Y-%m")

invisible(
  lapply(
    names(actionDateCluster[actionDateCluster == cluster]),
    function(act){
    
    points(x = as.Date(tBlo[[act]][,"EFFDT"]),
           y = tBlo[[act]][,"count"],
           col = actionColorMap[act],
           type = "l",
           lwd = 0.619
    )
      
    }
  )
)

mtext(side = 3, font=2, cex = 1.381, line = 0.619, text = paste0("Actions in Cluster ", cluster)  )

legend("topright",
       legend = names(actionDateCluster[actionDateCluster == cluster]),
       col = actionColorMap[actionDateCluster == cluster],
       pch = 19,
       pt.cex = 3,
       xpd = TRUE,
       inset = -0.1
       )



}


plotActions(cluster = 1)


plotActions(cluster = 2)
mtext(side = 3, font=2, cex = 1.381, line = 0.619, text ="Actions in Cluster 2"  )
legend("topright",
       legend = names(actionDateCluster[actionDateCluster == 2]),
       col = actionColorMap[actionDateCluster == 2],
       pch = 19,
       pt.cex = 3,
       xpd = TRUE,
       inset = -0.1
       )

plotActions(cluster = 3)
mtext(side = 3, font=2, cex = 1.381, line = 0.619, text ="Actions in Cluster 3"  )
legend("topright",
       legend = names(actionDateCluster[actionDateCluster == 3]),
       col = actionColorMap[actionDateCluster == 3],
       pch = 19,
       pt.cex = 3,
       xpd = TRUE,
       inset = -0.1
       )
```

### Seasonality


```{r}

plot(actionsPerWeek_all,
     xlab = "Week of year",
     ylab = "count of actions",
     type = "l",
     lty = 1,
     lwd = 0.619,
     col = "dodgerblue",
     main = "Weekly actions, all"
)

```


```{r}

# let's do week of year

actionsPerWeek <- lapply(actionsInOrder, function(x){ aggregate(count ~ week(EFFDT),  data = transform(journeyData[journeyData$ACTION == x,], count =1), length)}   )
names(actionsPerWeek) <- actionsInOrder # unique(journeyData$ACTION)


# Establish weekly maximum per cluster

clusterWeekMax <- sapply(1:3,
                     function(y) {
                       
                       sapply(actionsPerWeek[actionDateCluster == y ], function(x) {max(x[,"count"], na.rm=TRUE)}  ) 
                     }
                     )


plotWeeklyActions <- function(cluster) {

incomingPar <-  par(mar = c(4,4,3,6))
on.exit(par(incomingPar))


plot(1,
     type = "n",
     ylim = log(c(1,max(clusterWeekMax[[cluster]]))),
     xlim = range(week(journeyData$EFFDT)),
     ylab = "",
     xlab = "",
     xaxt = "n"
     )

axis(side = 1, at = pretty(week(journeyData$EFFDT)))

invisible(
  lapply(
    names(actionDateCluster[actionDateCluster == cluster]),
    function(act){
    
    points(x = actionsPerWeek[[act]][,"week(EFFDT)"],
           y = log(actionsPerWeek[[act]][,"count"]),
           col = actionColorMap[act],
           type = "l",
           lwd = 0.619
    )
      
    }
  )
)

mtext(side = 3, font=2, cex = 1.381, line = 0.619, text = paste0("Weekly actions in Cluster ", cluster)  )

mtext(side = 2, font = 1, cex = 1, line = 2.5, text = "Count of actions (log)")

mtext(side = 1, font = 1, cex = 1, line = 2.5, text = "Week of year")

legend("topright",
       legend = names(actionDateCluster[actionDateCluster == cluster]),
       col = actionColorMap[actionDateCluster == cluster],
       pch = 19,
       pt.cex = 3,
       xpd = TRUE,
       inset = -0.1
       )



}


```

```{r fig.height = 3}

# Plot the count of actions per date, one plot per cluster

plotWeeklyActions(cluster=1)
plotWeeklyActions(cluster=2)

```

```{r fig.height = 5}

plotWeeklyActions(cluster=3)


```

## (2) DEFINE "BOUNDARY" ACTIONS AND CHECK FOR INTERNAL CONSISTENCY

The actions and action reasons are reviewed, and grouped according to three main boundary types: primary, leave, and break.

These are then tallied per EMPLID to check for inconsistencies.  In particular, the number of exits should never exceed the number of entries for a type, and the number of entries should never be greater than one more than the number of exits.

Many inconsistencies were found, suggesting additional processing and refinement of the "boundary" definitions.

For example:    

  * EMPLID "00028879" on EFFDT == "2010-01-16" shows identical rows except for the EMPL_RCD field.   
  * EMPLID "00029479" on EFFDT == "2011-07-16" show them returning from leave twice, were terminated, and transferred all in one day (and the employment records continue until they retired in 2022.)   
  * EMPLID "00002624" on EFFDT == "2021-03-16" added a "Hire Concurrent Job", which suggests I should not consider this action as a boundary action.


```{r}

explainHeadCount()

```


```{r eval = FALSE}

# Table describing boundary actions

unique(actionReasonFrame[,c("ACTION", "ACTION_DESCR","boundary", "boundary_type")]) |>
  (\(x){x <- x[!is.na(x$boundary),] })() |>
  (\(x){x[order(x$boundary_type),]  })() |>
  
  kbl(caption = "Boundary actions grouped",
      col.names = c("","ACTION_DESCR","boundary",""),
      row.names=FALSE) |>
  column_spec(2:3, extra_css = "padding-left: 1em;")


```


```{r}

sapply(tooManyExits, nrow) |> kbl(caption = "Too many exits", col.names = "count of EMPLID's") |>
  kable_paper("hover", full_width = F)

tooManyExits[[1]][1:10,] |>
  kbl(caption = "Sample of EMPLIDs with more 'break' exits than entries", row.names=FALSE) |>
  kable_paper("hover", full_width = F)

tooManyExits[[2]][1:10,] |>
  kbl(caption = "Sample of EMPLIDs with more 'leave' exits than entries", row.names=FALSE) |>
  kable_paper("hover", full_width = F)

tooManyExits[[3]][1:10,] |>
  kbl(caption = "Sample of EMPLIDs with more 'primary' exits than entries", row.names=FALSE) |>
  kable_paper("hover", full_width = F)

```


```{r}

sapply(tooManyEntries, nrow) |> kbl(caption = "Too many entries", col.names = "count of EMPLID's") |>
  kable_paper("hover", full_width = F)

tooManyEntries[[1]][1:10,] |>
  kbl(caption = "More 'break' entries than exits", row.names=FALSE) |>
  kable_paper("hover", full_width = F)

tooManyEntries[[2]][1:10,] |>
  kbl(caption = "More 'leave' entries than exits", row.names=FALSE) |>
  kable_paper("hover", full_width = F)

tooManyEntries[[3]][1:10,] |>
  kbl(caption = "More 'primary' entries than exits", row.names=FALSE) |>
  kable_paper("hover", full_width = F)


```


```{r eval=FALSE}

# boundary consistency sandbox

# checking to see if boundary pairs (exit/entry per type) match

table(journeyData$boundary)
# entry   exit 
# 305797 303913 

# I'll need to do this per EMPLID

shobsho <- aggregate(boundary~EMPLID + boundary_type, data = journeyData, table) |> as.matrix.data.frame()

# that's not it

shobble <- as.data.frame(table(journeyData[, c("EMPLID","boundary_type","boundary")]))

# that's a lot more like it

# now I wonder if I can filter out the values of 0,
# and I wonder how to make that brief entry/exit map

# And I wonder if I want to re-shape this

shobble[shobble$EMPLID == monthOneTerm[1],]

# almost there to finding imbalance

# another aggregation, I think, to find them out



ohbsho <- lapply(levels(shobble$boundary_type), function(x) {aggregate(Freq ~ EMPLID, data = shobble[shobble$boundary_type == x,], length(unique(x)) )})


shobble[shobble$boundary_type == "primary",] |> head()


shaibo <- aggregate(Freq ~ EMPLID + boundary, data = shobble[shobble$boundary_type == "primary",], sum) |>
  (\(x){x[order(x$EMPLID),]})()

# I am going to re-shape this.
# I need to see where exit > entry
# or where entry is more than exit + 1

shaibobble <- reshape(shaibo,
                      idvar = "EMPLID",
                      timevar = "boundary",
                      v.names = "Freq",
                      direction = "wide")
names(shaibobble) <- c("EMPLID","entry","exit")

# ok, I like this
# I c/should turn this into a list

shazam <- lapply(levels(shobble$boundary_type), function(x){ 
  
  aggregate(Freq~EMPLID + boundary, data = shobble[shobble$boundary_type == x,], sum) |>
    (\(y){
      reshape(y,
              idvar = "EMPLID",
              timevar = "boundary",
              v.names = "Freq",
              direction = "wide")
    })() |>
    (\(z){ 
      names(z) <- c("EMPLID","entry","exit");
      return(z)
      })()

  })
names(shazam) <- levels(shobble$boundary_type)

# ok, now we are getting somewhere!


# too many cooks

incongruous <- lapply(shazam, function(x) { x[x[,"exit"] > x[,"entry"],] })


tooManyCooks <- lapply(shazam, function(x) {x[x[,"entry"] > x[,"exit"]+1, ] })

# ok now this is interesting
# check out "00028879", EFFDT == "2010-01-16"
# They are IDENTICAL except for the EMPL_RCD

# I should just shoot a quick e-mail and see if I can figure this out

# ...or, pick up a few examples of these

# View(journeyData[journeyData$EMPLID == "00029479",])

# check out "00029479", EFFDT == "2011-07-16"
# They returned from leave twice, were terminated, and transferred all in one day
# ...and continued employment until they retired on "2022-06-16"

# View(journeyData[journeyData$EMPLID == "00002624",])
# check out "00002624" on 2021-03-16
# Looks like this is a "Hire Concurrent Job" which needs some explaining

# let's just write this up and set up a meeting.


```



```{r eval=FALSE}

# print and work with actionReasonFrame in Excel

write.csv(actionReasonFrame, here::here("Robjects", "actionReasonFrame.csv"))


```


### Query

The query below is converted to a view, ds_hr.EMPL_AGE_RANGE_ACTION_MV_V.

"SELECT A.EMPLID, A.EMPL_RCD, A.EFFDT, A.EFFSEQ
    , A.ACTION, B.ACTION_DESCR
    , A.ACTION_REASON, C.DESCR ACTION_REASON_DESCR
    , Case when A.ACTION ! = 'TER' then ''
        when A.ACTION  = ('TER') 
        AND A.ACTION_REASON not in ('BNK', 'EVW', 'I9', 'INV', 'NER', 'RFN', 'RIF', 'RLS') 
        then 'Voluntary' else 'Involuntary' end VOLUNTARY_FLAG
    , Case    when (SYSDATE-D.BIRTHDATE)/365.25 < 20
        then 'Under 20' 
    when (SYSDATE-D.BIRTHDATE)/365.25 >= 20  and (SYSDATE-D.BIRTHDATE)/365.25 < 30
        then '20s'
    when (SYSDATE-D.BIRTHDATE)/365.25 >= 30  and (SYSDATE-D.BIRTHDATE)/365.25  < 40
        then '30s'
    when (SYSDATE-D.BIRTHDATE)/365.25 >= 40  and (SYSDATE-D.BIRTHDATE)/365.25  < 50
        then '40s'
    when (SYSDATE-D.BIRTHDATE)/365.25 >= 50  and (SYSDATE-D.BIRTHDATE)/365.25 < 60
        then '50s'
    when (SYSDATE-D.BIRTHDATE)/365.25 >= 60  and (SYSDATE-D.BIRTHDATE)/365.25  < 70
        then '60s'
    when (SYSDATE-D.BIRTHDATE)/365.25 >= 70  and (SYSDATE-D.BIRTHDATE)/365.25 < 80
        then '70s'
    when (SYSDATE-D.BIRTHDATE)/365.25 >= 80  and (SYSDATE-D.BIRTHDATE)/365.25 < 90
        then '80s'
    else '90 and Above'
        end Age_Band
        
FROM PS_UU_UNSEC_JOB_VW A
  JOIN PS_ACTION_TBL B
    ON (B.ACTION = A.ACTION
    AND B.EFFDT =
        (SELECT MAX(B_ED.EFFDT) FROM PS_ACTION_TBL B_ED
        WHERE B.ACTION = B_ED.ACTION
          AND B_ED.EFFDT <= SYSDATE))
  JOIN PS_ACTN_REASON_TBL C
    ON (C.ACTION = A.ACTION
     AND C.ACTION_REASON = A.ACTION_REASON
     AND C.EFFDT =
        (SELECT MAX(C_ED.EFFDT) FROM PS_ACTN_REASON_TBL C_ED
        WHERE C.ACTION = C_ED.ACTION
          AND C.ACTION_REASON = C_ED.ACTION_REASON
          AND C_ED.EFFDT <= SYSDATE))
    JOIN ps_personal_dt_fst D
        ON (D.EMPLID = A.EMPLID)
        
WHERE  A.EFFDT > TO_DATE('2010-01-01','YYYY-MM-DD')
    ORDER BY A.EMPLID, A.EMPL_RCD, A.EFFDT"

```{r eval = FALSE}

# Let's check on people terminated in the first month of the data set

dateCriteria <- journeyData$EFFDT <= as.Date("2010-01-31")
termCriteria <- journeyData$ACTION == "TER"
hireCriteria <- journeyData$ACTION == "HIR"
rehireCriteria <- journeyData$ACTION == "REH"

monthOneTerm <- unique(journeyData$EMPLID[dateCriteria & termCriteria])

dim(journeyData[journeyData$EMPLID == monthOneTerm[1],]) # looks like I found a re-hire.  His last record happened in 2018

journeyData[journeyData$EMPLID == monthOneTerm[1],] |>
  (\(x){x[order(x$EFFDT),] })()


journeyData[journeyData$EMPLID == monthOneTerm[2],] |>
  (\(x){x[order(x$EFFDT),] })() # see now this is what I'm looking for
# A termination but no matching hire date

# I need to understand the EFFSEQ, and there's actually value for EMPL_RCD

# 

```

