---
output:
  html_document:
    keep_md: true
---

```{r include = FALSE}

# Notes

# I could expand the summary, as it sure is concise the way it is.

# I need to finish the names. DONE

# Next to the first "population" bar chart I want a "percent of money won" bar chart that is identical. DONE

# I need an incoming "filter" graphic to explain the filtering. DONE

# The basic bar charts needs some help. DONE

# Maybe I need a separate "filter" child that applies to all of the following appendices. DONE

# What I need is a barchart of the population per cluster, # JUST THE TABLE
# possibly with an embedded table that gives the counts.  # OK

# possibly a table of the clusters (a simple aggregation) would be good, too. # DONE

# I would need to use "flextable" as "kableextra" is for html # NEEDS WORK BUT O.K.


```

```{r include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.height=7, fig.width=10)

# adjust graphic parameters

oldPar <- par(cex.main = 3,
    cex.lab = 3,
    cex.axis = 2,
    mar = c(5.1,4.1,4.1,2.1) # default is c(5.1,4.1,4.1,2.1)
    )

```




```{r}

##########
## LOAD ##
##########

source(here::here("Prep scripts", "Adjusting prepData and loading things.R"))

###############
## LIBRARIES ##
###############

library(viridis)
library(pheatmap)

######################
## CALCULATE PER PI ##
######################

piEmplid <- calculateWinRates(data = prepData, categoryColumn = "PROPOSAL_PI_EMPLID", functionList = list(mean = mean, median = median)) |> (\(x){ x[[1]] })()
# really starting to hate the list nature and why do I want "call" ?

piEmplid$count.total <- apply(piEmplid[,c("win.count","loss.count")], 1, sum, na.rm = TRUE )
piEmplid$sum.total <- apply(piEmplid[,c("win.sum","loss.sum")], 1, sum, na.rm = TRUE )

##########
## PREP ##
##########

piEmplid$emplid <- row.names(piEmplid)
filterThreeCount <- piEmplid$count.total > 3
piMap <- data.frame(college = row.names(piEmplid), abbrv = row.names(piEmplid), color = NA, pch = 19, cex = 0.3 )

#######################
## TEST FOR NORMALCY ##
#######################

normalTest <- apply(piEmplid[filterThreeCount,!colnames(piEmplid) %in% c("emplid", "count_cut","rate_cluster","totals_cluster", "complex_cluster")], 2, function(x) shapiro.test(x)$p.value) |>
  (\(x){any(x > 0.05) })() # FALSE

# unsurprisingly --- let's log transform before z-transform


#####################
## CREATE CLUSTERS ##
#####################

complexPCA <- piEmplid[filterThreeCount,!colnames(piEmplid) %in% c("emplid", "count_cut","rate_cluster","totals_cluster", "complex_cluster")]|>
  (\(x){data.frame(
    log(x[,-which(colnames(x) %in% c("count.rate","sum.rate"))]), 
    x[,c("count.rate","sum.rate")]
    ) })() |> # log transformation of none-rate columns
  (\(x){
    x[x==-Inf] <- 0;
    return(x)
  })() |>
  scale() |>
  PCA(graph = FALSE)
  
complexHCPC <- complexPCA |>
  HCPC(nb.clust = 5, graph = FALSE, metric = "euclidean", method = "ward") 

```



```{r}
             
# Manage colors

if(length(complexHCPC[["desc.var"]][[2]]) <= 9) {
  
  complexClusterColors <- c("forestgreen", "deepskyblue", "goldenrod",  
                            "firebrick", "darkslategray", "chartreuse",  
                            "slateblue", "darkkhaki", "coral")[1:(length(complexHCPC[["desc.var"]][[2]]))]
  
} else {
  complexClusterColors <- c("forestgreen", "deepskyblue", "goldenrod",  
                            "firebrick", "darkslategray", "chartreuse",  
                            "slateblue", "darkkhaki", "coral", 
                            rep(c("darkorange","purple"), length.out = length(complexHCPC[["desc.var"]][[2]])-9 ))
}

clusterComplexMapping <- piMap
clusterComplexMapping[,"pch"] <- rep(19, nrow(clusterComplexMapping))
clusterComplexMapping[,"color"] <- complexClusterColors[complexHCPC$data.clust$clust[match(clusterComplexMapping[,"college"], row.names(complexHCPC$data.clust))] ]

```


```{r}

# Merge clusters back into piEmplid
piEmplid <- merge(piEmplid, complexHCPC$data.clust[,"clust", drop = FALSE], by.x = "emplid", by.y= "row.names")

row.names(piEmplid) <- piEmplid$emplid
names(piEmplid)[names(piEmplid) == "clust"] <- "complex_cluster" 


# Merge into prepData
prepData <- merge(prepData, piEmplid[,c("emplid","complex_cluster")], by.x = "PROPOSAL_PI_EMPLID", by.y = "emplid", all.x = TRUE)

```


```{r}

# Calculate win rates

complexClusterWins <- calculateWinRates(prepData, categoryColumn = "complex_cluster", functionList = list(mean = mean, median = median)) |>
  (\(x){
    return(x$summary[match(levels(prepData$complex_cluster), row.names(x$summary)),])
  })()

complexClusterWins$color <- complexClusterColors

complexClusterWins$population <- aggregate(emplid ~ complex_cluster,  data = piEmplid, length )[,2]

complexClusterWins$pop_perc <- round(100*(complexClusterWins$population / sum(complexClusterWins$population)),1)


complexClusterWins$sum_perc <- round(100*(complexClusterWins$win.sum/sum(complexClusterWins$win.sum)),1)

complexClusterWins$count_total <- apply(complexClusterWins[,c("win.count","loss.count")], 1, sum)

complexClusterWins$count_perc <- round(100*(apply(complexClusterWins[,c("win.count","loss.count")], 1, sum)/sum(apply(complexClusterWins[,c("win.count","loss.count")], 1, sum))),1)


medianPIbyCluster <- aggregate(win.sum ~ complex_cluster, data = piEmplid, median) # for plot

```



```{r}

# Population heatmaps and table

piNames <- row.names(piEmplid)[piEmplid$count.total > 3]
piFilter <- prepData$PROPOSAL_PI_EMPLID %in% piNames

collegePopulation <- unique(prepData[piFilter, c("PROPOSAL_PI_EMPLID", "college","complex_cluster" )])

# > length(unique(collegePopulation[,1]))
# [1] 1672
# > length((collegePopulation[,1]))
# [1] 1922

# I think I'll just allow the duplication

```

```{r}

# Comparing rates of submissions by 4 and 5

longevity <- aggregate(interval.PI ~ complex_cluster, data = prepData, median)

PI_yr_cluster <- aggregate(PROPOSAL_ID ~ PROPOSAL_PI_EMPLID + floor(interval.PI) + complex_cluster, data = prepData, length )

pace <- aggregate(PROPOSAL_ID ~ complex_cluster, data = PI_yr_cluster, median) # median number of proposals submitted per year per cluster

```

# COMPLEX CLUSTERS BASED ON MULTIPLE DIMENSIONS

### SECTIONS

  (1) Summary   
  (2) Method    
  (3) Clusters verbally described 
  (4) Clusters visually described
  (5) Table of aggregated values by cluster   
  (6) Hierachical clustering on principal components    
  (7) Clusters displayed on scatter plots   
  (8) Variable relevance per cluster   
  (9) Clusters over time    
  (10) Cluster populations by college and major institutions    

### 1. SUMMARY

This shows principal investigators sorted into five clusters using multiple criteria.

### 2. METHOD

First, principal investigators (PI's) with three or fewer proposals are filtered out.  `r format(table(filterThreeCount)[1], big.mark = ",")` principal investigators (`r  round(100*proportions(table(filterThreeCount))[1], 0) `%) are removed and `r format(table(filterThreeCount)[2], big.mark = ",")` principal investigators (`r  round(100*proportions(table(filterThreeCount))[2], 0) `%) are kept.    

Second, several variables are calculated per principal investigator:    

  * Sum of funds requested (won and lost)       
  * Count of proposals (won and lost)   
  * Mean and median of funds requested (won and lost)
  * Rate of proposals won (count rate)        
  * Rate of funds requested won (sum rate)    
  * Total count and sum   

Third, variables except rates are scaled with a natural log transformation first, then all variables are centered and scaled.

Fourth, principle components are extracted.

Fifth, hierarchical clustering using Euclidean distance and Ward's method is used to produce five clusters.

The clusters are named, described verbally and visually, and the cluster populations by colleges and major institutions are shown.  

### 3. CLUSTERS VERBALLY DESCRIBED

**CLUSTER 1: PERFECT**    
  Principal investigators in Cluster 1 are perfect in their attempts with 100% win rates of relatively smaller proposals.    
  
  A high proportion of PI's from the School of Business are found here.  
  
  It has a population of `r complexClusterWins[1,"population"]` PIs (`r complexClusterWins[1,"pop_perc"]`%) and accounts for only `r complexClusterWins$sum_perc[1] `% of total requested funds won.

**CLUSTER 2:  PRECISE**   
  Principal investigators in Cluster 2 don't have the perfect record of Cluster 1, but win a high proportion of their proposals on both a count and sum basis. They bring in as much as Cluster 4 in total funds requested won through racking up lots of relatively smaller wins.   
  
  It has a population of `r complexClusterWins[2,"population"]` PIs (`r complexClusterWins[2,"pop_perc"]`%) and accounts for `r complexClusterWins$sum_perc[2] `% of total requested funds won.     
  
**CLUSTER 3: PIPE DREAMS**    
Cluster 3 contains principal investigators with zero and near-zero win rates.  They ambitiously attempt proposals as large as the principal investigators in Clusters 4 and 5 -- just without success.    

  It has high proportions of PI's from the colleges of Law, Health, and Social and Behavioral Sciences.
  
  It has a population of `r complexClusterWins[3,"population"]` PIs (`r complexClusterWins[3,"pop_perc"]`%) and accounts for  `r complexClusterWins$sum_perc[3] `% of total requested funds won.
  
**CLUSTER 4:  PLUCKY**    
Cluster 4 contains principal investigators who mainly differ from Cluster 5 in the count of proposals submitted.  They submit equivalently large proposals, just far fewer -- less than a third of the proposals submitted by Cluster 5.  Some of this is due to longevity (PI's in Cluster 4 have been submitting bids for a median of `r round( longevity[4,2],2)` years compared to `r round( longevity[5,2],2)` years in Cluster 5), but some is also due to the rate of submission:  PI's in Cluster 5 submit a median of 30% more bids per year.   

Cluster 4 PI's also slightly lag Cluster 5 PI's in win rates (count and sum) as well.    

  It has a population of `r complexClusterWins[4,"population"]` PIs (`r complexClusterWins[4,"pop_perc"]`%) and accounts for  `r complexClusterWins$sum_perc[4] `% of total requested funds won.
  
**CLUSTER 5:  PROLIFIC**    
Cluster 5 contains principal investigators with a prolific record of submission.  They submit a large number of proposals -- triple Cluster 4, the next highest -- with impressive win rates on large proposals.  These are the workhorses of University of Utah research.    

  Principal investigators from the institutions (EGI, CTSI, CVRTI, ICSE, and SCI) tend to show up here.

It has a population of `r complexClusterWins[5,"population"]` PIs (`r complexClusterWins[5,"pop_perc"]`%) and accounts for a large majority (`r complexClusterWins$sum_perc[5] `%) of total requested funds won.    

### 4. CLUSTERS VISUALLY DESCRIBED

```{r fig.height = 4.5, fig.width = 9}

#########################################
## THREE PROPORTION BARS AND A BOXPLOT ##
#########################################


####################
## POPULATION BAR ##
####################

par(bg = "ivory",
    fg = "gray10",
    oma = c(0,0,3,0),
    mar = c(3,3,3,1),
    mfrow= c(1,2))

# dummy graph to create the background

theBat <- barplot(
  complexClusterWins[,"population"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  complexClusterWins[,"population"],
  col = complexClusterWins[,"color"],
  las = 2
)

mtext(text = row.names(complexClusterWins),
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font =2,
      col = complexClusterWins[,"color"])

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Population"
      )

text(x = theBar,
     y = complexClusterWins[,"population"],
     pos = c(3,3,3,1,1),
     font = 2,
     col = c(rep("gray10",4),"gray90"),
     labels = paste0(complexClusterWins$pop_perc,"%")
     )


###############
## FUNDS BAR ##
###############

# adjust margins

par(mar = c(3,5,3,1))

# dummy graph to create the background

theBat <- barplot(
  complexClusterWins[,"win.sum"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  complexClusterWins[,"win.sum"],
  col = complexClusterWins[,"color"],
  las = 2
)

mtext(text = row.names(complexClusterWins),
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font = 2,
      col = complexClusterWins[,"color"]
        )

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Funds requested won"
      )

text(x = theBar,
     y = complexClusterWins[,"win.sum"],
     pos = c(3,3,3,3,1),
     cex = c(1,1,1,1,0.9),
     font = 2,
     col = c(rep("gray10",4),"gray90"),
     labels = paste0(complexClusterWins$sum_perc,"%")
     )

# float the median over THAT !

mtext(text = "Proportions and proposal amounts per cluster",
      side = 3,
      line = 0.309,
      font = 2,
      cex = 2,
      outer = TRUE)

###############
## COUNT BAR ##
###############

# adjust margins

par(mar = c(3,4,3,1),
    oma = c(0,0,0,0))

# par(bg = "ivory",
#    fg = "gray10",
#    oma = c(0,0,0,0),
#    mar = c(3,3,3,1)
#    )

# dummy graph to create the background

theBat <- barplot(
  complexClusterWins[,"count_total"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  complexClusterWins[,"count_total"],
  col = complexClusterWins[,"color"],
  las = 2
)

mtext(text = row.names(complexClusterWins),
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font = 2,
      col = complexClusterWins[,"color"]
        )

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Count of proposals"
      )

text(x = theBar,
     y = complexClusterWins[,"count_total"],
     pos = c(3,3,3,3,1),
     cex = c(1,1,1,1,0.9),
     font = 2,
     col = c(rep("gray10",4),"gray90"),
     labels = paste0(complexClusterWins$count_perc,"%")
     )


#############
## BOXPLOT ##
#############

par(mar = c(3,4,3,1))

# par(bg = "ivory",
#    fg = "gray10",
#    oma = c(0,0,0,0),
#    mar = c(3,4,4,1))
    
#    ,
#    mfrow= c(1,1))

boxplot(log(PROPOSAL_TOTAL_SPONSOR_BUDGET) ~ complex_cluster,
        data = prepData,
        names = rep("",5),
        yaxt = "n",
        ylab = ""
        )

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBox <- boxplot(log(PROPOSAL_TOTAL_SPONSOR_BUDGET) ~ complex_cluster,
        data = prepData,
        col = complexClusterWins$color,
        names = rep("",5),
        ylab = "",
        las = 1
        )

mtext(text = c("Proposal funds requested", "Funds requested (log)" ),
      side = c(3,2),
      font = c(1,1),
      cex = c(1.75,1.5),
      line = c(0.309, 2.427)
      
      )

mtext(text = row.names(complexClusterWins),
      line = 1,
      at = 1:5,
      side = 1,
      cex = 1.5,
      font = 2,
      col = complexClusterWins[,"color"]
        )

```

```{r}

# Simple bars describing the clusters

par(oma = c(0,0,3,0), mfrow = c(2,1))

plotWinRates(complexClusterWins,
             agg = "count",
             line = "count.rate",
             bar_params = list(mar = c(2,7,4,5)),
             bar_args = list(las = 1),
             bar_mtext_args = list(
               text = c("Proposals per cluster",
                        "",
                        "Count of proposals"
                        ),
               line = c(1.618,0,5),
               col = c("gray10", "gray10", "gray10"),
               font = c(1,3,1),
               cex = c(1.75,1.25,1.15),
               side = c(3,3,2)
                        ),
             bar_legend_args = list(x = "topleft")
             )

mtext(text = "Clusters described by count and sum",
      side = 3,
      line = 0.309,
      font = 2,
      cex = 2,
      outer = TRUE)

plotWinRates(complexClusterWins,
             agg = "sum",
             line = "sum.rate",
             bar_params = list(mar = c(2,7,4,5)),
             bar_args = list(las = 1),
             bar_mtext_args = list(
               text = c("Funds requested per cluster",
                        "",
                        "Sum of funds requested"
                        ),
               line = c(1.618,0,5),
               col = c("gray10", "gray10", "gray10"),
               font = c(1,3,1),
               cex = c(1.75,1.25,1.15),
               side = c(3,3,2)
                        ),
             bar_legend_args = list(x = "topleft")
)

```


```{r eval=FALSE}

# I did this using "college" column and "pheatmap" at the end of the report

# I'd like to make  heatmap of clusters per college and 5 institutions

table(prepData[,c("college16_bigInst", "complex_cluster")]) |>
  log() |>
  (\(x){ 
   x[x==-Inf] <- 0;
    return(x) 
    })() |>
  heatmap(
  scale = "row"
)



# let's show the row-wise percentages

table(prepData[,c("college16_bigInst", "complex_cluster")]) |>
  as.data.frame.matrix() |>
  apply(1, proportions) |>
  t() |>
  heatmap()



table(prepData[,c("college16_bigInst", "complex_cluster")]) |>
  as.data.frame.matrix() |>
  apply(1, proportions) |>
  t() |>
  heatmap(
    scale = "column"
  )

# it's all kind of interesting
# But this is just the wrong basis.
# I need to do this per PI, not per proposal

# I'll have to think about that one a little bit

```



```{r, eval = FALSE}

# If you look at clYr[[2]] and that last final aggregation,
# you'll see that Cluster 5 produces at a much higher rate (double the proposals per year)

# But I think I'll leave that for an investigation using time clusters


curious <- aggregate(interval.PI ~ complex_cluster, data = prepData, median)

# so cluster 5 has been at it longer
# let's see the count per elapsed year
# I need some work on the logic

stillCurious <- aggregate(interval.PI ~ floor(interval.PI) + complex_cluster, data = prepData, length )

# seems like I can do an aggregation that manages "elapsed"
# should I attempt a new one for the 

# or just see what my "elapsed" ones do?

piYr <- elapsed(data = prepData)
piYrMon <- elapsedCash(data = prepData)

clYr <- elapsed(target = "complex_cluster", data = prepData )
clYrMon <- elapsedCash(target = "complex_cluster", data = prepData )

# or-r-r-r-r
# before figuring that out (and implying I've analyzed it)
# I just want to figure out the pace.

catsCurious <- aggregate(PROPOSAL_ID ~ PROPOSAL_PI_EMPLID + floor(interval.PI) + complex_cluster, data = prepData, length )

aggregate(PROPOSAL_ID ~ complex_cluster ,data = catsCurious, mean) # MUCH higher pace

# timeDots(data = clYr[[2]][clYr[[2]][,"complex_cluster"] == 1, ] ) error

filter <- clYrMon$each[,"complex_cluster"] == 1
timeDots(data = clYrMon$each[filter,])
# not happy

filter <- piYrMon$each[,"PROPOSAL_PI_EMPLID"] %in% piEmplid$emplid[piEmplid$complex_cluster == 1]
timeDots(data = piYrMon$each[filter,])
# Happy !

filter <- piYr$each[,"PROPOSAL_PI_EMPLID"] %in% piEmplid$emplid[piEmplid$complex_cluster == 1]
timeline(data = piYr$each[filter,]) # happy (!)

# Honestly not a bad graphical representation

# Is it sorted somehow?  Seems like it




```

### 5. TABLE OF AGGREGATED VALUES BY CLUSTER

```{r}

displayColumns <- c(
    "population",
    "pop_perc",
    "count_total", 
    "count_perc",
    "win.sum",
    "sum_perc",
    "win.count",   
    "win.mean",    
    "win.median", 
  "loss.sum",    
  "loss.count",  
  "loss.mean",   
  "loss.median",
  "sum.rate",    
  "count.rate",  
  "color"
)


columnRounding <- c(
    "population" = 0,
    "pop_perc" =1,
    "count_total" =-1, 
    "count_perc" = 0,
    "win.sum" = -6,
    "sum_perc" = 0,
    "win.count" = -1,   
    "win.mean" = -3,    
    "win.median" = -3, 
  "loss.sum" = -6,    
  "loss.count" = -1,  
  "loss.mean" = -2,   
  "loss.median" = -2,
  "sum.rate" = 2,    
  "count.rate" = 2  
)

# Convert result_df to a flextable
ft <- data.frame(cluster = rownames(complexClusterWins), complexClusterWins[,displayColumns] ) |>
  (\(x){ 
      x[, names(columnRounding)] <- Map(round, x[, names(columnRounding)], columnRounding);
    x  # Return the modified data frame
    })() |>
  flextable::flextable() |>
  flextable::theme_vanilla() %>%  # Apply a clean theme
  flextable::set_caption("Aggregated Metrics by Complex Cluster") %>%  # Add a title
  flextable::colformat_num(j = 2:ncol(complexClusterWins), digits = 2) %>%  # Format numeric columns
  flextable::autofit()  # Adjust column widths for better display

# Print the table
ft


```



```{r fig.width=7, eval=FALSE}

# moved this below the dendrogram to appear next to the factor map

fviz_pca_biplot(complexPCA,
                geom = "none",
                title = "Principal components of multiple variables")

```


### 6. HIERARCHICAL CLUSTERING ON PRINCIPAL COMPONENTS


```{r fig.height = 4.5, fig.width = 9}

library(gridExtra)

p1 <- fviz_pca_biplot(complexPCA,
                       geom = "none",
                       title = "Principal components of multiple variables")

p2 <- fviz_cluster(complexHCPC, 
                   geom = "point",
                   main = "PI clusters plotted on component dimensions",
                   show.clust.cent = TRUE)

# Arrange plots side by side
grid.arrange(p1, p2, ncol = 2)  # Change `ncol = 1` for vertical arrangement

```

```{r fig.height = 4.5, fig.width=4.5}

# Dendrogram

if(length(complexHCPC[["desc.var"]][[2]]) <= 12) {
  plot(complexHCPC, choice = "tree", cex = 0.05)
}

```


### 7. CLUSTERS DISPLAYED ON SCATTER PLOTS

```{r}

# Plot by win sum and rate clusters

par(oma = c(0,3,4,0), mfrow =c(1,1))

yBYx(data = piEmplid,  
     x_axis =  "win.sum", 
     y_axis = "sum.rate", 
     detailMapping = clusterComplexMapping, 
     log = c("x"), 
     scatter_args = list(xlim = c(0,18.5)),
     scatter_text_args = list(
       labels = "",
       cex = 1.618),
     scatter_mtext_args = list(text = c("", "Sum of funds requested won (log)", "Fraction of funds requested won"), 
                               cex = c(1.75,1.5,1.5),
                               font = c(0,0,0)   ))

mtext(text = "Sum rate by total funds won" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

text(x = log(medianPIbyCluster$win.sum) |> (\(x){ x[is.infinite(x)] <- 0; return(x)})() ,
     y = complexClusterWins$sum.rate,
     labels = 1:(length(complexHCPC[["desc.var"]][[2]])),
     cex = 2,
     font = 2,
     col = complexClusterWins$color,
     xpd = TRUE
) 

```


```{r fig.width=10}

# Tiled sum scatter plots of complex clusters

par(oma = c(5,7,4,1), mfrow = c(3,2))

invisible(
  lapply(levels(piEmplid$complex_cluster), function(x){ 
    
    yBYx(data = piEmplid[piEmplid$complex_cluster == x,],  
         x_axis = "win.sum", 
         y_axis = "sum.rate", 
         detailMapping = clusterComplexMapping, 
         log = c("x"), 
         scatter_params = list(mar = c(0,0,0,0)),
         scatter_args = list(cex = 1.618, xlim = c(0,20), ylim = c(0,1)),
         scatter_text_args = list(
           labels = "",
           cex = 1.618),
         scatter_mtext_args = list(text = c("", "", "") )
         #                              cex = c(1.75,1.5,1.5),
         #                              font = c(0,0,0)   )
    )
    
    textFilter <- row.names(complexClusterWins) == x
    #text(x = totalsClusterWins$count.rate[textFilter],
    #    y = totalsClusterWins$sum.rate[textFilter],
    #    labels = row.names(totalsClusterWins)[textFilter],
    #    cex = 2,
    #    font = 2,
    #    col = totalsClusterWins$color[textFilter]
    #    )   
    
    text(x = 0.05,
         y = 0.9,
         labels = row.names(complexClusterWins)[textFilter],
         cex = 3,
         font = 2,
         col = complexClusterWins$color[textFilter],
         xpd = TRUE
    )  
    
    
  })    
)    



# mtext(text = "Clusters based on multiple dimensions" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

mtext(text = "Sum of funds requested won (log)" , side = 1, outer = TRUE, font = 1, cex = 1.5, line = 3) 

mtext(text = "Fraction of funds requested won" , side = 2, outer = TRUE, font = 1, cex = 1.5, line = 3) 

```


```{r}

# Rate scatter plot

par(oma = c(0,3,4,0), mfrow =c(1,1))

yBYx(data = piEmplid,  
     x_axis = "count.rate", 
     y_axis = "sum.rate", 
     detailMapping = clusterComplexMapping, 
     log = NA, 
     scatter_text_args = list(
       labels = "",
       cex = 1.618),
     scatter_mtext_args = list(text = c("", "Fraction of proposals won", "Fraction of funds requested won"), 
                               cex = c(1.75,1.5,1.5),
                               font = c(0,0,0)   ))

text(x = complexClusterWins$count.rate,
     y = complexClusterWins$sum.rate,
     labels = row.names(complexClusterWins),
     cex = 2,
     font = 2,
     col = complexClusterWins$color
) 

mtext(text = "Sum rate and count rate" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

```

```{r}

# Tiled rate scatter plots of complex clusters

par(oma = c(5,7,4,1), mfrow = c(3,2))

invisible(
  lapply(levels(piEmplid$complex_cluster), function(x){ 
    
    
    
    yBYx(data = piEmplid[piEmplid$complex_cluster == x,],  
         x_axis = "count.rate", 
         y_axis = "sum.rate", 
         detailMapping = clusterComplexMapping, 
         log = NA, 
         scatter_params = list(mar = c(0,0,0,0)),
         scatter_args = list(cex = 1.618, xlim = c(0,1), ylim = c(0,1)),
         scatter_text_args = list(
           labels = "",
           cex = 1.618),
         scatter_mtext_args = list(text = c("", "", "") )
         #                              cex = c(1.75,1.5,1.5),
         #                              font = c(0,0,0)   )
    )
    
    textFilter <- row.names(complexClusterWins) == x
    #text(x = totalsClusterWins$count.rate[textFilter],
    #    y = totalsClusterWins$sum.rate[textFilter],
    #    labels = row.names(totalsClusterWins)[textFilter],
    #    cex = 2,
    #    font = 2,
    #    col = totalsClusterWins$color[textFilter]
    #    )   
    
    text(x = 0.05,
         y = 0.9,
         labels = row.names(complexClusterWins)[textFilter],
         cex = 3,
         font = 2,
         col = complexClusterWins$color[textFilter],
         xpd = TRUE
    )  
    
    
  })    
)    



# mtext(text = "Clusters based on multiple dimensions" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

mtext(text = "Fraction of funds requested won" , side = 2, outer = TRUE, font = 1, cex = 1.5, line = 3) 

mtext(text = "Fraction of proposals won" , side = 1, outer = TRUE, font = 1, cex = 1.5, line = 3) 

```

### 8. VARIABLE RELEVANCE PER CLUSTER

```{r}

# Variable relevance per cluster

oldPar <- par(oma = c(0,0,4,1),
              bg = "ivory",
              fg = "gray10",
              mfrow = c(3,2), 
              mar = c(3,7,0,0))

invisible(
  lapply(levels(piEmplid$complex_cluster), function(x){ 
 
    # Extract v.test values for the cluster
    vtest_values <- complexHCPC$desc.var$quanti[[x]][, "v.test"]
         
    # Generate a color gradient using viridis, centered around 0
    colors <- viridis(length(vtest_values), option = "G")  # "C" is a blue-to-yellow scale
             
    # Order values for proper coloring
    ordered_idx <- order(vtest_values)  
    sorted_values <- vtest_values[ordered_idx]  
    sorted_colors <- colors[rank(sorted_values, ties.method = "first")]  # Match colors to sorted values
    
    complexHCPC$desc.var$quanti[[x]][,"v.test"] |> 
      (\(x){ 
      barplot(
        rev(x),
        las = 2,
        horiz = TRUE,
        col = sorted_colors)   
      })()
        
    legend("topleft", legend = x, bty = "n", text.font = 2, cex = 2, text.col = complexClusterWins[as.numeric(x),"color"])    
})
)

mtext(side = 3, 
      font = 2, 
      cex = 2, 
      text = "Variable relevance per cluster",
      col = "gray10",
      line = 0.618,
      outer = TRUE)
```


```{r, eval = FALSE}

# rather than create an aggregate of piEmplid per cluster (which is the accurate way to represent how the cluster was calculated), I added a "population" column to complexClusterWins and I'll display that instead

# simple metrics per cluster

count.per <- aggregate(emplid ~ complex_cluster,  data = piEmplid, length )
median.per <- aggregate(cbind(win.median, loss.median) ~ complex_cluster,  data = piEmplid, median )
sum.per <- aggregate(cbind(win.sum, loss.sum) ~ complex_cluster,  data = piEmplid, sum )
count.per <- aggregate(cbind(win.count, loss.count) ~ complex_cluster,  data = piEmplid, sum )
  
median.sum <- aggregate(cbind(win.sum, loss.sum) ~ complex_cluster,  data = piEmplid, median )
median.rate <- aggregate(cbind(count.rate, sum.rate) ~ complex_cluster,  data = piEmplid, median )

# chat gave me this

cluster_summary <- by(piEmplid[, c("emplid", "win.median", "loss.median", 
                           "win.sum", "loss.sum", "win.count", 
                           "loss.count", "count.rate", "sum.rate")], 
             piEmplid$complex_cluster, 
             function(x) c(population = nrow(x), 
                           pop_perc = round((nrow(x)/nrow(piEmplid))*100,0),
                           median_win = round(median(x$win.median), -3) ,  
                           median_loss = round(median(x$loss.median), -3), 
                           sum_win = round(sum(x$win.sum),-6) , 
                           sum_loss = round(sum(x$loss.sum), -6), 
                           count_win = sum(x$win.count), 
                           count_loss = sum(x$loss.count), 
                           median_count_rate = round(median(x$count.rate),2), 
                           median_sum_rate = round(median(x$sum.rate),2) ) ) |>
  (\(x){do.call(rbind,x) })()
  

# Load required packages
library(flextable)

# Convert result_df to a flextable
ft <- flextable::flextable(data.frame(complex_cluster = rownames(cluster_summary), cluster_summary)) %>%
  flextable::theme_vanilla() %>%  # Apply a clean theme
  flextable::set_caption("Aggregated Metrics by Complex Cluster") %>%  # Add a title
  flextable::colformat_num(j = 2:ncol(cluster_summary), digits = 2) %>%  # Format numeric columns
  flextable::autofit()  # Adjust column widths for better display

# Print the table
ft

# I actually have all but the population in the "complexClusterWins". Well, except the calculation of the rates and medians have a different base -- per proposal instead of per emplid, so it's not exactly the same thing.

# the cluster summary is a summary of what was actually clustered (piEmplid, per PI) while the complexClusterWins goes back to the original per-proposal data that piEmplid was aggregated from.

# I should just merge the population to complexClusterWins and use that in my table, or figure out a function to do it when I calculcate complexClusterWins



```


### 9. CLUSTERS OVER TIME

Although this clustering did not include any time variables, the submission patterns by cluster are shown.

```{r fig.height = 4}

# I need to fix the timeDots graphics before I can get this to tile.  I should  probably leave this out.

piYr <- elapsed(data = prepData[!is.na(prepData$complex_cluster), ])


# I haven't improved my timeline graphics yet

par(mfrow = c(1,1), bg = "ivory", fg = "gray10",
    mar = c(0,0,0,0))

invisible(
lapply(levels(prepData$complex_cluster), function(x){ 

  filter <- piYr$each[,"PROPOSAL_PI_EMPLID"] %in% piEmplid$emplid[piEmplid$complex_cluster == x]
  
# par(new = TRUE)  
timeline(data = piYr$each[filter,])
mtext(x, side = 3, line = -2, cex = 2, font = 2, outer = FALSE, col = complexClusterWins$color[as.numeric(x)])
  
  })
)



```

### 10. CLUSTER POPULATIONS BY COLLEGE AND MAJOR INSTITUTIONS

 `r nrow(collegePopulation) - length(unique(collegePopulation[,1]))` (`r  ( (nrow(collegePopulation) - length(unique(collegePopulation[,1]))) /length(unique(collegePopulation[,1])) ) |> (\(x){round(x*100, 1)})()`)%  PI's appear in multiple colleges or institutions.


```{r}

table(collegePopulation[,c("college", "complex_cluster")]) |>
  as.data.frame.matrix() |>
  apply(1, proportions) |>
  t() |>
  pheatmap(
    scale = "column",
  #  Rowv = NA,
  #  Colv = NA,
    main = "Organization by cluster",
  #  margins = c(3,3)
  fontsize_col = 15,
  angle_col = 0
  )


# that's pretty much what I want

# let's publish the table now

collegeTable <- table(collegePopulation[,c("college", "complex_cluster")]) |>
  as.data.frame.matrix() |>
  apply(1, proportions) |>
  t() |>
  as.data.frame() |>
  (\(x){round(100*x,1) })() |>
  (\(x){cbind(Org = rownames(x), x) })() |>  #
  flextable::flextable() |>
  flextable::theme_vanilla() |>  # Apply a clean theme
  flextable::set_caption("Percent of organization population per cluster") |>  # Add a title
  flextable::colformat_num(j = 2:6, digits = 2) |>  # Format numeric columns
  flextable::autofit()  # Adjust column widths for better display

# Print the table
collegeTable

# man that's beautiful


# let's reverse the table -- 

table(collegePopulation[,c("college", "complex_cluster")]) |>
  as.data.frame.matrix() |>
  apply(2, proportions) |>
  t() |>
  pheatmap(
    scale = "column",
  #  Rowv = NA,
  #  Colv = NA,
    main = "Cluster by organization",
  #  margins = c(3,3)
  las = 3
  )


collegeTable <- table(collegePopulation[,c("college", "complex_cluster")]) |>
  as.data.frame.matrix() |>
  apply(2, proportions) |>
  t() |>
  as.data.frame() |>
  (\(x){round(100*x,1) })() |>
  (\(x){cbind(Org = rownames(x), x) })() |>  #
  flextable::flextable() |>
  flextable::theme_vanilla() |>  # Apply a clean theme
  flextable::set_caption("Percent of cluster population per organization") |>  # Add a title
  flextable::colformat_num(j = 2:6, digits = 2) |>  # Format numeric columns
  flextable::autofit()  # Adjust column widths for better display

# Print the table
collegeTable



```



```{r}

# Save "piEmplid" object

saveRDS(piEmplid, here::here("Robjects", "Clustering PIs", "Complex cluster from complex cluster child.rds"))

```