---
output:
  html_document:
    keep_md: true
---


```{r include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.height=7, fig.width=10)

# adjust graphic parameters

oldPar <- par(cex.main = 3,
    cex.lab = 3,
    cex.axis = 2,
    mar = c(5.1,4.1,4.1,2.1) # default is c(5.1,4.1,4.1,2.1)
    )

```




```{r}

##########
## LOAD ##
##########


source(here::here("Prep scripts", "Prep Script from Win Rates by College.R"))

###############
## LIBRARIES ##
###############

library(viridis)

######################
## CALCULATE PER PI ##
######################

piEmplid <- calculateWinRates(data = prepData, categoryColumn = "PROPOSAL_PI_EMPLID", functionList = list(mean = mean, median = median)) |> (\(x){ x[[1]] })()
# really starting to hate the list nature and why do I want "call" ?

piEmplid$count.total <- apply(piEmplid[,c("win.count","loss.count")], 1, sum, na.rm = TRUE )
piEmplid$sum.total <- apply(piEmplid[,c("win.sum","loss.sum")], 1, sum, na.rm = TRUE )

##########
## PREP ##
##########

piEmplid$emplid <- row.names(piEmplid)
filterThreeCount <- piEmplid$count.total > 3
piMap <- data.frame(college = row.names(piEmplid), abbrv = row.names(piEmplid), color = NA, pch = 19, cex = 0.3 )


```





```{r identify.natural.breaks}


proposal_count_clusters <- table(piEmplid$count.total) |>
  as.data.frame() |>
    (\(x){data.frame(proposals = as.numeric(as.character(x[,1])), pi = x[,2])  })() |>
  # scale() |>
  dist(method = "canberra") |>
  hclust() |>
  #plot() 
  cutree(k=8) 
  #cutree(h = 0.68)

# manual cleaning of the long tail
breaks <- which(!duplicated(proposal_count_clusters))

proposal_count_clusters[30:45] <- 6
proposal_count_clusters[46:84] <- 7


breaks <- which(!duplicated(proposal_count_clusters))

```

# Filtering based on the count of proposals per principal investigator

### EXECUTIVE SUMMARY

Using the natural breaks identified in clustering the counts of proposals per principal investigators, 1,264 principal investigators submitting three or fewer proposals are filtered out.  1,672 principal investigators submitting four or more proposals are kept for further clustering.

```{r}


# Display and put the breaks in

par(oma = c(0,0,3,0),
  mar = c(5,5,4,1),
    bg = "ivory",
    fg = "gray10")

table(piEmplid$count.total) |> 
  barplot(xaxt = "n",
          yaxt = "n")

rect(xleft = par("usr")[1], 
    ybottom = par("usr")[3], 
    xright = par("usr")[2], 
    ytop = par("usr")[4],
    col = "gray95", 
    border = NA)

grid(nx = NA, ny = NULL, col = "gray10", lwd = 0.618)

par(new = TRUE)

theBar <- table(piEmplid$count.total) |> 
  barplot(col = proposal_count_clusters,
          las = 1,
          cex.axis = 1.2,
          cex.names = 1.2
          )

mtext(text = c("Clusters shown", 
               "Count of PI's",
               "Number of proposals submitted per PI"
               ),
      side = c(3,2,1),
      font = c(1,1,1),
      cex = c(2,1.3,1.3),
      line = c(1.309,3.3,3)
      
      )


mtext(text = expression(bold("Count of PI's submitting " * italic(x) * " proposals")),
      side = 3,
      font = 2,
      cex = 2,
      line = 0.309,
      outer = TRUE
      )

text(x = theBar[breaks][1],
     y = table(piEmplid$count.total)[breaks][1] - 20,
     labels = paste0(table(piEmplid$count.total)[breaks][1], " PI's submitted ", breaks[1], " proposal."),
     col = "red",
     pos = 4,
     cex = 1.0,
     xpd = TRUE)

text(x = theBar[breaks][2:4],
     y = table(piEmplid$count.total)[breaks][2:4],
     labels = paste0(table(piEmplid$count.total)[breaks][2:4], " PI's submitted ", breaks[2:4], " proposals."),
     col = "red",
     pos = 4,
     cex = 1.0,
     xpd = TRUE)

text(x = theBar[breaks][5:(length(breaks))],
     y = c(45,30,25),    # table(piEmplid$count.total)[breaks][5:(length(breaks))],
     labels = table(piEmplid$count.total)[breaks][5:(length(breaks))],
     col = "red",
     pos = 4,
     cex = 1.2,
     xpd = TRUE)

legend("topright",
       pch = 15,
       pt.cex = 2,
       col = 1:8,
       legend = c("1","2 to 3", "4 to 7", "8 to 19", "20 to 29", "30 to ~45", ">46"),
       title = "Clusters\nProposal counts per PI",
       bty = "n")

```


```{r}

# filter graphic

par(oma = c(0,0,0,0),
    mar = c(5,5,4,1),
    bg = "ivory",
    fg = "gray10")

table(piEmplid$count.total) |> 
  barplot(xaxt = "n",
          yaxt = "n")

rect(xleft = par("usr")[1], 
    ybottom = par("usr")[3], 
    xright = par("usr")[2], 
    ytop = par("usr")[4],
    col = "gray95", 
    border = NA)

grid(nx = NA, ny = NULL, col = "gray10", lwd = 0.618)

par(new = TRUE)

theBar <- table(piEmplid$count.total) |> 
  barplot(col = c(rep("coral",2), rep("darkcyan",4) ) [proposal_count_clusters],
          las = 1,
          cex.axis = 1.2,
          cex.names = 1.2
          )

mtext(text = c("Filter shown",
               "Count of PI's",
               "Number of proposals submitted per PI"
               ),
      side = c(3,2,1),
      font = c(2,1,1),
      cex = c(2,1.3,1.3),
      line = c(1.309,3.3,3)
      
      )

text(x = theBar[breaks][2], 
     y = table(piEmplid$count.total)[breaks][2] - 20,
     labels = paste( format(sum(table(piEmplid$count.total)[1:3]),big.mark = ","), "PIs submitting three or fewer proposals are removed"),
     col = "coral",
     pos = 4,
     cex = 1.2,
     xpd = TRUE  
       )

text(x = theBar[breaks][3], 
     y = table(piEmplid$count.total)[breaks][3] - 20,
     labels = paste( format(sum(table(piEmplid$count.total)[4: length(table(piEmplid$count.total))]),big.mark = ",") , "PIs submitting four or more proposals are kept"),
     col = "darkcyan",
     pos = 4,
     cex = 1.2,
     xpd = TRUE  
       )


legend("topright",
       pch = 15,
       pt.cex = 2,
       col = c("coral","darkcyan"),
       legend = c("Removed","kept"),
       title = "Filter",
       bty = "n")



```

