---
output:
  html_document:
    keep_md: true
---

```{r include = FALSE}

#-- I'd like some visualizations that walk you through how I've identified my low -performers
#-- I'd like some visualizations that clearly overlap the three methods -- pareto, rate, and complex --
#   ...and clarify what some of these overlaps mean.
#-- I'd like to change "many at bats, few hits" cluster name to start with the letter S.
#   ...then, there's a subset of this one that I can call "inefficient (?)" and call out their high total despite their low win rates.  Curious in itself.  This would be >80% performers.

#-- I'd like to call out that the strike-outs have a very high median submission.  Delusional?  Not serious?

#-- the timelines.... not sure what they are adding to the conversation.  Lots of graphs are experimental here and not adding a lot.  

#-- the Sum of funds requested break-out should be 1 column of three rows

#-- my Pareto child should have 10 pareto clusters for ten percentiles, and show the same charts as I did the others.

# And I would've could've should've adjusted for inflation since it's been ten years.
# Not too late, I guess



# Maybe my "At Bat" cluster is just called "Inefficient" -- becuase aren't I calling my "Cluster 5" my high performers?  I mean, at least some of Cluster 5 are high performers.  Where's the dividing line between "high" and "low" performers in cluster 5?  I think I want to look at that "At Bat" overlap a little closer.

# How big of a gap is it between Complex 1 (no wins) and Complex 2 (low wins) ?  Is it mild changes by degree, or is there a clear gap?

# An idea for a graph is to plot the Complex Clusters on top of the percentile bands.  I might need to change pch for added clarity.  I'm not sure if I want to show Win Rates by Funds Requested Won colored by percentile .... that's a lot to walk through (?)  Sounds like a report?  Maybe put them side by side, with a line over the first one?

# I am going to need a philosophical discussion on who is a "low performer."

# This report is going to need a pretty big over-haul in flow and graphics.


# I have copied and pasted from "Comparing clusters" child

```



```{r include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.height=7, fig.width=10)

# adjust graphic parameters

oldPar <- par(cex.main = 3,
    cex.lab = 3,
    cex.axis = 2,
    mar = c(5.1,4.1,4.1,2.1) # default is c(5.1,4.1,4.1,2.1)
    )

```



```{r}

##########
## LOAD ##
##########

# PREP SCRIPT

source(here::here("Prep scripts", "Adjusting prepData and loading things.R"))

# PER PI

piClusters <- lapply(list.files(here::here("Robjects", "Clustering PIs"), full.names = TRUE), readRDS)

###########
## MERGE ## 
###########

piEmplid <- Reduce(function(x, y) {
  merged <- merge(x, y, by = "emplid", all = FALSE)
  merged <- merged[, !duplicated(sub("\\.x$|\\.y$", "", names(merged)))]  # Remove duplicate columns
  
  # Rename columns to remove ".x"
  names(merged) <- sub("\\.x$", "", names(merged))
  
  merged
}, piClusters)
# piEmplid <- Reduce(function(x, y) merge(x, y, by = "emplid", all = FALSE), piClusters)
# piEmplid <- do.call(merge, piClusters)



prepData <- merge(prepData, piEmplid[,c("emplid","complex_cluster","rate_cluster")], by.x =  "PROPOSAL_PI_EMPLID", by.y = "emplid", all.x = TRUE)

##########
## PREP ##
##########

piEmplid$combined_cluster <-  factor(paste(piEmplid$complex_cluster, piEmplid$rate_cluster, sep = ", "))

prepData$combined_cluster <- factor(paste(prepData$complex_cluster, prepData$rate_cluster, sep = ", "))

###############
## LIBRARIES ##
###############

library(viridis)
library(pheatmap)

# I'll probably want my per-cluster wins as well

# I am going to want to bring in the colors

filterThreeCount <- piEmplid$count.total > 3

piMap <- data.frame(college = row.names(piEmplid), abbrv = row.names(piEmplid), color = NA, pch = 19, cex = 0.7 )


```



```{r}
             
# Manage colors

#############
## COMPLEX ##
#############

complexClusterColors <- c("forestgreen", "deepskyblue", "goldenrod",  
                            "firebrick", "darkslategray", "chartreuse",  
                            "slateblue", "darkkhaki", "coral")

clusterComplexMapping <- piMap
clusterComplexMapping[,"pch"] <- rep(19, nrow(clusterComplexMapping))

clusterComplexMapping[,"color"] <- complexClusterColors [piEmplid$complex_cluster]

# clusterComplexMapping[,"color"] <- complexClusterColors[complexHCPC$data.clust$clust[match(clusterComplexMapping[,"college"], row.names(complexHCPC$data.clust))] ]

##############
## COMBINED ##
##############

combinedClusterColors <- c("forestgreen", "deepskyblue", "goldenrod",  
  "firebrick", "darkslategray", "chartreuse",  
  "slateblue", "darkkhaki", "coral",  
  "mediumorchid", "dodgerblue", "tomato",  
  "orchid", "darkseagreen", "sienna",  
  "royalblue", "indianred", "seagreen",  
  "peru", "cadetblue", "plum",  
  "midnightblue", "lawngreen", "darkorange",  
  "lightsteelblue")

clusterCombinedMapping <- piMap
clusterCombinedMapping[,"pch"] <- rep(19, nrow(clusterCombinedMapping))

clusterCombinedMapping[,"color"] <- combinedClusterColors [piEmplid$combined_cluster]




################
## PERCENTILE ##
################

percentileColors <- viridis(10,direction = -1)[c(1:7,10)]
percentileMapping <- piMap

percentileMapping[,"pch"] <- rep(19, nrow(percentileMapping))

percentileMapping[,"color"] <- percentileColors[as.numeric(cut(piEmplid$win.sum_percentile, breaks = seq(0.2,1,by=0.1)))]

```



# Low Performers

Low performers are in Rates Cluster 1 (<20% win rates by count and sum.)

**Strike outs:**

These PI's have never won a proposal.  Strike-outs are located in the overlap with Complex Cluster 3 ("Pipe Dreams").

**Sputtering:**

These PI's submit a decent number of proposals with relatively few wins.

They are located in the overlap with Complex Cluster 4 ("Plucky"). 


**Many at bats, few hits:**

These PI's submit a large number of proposals with relatively few wins.

They are located in the overlap with Complex Cluster 5 ("Prolific"). 

# "Inefficent" Performers

These PI's simultaneously carry less than a 20% win rate while also being a Top 20% performer.  They contribute to 80% of the total funds won over the last ten years.  (Maybe exclude them from the other clusters?)


Notes:

Try showing the boxplots, win and loss, for each of these sub-categories

I need to set the same scale for timeDots
I need to modify timeDots according to my new and improved functions
I should show the tick marks (timeLine) as well as the timeDots

```{r}

# Elapsed calculations

piYr <- elapsed(data = prepData)
piYrMon <- elapsedCash(data = prepData)

# clYr <- elapsed(target = "complex_cluster", data = prepData )
# clYrMon <- elapsedCash(target = "complex_cluster", data = prepData )

```


```{r}

# strike-out
strikeOutPI <- piEmplid$emplid[piEmplid$rate_cluster == 1 & piEmplid$complex_cluster == 3]

strikeOutFilter <- piYrMon$each$PROPOSAL_PI_EMPLID %in% strikeOutPI

timeDots(piYrMon$each[strikeOutFilter,])


```

```{r}

# sputtering
sputteringPI <- piEmplid$emplid[piEmplid$rate_cluster == 1 & piEmplid$complex_cluster == 4]

sputteringFilter <- piYrMon$each$PROPOSAL_PI_EMPLID %in% sputteringPI

timeDots(piYrMon$each[sputteringFilter,])

```

```{r}

# at Bat
atBatPI <- piEmplid$emplid[piEmplid$rate_cluster == 1 & piEmplid$complex_cluster == 5]

atBatFilter <- piYrMon$each$PROPOSAL_PI_EMPLID %in% atBatPI

timeDots(piYrMon$each[atBatFilter,])

# at bat might be better represented with tick marks

```

```{r}

# Do it with one lapply

par(mfrow = c(3,1), mar = c(0,10,0,1), bg = "ivory", fg = "gray10" )

invisible(
lapply(c(3:5), function(x){
 
  # identify PI's
  thePIs <- piEmplid$emplid[piEmplid$rate_cluster == 1 & piEmplid$complex_cluster == x]
  
  # define a filer
  theFilter <- piYrMon$each$PROPOSAL_PI_EMPLID %in% thePIs

  # draw the graph
  timeDots(piYrMon$each[theFilter,])  
  
  
})
)

```


What about people with a low rate of wins, but have a high total sum?  At the bottom right of "Funds requested" graph?

Maybe I have two criteria for low performers --
(1)  Low win rate (what I have so far)
(2)  Low total funds requested won (who would this be?  I guess I want a pareto already!)

I need to see these by "total funds requested won" per PI


```{r}

# I need a detailMapping

# Plot by win sum and rate clusters

par(oma = c(0,3,4,0), mfrow =c(1,1))

yBYx(data = piEmplid[piEmplid$emplid %in% c( sputteringPI, strikeOutPI,atBatPI),],  
     x_axis =  "win.sum", 
     y_axis = "sum.rate", 
     detailMapping = clusterComplexMapping, 
     log = c("x"), 
     scatter_args = list(xlim = c(0,18.5)),
     scatter_text_args = list(
       labels = "",
       cex = 1.618),
     scatter_mtext_args = list(text = c("", "Sum of funds requested won (log)", "Fraction of funds requested won"), 
                               cex = c(1.75,1.5,1.5),
                               font = c(0,0,0)   ))

mtext(text = "Sum rate by total funds won" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

# text(x = log(medianPIbyCluster$win.sum) |> (\(x){ x[is.infinite(x)] <- 0; return(x)})() ,
#     y = combinedClusterWins$sum.rate,
#     labels = 1:(length(complexHCPC[["desc.var"]][[2]])),
#     cex = 2,
#     font = 2,
#     col = combinedClusterWins$color,
#     xpd = TRUE
# ) 

```


```{r fig.width=10}

# Tiled sum scatter plots of complex clusters

par(oma = c(5,7,4,1), mfrow = c(3,2))

invisible(
  lapply(3:5, function(x){ 
    
    yBYx(data = piEmplid[piEmplid$rate_cluster == 1 & piEmplid$complex_cluster == x,],  
         x_axis = "win.sum", 
         y_axis = "sum.rate", 
         detailMapping = clusterComplexMapping, 
         log = c("x"), 
         scatter_params = list(mar = c(0,0,0,0)),
         scatter_args = list(cex = 1.618, xlim = c(0,20), ylim = c(0,1)),
         scatter_text_args = list(
           labels = "",
           cex = 1.618),
         scatter_mtext_args = list(text = c("", "", "") )
         #                              cex = c(1.75,1.5,1.5),
         #                              font = c(0,0,0)   )
    )
    
    # textFilter <- row.names(combinedClusterWins) == x
    #text(x = totalsClusterWins$count.rate[textFilter],
    #    y = totalsClusterWins$sum.rate[textFilter],
    #    labels = row.names(totalsClusterWins)[textFilter],
    #    cex = 2,
    #    font = 2,
    #    col = totalsClusterWins$color[textFilter]
    #    )   
    
    text(x = 0.05,
         y = 0.9,
         labels = x,
         cex = 3,
         font = 2,
         #col = combinedClusterWins$color[textFilter],
         xpd = TRUE
    )  
    
    
  })    
)    



# mtext(text = "Clusters based on multiple dimensions" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

mtext(text = "Sum of funds requested won (log)" , side = 1, outer = TRUE, font = 1, cex = 1.5, line = 3) 

mtext(text = "Fraction of funds requested won" , side = 2, outer = TRUE, font = 1, cex = 1.5, line = 3) 

```


I need to examine the values that have a very low win rate, but have still pulled in enough money
to be a significant contributor to the U

I can color the ones in the graph above according to 80th percentile ....?

It would be really cool to replace my background gray grid with transparently changing viridis bands according to the percentiles, or the cumulative percentile or something like that.

I need to include complex cluster 2, it's just incomplete otherwise.

And maybe we don't call these "low performers" but we call them "wasted effort" or something like that?

I mean, how can you call someone a low performer when they are in the 90th percentile for funds brought in?

And how can you call someone a low performer when they nabbed 75% of their proposals, but are in the cumulative 10th percentile?

"Low" would be a measure of performance against your own potential, or of your peers'.

I guess I'd want to carefully designate the "low" performers that are above the 80th percentile.  That's a curious sub-set.


```{r}

# Percentile mapping

par(oma = c(0,3,4,0), mfrow =c(1,1))

yBYx(data = piEmplid[piEmplid$emplid %in% c( sputteringPI, strikeOutPI,atBatPI),],  
     x_axis =  "win.sum", 
     y_axis = "sum.rate", 
     detailMapping = percentileMapping, 
     log = c("x"), 
    # scatter_args = list(xlim = c(0,18.5)),
     scatter_text_args = list(
       labels = "",
       cex = 1.618),
     scatter_mtext_args = list(text = c("", "Sum of funds requested won (log)", "Fraction of funds requested won"), 
                               cex = c(1.75,1.5,1.5),
                               font = c(0,0,0)   ))

mtext(text = "Sum rate by total funds won" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

# text(x = log(medianPIbyCluster$win.sum) |> (\(x){ x[is.infinite(x)] <- 0; return(x)})() ,
#     y = combinedClusterWins$sum.rate,
#     labels = 1:(length(complexHCPC[["desc.var"]][[2]])),
#     cex = 2,
#     font = 2,
#     col = combinedClusterWins$color,
#     xpd = TRUE
# ) 

```


```{r}

# Percentile mapping

par(oma = c(0,3,4,0), mfrow =c(1,1))

yBYx(data = piEmplid[1:nrow(piEmplid),], # no idea why it needs the subset to work.  
     x_axis =  "win.sum", 
     y_axis = "sum.rate", 
     detailMapping = percentileMapping, 
     log = c("x"), 
    # scatter_args = list(xlim = c(0,18.5)),
     scatter_text_args = list(
       labels = "",
       cex = 1.618),
     scatter_mtext_args = list(text = c("", "Sum of funds requested won (log)", "Fraction of funds requested won"), 
                               cex = c(1.75,1.5,1.5),
                               font = c(0,0,0)   ))

mtext(text = "Sum rate by total funds won" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

# text(x = log(medianPIbyCluster$win.sum) |> (\(x){ x[is.infinite(x)] <- 0; return(x)})() ,
#     y = combinedClusterWins$sum.rate,
#     labels = 1:(length(complexHCPC[["desc.var"]][[2]])),
#     cex = 2,
#     font = 2,
#     col = combinedClusterWins$color,
#     xpd = TRUE
# ) 

```

```{r}

# >80th percentile

par(oma = c(0,3,4,0), mfrow =c(1,1))

yBYx(data = piEmplid[piEmplid$win.sum_percentile >= 0.8  ,], 
     x_axis =  "win.sum", 
     y_axis = "sum.rate", 
     detailMapping = percentileMapping, 
     log = c("x"), 
    # scatter_args = list(xlim = c(0,18.5)),
     scatter_text_args = list(
       labels = "",
       cex = 1.618),
     scatter_mtext_args = list(text = c("", "Sum of funds requested won (log)", "Fraction of funds requested won"), 
                               cex = c(1.75,1.5,1.5),
                               font = c(0,0,0)   ))

mtext(text = "Sum rate by total funds won" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

# text(x = log(medianPIbyCluster$win.sum) |> (\(x){ x[is.infinite(x)] <- 0; return(x)})() ,
#     y = combinedClusterWins$sum.rate,
#     labels = 1:(length(complexHCPC[["desc.var"]][[2]])),
#     cex = 2,
#     font = 2,
#     col = combinedClusterWins$color,
#     xpd = TRUE
# ) 

```


```{r}

# "Inefficient"

par(oma = c(0,3,4,0), mfrow =c(1,1))

yBYx(data = piEmplid[piEmplid$win.sum_percentile >= 0.8 & piEmplid$rate_cluster ==1,], 
     x_axis =  "win.sum", 
     y_axis = "sum.rate", 
     detailMapping = percentileMapping, 
     log = c("x"), 
    # scatter_args = list(xlim = c(0,18.5)),
     scatter_text_args = list(
       labels = "",
       cex = 1.618),
     scatter_mtext_args = list(text = c("", "Sum of funds requested won (log)", "Fraction of funds requested won"), 
                               cex = c(1.75,1.5,1.5),
                               font = c(0,0,0)   ))

mtext(text = "Sum rate by total funds won" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

# text(x = log(medianPIbyCluster$win.sum) |> (\(x){ x[is.infinite(x)] <- 0; return(x)})() ,
#     y = combinedClusterWins$sum.rate,
#     labels = 1:(length(complexHCPC[["desc.var"]][[2]])),
#     cex = 2,
#     font = 2,
#     col = combinedClusterWins$color,
#     xpd = TRUE
# ) 

```



```{r}

# inefficient

inefficientPI <- piEmplid$emplid[piEmplid$rate_cluster == 1 & piEmplid$win.sum_percentile >= 0.8]

inefficientFilter <- piYrMon$each$PROPOSAL_PI_EMPLID %in% inefficientPI

timeDots(piYrMon$each[inefficientFilter,])

timeline(piYr$each[inefficientFilter,])


```



```{r eval=FALSE}

# experimenting with the background

barplot(100:110)

lapply(1:7, function(x) {rect(seq(par("usr")[1], par("usr")[2], length.out = 8 )[x], par("usr")[3], seq(par("usr")[1], par("usr")[2], length.out = 8 )[x+1], par("usr")[4], col = percentileColors[x]  ) 
 })

par(new = TRUE)

barplot(100:110, col = viridis::viridis(11, direction = 1))

# wow that's hard on the eyes

# dynamic cut

filter <- 1:nrow(piEmplid)
plot(log(piEmplid[filter, "win.sum"]),
     piEmplid[filter, "sum.rate"],
     col = percentileMapping[,"color"]
     )

# plotBreaks <- seq(1000,70000000,length.out = 9 )

plotBreaks <-
  piEmplid$win.sum |>
  (\(x){split(x,
              cut(piEmplid$win.sum_percentile, breaks = seq(0.2,1,by=0.1))
              )})() |>
  (\(x){ 
  sapply(x, max, na.rm=TRUE)
  })() |>
  log() |>
  (\(x){
    c(0,x)
  })()
  
lapply(1:(length(plotBreaks) - 1), function(i) {
  rect(plotBreaks[i], par("usr")[3], plotBreaks[i + 1], par("usr")[4], col = adjustcolor(percentileColors[i], alpha.f= 0.1 ) , border = NA)
})

par(new=TRUE)
plot(log(piEmplid[filter, "win.sum"]),
     piEmplid[filter, "sum.rate"],
     col = percentileMapping[,"color"]
     )

  
#  (cut(piEmplid$win.sum_percentile, breaks = seq(0.2,1,by=0.1)))


#  percentileColors[as.numeric(cut(piEmplid$win.sum_percentile, breaks = seq(0.2,1,by=0.1)))]

plotBreaks <- piEmplid$win.sum[cut(piEmplid$win.sum_percentile, breaks = seq(0.2,1,by=0.1))]
  
# this is pretty much what I'm looking for.

# Next, do I need to modify yBYx to enable this?

# I can figure that out later.  Let's use what I've got to finish up this report.


filter <- piEmplid$rate_cluster == 1
plot(log(piEmplid[filter, "win.sum"]),
     piEmplid[filter, "sum.rate"],
     col = percentileMapping[filter,"color"],
     xlab = "",
     ylab = "",
     xaxt = "n",
     yaxt = "n",
     plot = FALSE
     )


plotBreaks <-
  piEmplid$win.sum |>
  (\(x){split(x,
              cut(piEmplid$win.sum_percentile, breaks = c(0,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1), include.lowest=TRUE)

              # cut(piEmplid$win.sum_percentile, breaks = seq(0.2,1,by=0.1))
              )})() |>
  (\(x){ 
  sapply(x, max, na.rm=TRUE)
  })() |>
  log() |>
  (\(x){
    c(0,x)
  })()

lapply(1:(length(plotBreaks) - 1), function(i) {
  rect(plotBreaks[i], par("usr")[3], plotBreaks[i + 1], par("usr")[4], col = adjustcolor(percentileColors[i], alpha.f= 0.1 ) , border = NA)
})

par(new=TRUE)
plot(log(piEmplid[filter, "win.sum"]),
     piEmplid[filter, "sum.rate"],
     col = percentileMapping[filter,"color"],
     xlab = "",
     ylab = "",
     las = 1,
     )

legend("topleft",
       title = "Percentile ranges",
       col = percentileColors,
       #legend = names(plotBreaks[-1]),
       legend = paste0("<=",seq(from =0.3, to = 1, by = 0.1)),
       pch = 15,
       pt.cex = 3
       )

mtext(text = c("Some PI's with a low win rate\nare in the top 20% of funds requested won",
              "Win rate",
              "Total funds requested won (log)"),
      side = c(3,2,1) ,
      font = c(2,1,1) ,
      line = c(0,3,4),
      cex = c(1.5,1,1)
)


```


```{r}


filter <- piEmplid$rate_cluster == 1 & piEmplid$complex_cluster == 5
plot(log(piEmplid[filter, "win.sum"]),
     piEmplid[filter, "sum.rate"],
     col = percentileMapping[filter,"color"],
     xlab = "",
     ylab = "",
     xaxt = "n",
     yaxt = "n",
     plot = FALSE
     )


plotBreaks <-
  piEmplid$win.sum |>
  (\(x){split(x,
              cut(piEmplid$win.sum_percentile, breaks = c(0,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1), include.lowest=TRUE)

              # cut(piEmplid$win.sum_percentile, breaks = seq(0.2,1,by=0.1))
              )})() |>
  (\(x){ 
  sapply(x, max, na.rm=TRUE)
  })() |>
  log() |>
  (\(x){
    c(0,x)
  })()

invisible(
lapply(1:(length(plotBreaks) - 1), function(i) {
  rect(plotBreaks[i], par("usr")[3], plotBreaks[i + 1], par("usr")[4], col = adjustcolor(percentileColors[i], alpha.f= 0.1 ) , border = NA)
})
)

par(new=TRUE)
plot(log(piEmplid[filter, "win.sum"]),
     piEmplid[filter, "sum.rate"],
     col = percentileMapping[filter,"color"],
     xlab = "",
     ylab = "",
     las = 1,
     )

legend("topleft",
       title = "Percentile ranges",
       col = percentileColors,
       #legend = names(plotBreaks[-1]),
       legend = paste0("<=",seq(from =0.3, to = 1, by = 0.1)),
       pch = 15,
       pt.cex = 3
       )

mtext(text = c("Some PI's with a low win rate\nare in the top 20% of funds requested won",
              "Win rate",
              "Total funds requested won (log)"),
      side = c(3,2,1) ,
      font = c(2,1,1) ,
      line = c(0,3,4),
      cex = c(1.5,1,1)
)


```


```{r}

# Should I identify any "at bat" as low performers?

hist(piEmplid$win.sum_percentile,
     col = c(rep(percentileColors, each = 2)))
     #col = c(percentileColors[1], rep(percentileColors[-1], each = 2) ) )
# almost got the colors right.

# I think I'm comfortable drawing a line at the 80th percentile
abline(v = 0.8, col = "red", lwd = 3)

# should I show this histogram?  Not sure what it's saying.

# and how much of Complex 5 is in Rate 1 again?  Looks like 7%, so I can comfortably isolate these as a kind of "low performers"

```



```{r}

# Calculate win rates

combinedClusterWins <- calculateWinRates(prepData, categoryColumn = "combined_cluster", functionList = list(mean = mean, median = median)) |>
  (\(x){
    return(x$summary[match(levels(prepData$combined_cluster), row.names(x$summary)),])
  })()

combinedClusterWins$color <- combinedClusterColors

combinedClusterWins$population <- aggregate(emplid ~ complex_cluster,  data = piEmplid, length )[,2]

combinedClusterWins$pop_perc <- round(100*(combinedClusterWins$population / sum(combinedClusterWins$population)),1)


combinedClusterWins$sum_perc <- round(100*(combinedClusterWins$win.sum/sum(combinedClusterWins$win.sum)),1)

combinedClusterWins$count_total <- apply(combinedClusterWins[,c("win.count","loss.count")], 1, sum)

combinedClusterWins$count_perc <- round(100*(apply(combinedClusterWins[,c("win.count","loss.count")], 1, sum)/sum(apply(combinedClusterWins[,c("win.count","loss.count")], 1, sum))),1)


medianPIbyCluster <- aggregate(win.sum ~ complex_cluster, data = piEmplid, median) # for plot

```


```{r fig.height = 4.5, fig.width = 9}

#########################################
## THREE PROPORTION BARS AND A BOXPLOT ##
#########################################

filter <- row.names(combinedClusterWins) %in% c("3, 1", "4, 1", "5, 1")

####################
## POPULATION BAR ##
####################

par(bg = "ivory",
    fg = "gray10",
    oma = c(0,0,3,0),
    mar = c(3,3,3,1),
    mfrow= c(1,2))

# dummy graph to create the background

theBat <- barplot(
  combinedClusterWins[filter,"population"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  combinedClusterWins[filter,"population"],
  col = combinedClusterWins[filter,"color"],
  las = 2
)

mtext(text = row.names(combinedClusterWins)[filter],
      line = 1,
      at = theBar,
      side = 1,
      cex = 0.9,
      font =2,
      las = 2,
      col = combinedClusterWins[filter,"color"])

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Population"
      )

text(x = theBar,
     y = combinedClusterWins[filter,"population"],
     pos = c(3,1,3),
     font = 2,
     col = c(rep("gray10",4),"gray90"),
     labels = paste0(combinedClusterWins$pop_perc[filter],"%")
     )


###############
## FUNDS BAR ##
###############

# adjust margins

par(mar = c(3,5,3,1))

# dummy graph to create the background

theBat <- barplot(
  combinedClusterWins[filter,"win.sum"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  combinedClusterWins[filter,"win.sum"],
  col = combinedClusterWins[filter,"color"],
  las = 2
)

mtext(text = row.names(combinedClusterWins),
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font = 2,
      col = combinedClusterWins[filter,"color"]
        )

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Funds requested won"
      )

text(x = theBar,
     y = combinedClusterWins[filter,"win.sum"],
     pos = c(3,1,3),  #,3,1),
     cex = c(1,1,1), #1,0.9),
     font = 2,
     col = c(rep("gray10",4),"gray90"),
     labels = paste0(combinedClusterWins$sum_perc[filter],"%")
     )

# float the median over THAT !

mtext(text = "Proportions and proposal amounts per cluster",
      side = 3,
      line = 0.309,
      font = 2,
      cex = 2,
      outer = TRUE)

###############
## COUNT BAR ##
###############

# adjust margins

par(mar = c(3,4,3,1),
    oma = c(0,0,0,0))

# par(bg = "ivory",
#    fg = "gray10",
#    oma = c(0,0,0,0),
#    mar = c(3,3,3,1)
#    )

# dummy graph to create the background

theBat <- barplot(
  combinedClusterWins[filter,"count_total"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  combinedClusterWins[filter,"count_total"],
  col = combinedClusterWins[filter,"color"],
  las = 2
)

mtext(text = row.names(combinedClusterWins)[filter],
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font = 2,
      col = combinedClusterWins[filter,"color"]
        )

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Count of proposals"
      )

text(x = theBar,
     y = combinedClusterWins[filter,"count_total"],
     pos = c(3,3,3,3,1),
     cex = c(1,1,1,1,0.9),
     font = 2,
     col = c(rep("gray10",4),"gray90"),
     labels = paste0(combinedClusterWins$count_perc[filter],"%")
     )


#############
## BOXPLOT ##
#############

par(mar = c(3,4,3,1))

# par(bg = "ivory",
#    fg = "gray10",
#    oma = c(0,0,0,0),
#    mar = c(3,4,4,1))
    
#    ,
#    mfrow= c(1,1))

boxplot(log(PROPOSAL_TOTAL_SPONSOR_BUDGET) ~ combined_cluster,
        data = prepData, #[prepData$combined_cluster %in% c("3, 1","4, 1","5, 1"),],
        names = rep("",25),
        yaxt = "n",
        ylab = ""
        )

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

boxColors <- rep("lightgray", nrow(combinedClusterWins))
colorFilter <- row.names(combinedClusterWins) %in% c("3, 1","4, 1","5, 1")
boxColors[colorFilter] <- combinedClusterWins[colorFilter, "color"]

theBox <- boxplot(log(PROPOSAL_TOTAL_SPONSOR_BUDGET) ~ combined_cluster,
        data = prepData,
        col = boxColors,
        names = rep("",25),
        ylab = "",
        las = 1
        )

mtext(text = c("Proposal funds requested", "Funds requested (log)" ),
      side = c(3,2),
      font = c(1,1),
      cex = c(1.75,1.5),
      line = c(0.309, 2.427)
      
      )

mtext(text = row.names(combinedClusterWins)[colorFilter],
      line = 1,
      at = (1:nrow(combinedClusterWins))[colorFilter],
      side = 1,
      cex = 0.9,
      font = 2,
      las = 2,
      col = combinedClusterWins[colorFilter,"color"]
        )

```



