---
output:
  html_document:
    keep_md: true
---

```{r include = FALSE}

# Notes

# This follows "Quick Pareto of winning funds requested.Rmd" and "Complex cluster of PIs child.Rmd"

```



```{r include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.height=10, fig.width=9)

# adjust graphic parameters

oldPar <- par(cex.main = 3,
    cex.lab = 3,
    cex.axis = 2,
    mar = c(5.1,4.1,4.1,2.1) # default is c(5.1,4.1,4.1,2.1)
    )

```


```{r}

##########
## LOAD ##
##########


source(here::here("Prep scripts", "Prep Script from Win Rates by College.R"))

###############
## LIBRARIES ##
###############

library(viridis)
library(pheatmap)

######################
## CALCULATE PER PI ##
######################

piEmplid <- calculateWinRates(data = prepData, categoryColumn = "PROPOSAL_PI_EMPLID", functionList = list(mean = mean, median = median)) |> (\(x){ x[[1]] })()
# really starting to hate the list nature and why do I want "call" ?

piEmplid$count.total <- apply(piEmplid[,c("win.count","loss.count")], 1, sum, na.rm = TRUE )
piEmplid$sum.total <- apply(piEmplid[,c("win.sum","loss.sum")], 1, sum, na.rm = TRUE )

##########
## PREP ##
##########

piEmplid$emplid <- row.names(piEmplid)
filterThreeCount <- piEmplid$count.total > 3
piMap <- data.frame(college = row.names(piEmplid), abbrv = row.names(piEmplid), color = NA, pch = 19, cex = 0.3 )

################
## PERCENTILE ##
################

piEmplid$win.sum_prop <- proportions(piEmplid$win.sum)
piEmplid$win.sum_rank <- rank(-piEmplid$win.sum,na.last = "keep", ties.method = "min")
ecdf_fun <- ecdf(piEmplid$win.sum)
piEmplid$win.sum_percentile <- ecdf_fun(piEmplid$win.sum)

#####################
## USEFUL FUNCTION ##
#####################

pickIndex <- function(percentageTarget = 0.8, propColumn) {
  
 index_value <-  which(cumsum(propColumn[order(propColumn, decreasing = TRUE)]) > percentageTarget-0.01 & cumsum(propColumn[order(propColumn, decreasing = TRUE)]) < percentageTarget + 0.01 ) |>
    (\(x)({
      x[x %in% floor(median(x))]
    }))()
  
  return(index_value)
  
}

```


**EXECUTIVE SUMMARY:** 

Over the last ten years, the U has requested some ~\$5.5 B that received some level of support or funding, and are called "won" for this report.

Only `r pickIndex(propColumn = piEmplid[,"win.sum_prop"])` PI's, or `r round(100*pickIndex(propColumn = piEmplid[,"win.sum_prop"])/length(piEmplid[,"win.sum_prop"]),0)`% of all PI's, are responsible for 80% of these 'won' funds requested.

This follows a standard 80/20 rule or Pareto principle, and suggests prioritizing principal investigators based on their productivity.


```{r pareto.plot}


#################
## PARETO PLOT ##
#################

par(mar = c(6,6,7,1)) # c(5.1,4.1,4.1,2.1)

plot(cumsum(piEmplid[,"win.sum_prop"][order(piEmplid[,"win.sum_prop"], decreasing = TRUE)]),
     yaxt = "n",
     ylab = "",
     #ylab = "Fraction of total requested funds",
     xaxt = "n",
     xlab = "",
     # xlab = "Count of PI's",
     
     panel.first = list(
       rect(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[4], col = "gray80", border = NA ),
     grid(col="gray90", lwd = 5, lty="solid")
     )
     
     
     )

# 80%
points(x = pickIndex(propColumn = piEmplid[,"win.sum_prop"]),
       y = cumsum(piEmplid[,"win.sum_prop"][order(piEmplid[,"win.sum_prop"], decreasing = TRUE)])[pickIndex(propColumn = piEmplid[,"win.sum_prop"])],
       col = "red",
       cex =1.5,
       pch = 19)

text(x = pickIndex(propColumn = piEmplid[,"win.sum_prop"]),
       y = cumsum(piEmplid[,"win.sum_prop"][order(piEmplid[,"win.sum_prop"], decreasing = TRUE)])[pickIndex(propColumn = piEmplid[,"win.sum_prop"])],
       col = "red",
       cex =1.25,
      #pos = 4,
       adj = c(-0.01,1.5),
       label = paste0(pickIndex(propColumn = piEmplid[,"win.sum_prop"]),
                      " (",
                      round(100*pickIndex(propColumn = piEmplid[,"win.sum_prop"])/length(piEmplid[,"win.sum_prop"]),0),
                      "%)",
                      " PI's are responsible for",
                      "\n   80%",
                      " of requested funds won"
                      )
 )

# 40%
points(x = pickIndex(percentageTarget = 0.4, propColumn = piEmplid[,"win.sum_prop"]),
       y = cumsum(piEmplid[,"win.sum_prop"][order(piEmplid[,"win.sum_prop"], decreasing = TRUE)])[pickIndex(percentageTarget = 0.4, propColumn = piEmplid[,"win.sum_prop"])],
       col = "red",
       cex =1.5,
       pch = 19)

text(x = pickIndex(percentageTarget = 0.4, propColumn = piEmplid[,"win.sum_prop"]),
       y = cumsum(piEmplid[,"win.sum_prop"][order(piEmplid[,"win.sum_prop"], decreasing = TRUE)])[pickIndex(percentageTarget = 0.4, propColumn = piEmplid[,"win.sum_prop"])],
       col = "red",
       cex =1.25,
      #pos = 4,
       adj = c(-0.01,1.5),
       label = paste0(pickIndex(percentageTarget = 0.4, propColumn = piEmplid[,"win.sum_prop"]),
                      " (",
                      round(100*pickIndex(percentageTarget = 0.4, propColumn = piEmplid[,"win.sum_prop"])/length(piEmplid[,"win.sum_prop"]),0),
                      "%)",
                      " PI's are responsible for",
                      "\n   40%",
                      " of requested funds won"
                      )
 )


# 10%
points(x = pickIndex(percentageTarget = 0.1, propColumn = piEmplid[,"win.sum_prop"]),
       y = cumsum(piEmplid[,"win.sum_prop"][order(piEmplid[,"win.sum_prop"], decreasing = TRUE)])[pickIndex(percentageTarget = 0.1, propColumn = piEmplid[,"win.sum_prop"])],
       col = "red",
       cex =1.5,
       pch = 19)

text(x = pickIndex(percentageTarget = 0.1, propColumn = piEmplid[,"win.sum_prop"]),
       y = cumsum(piEmplid[,"win.sum_prop"][order(piEmplid[,"win.sum_prop"], decreasing = TRUE)])[pickIndex(percentageTarget = 0.1, propColumn = piEmplid[,"win.sum_prop"])],
       col = "red",
       cex =1.25,
      #pos = 4,
       adj = c(-0.01,1.5),
       label = paste0(pickIndex(percentageTarget = 0.1, propColumn = piEmplid[,"win.sum_prop"]),
                    #  " (",
                    #  round(100*pickIndex(propColumn = piEmplid[,"win.sum_prop"])/length(piEmplid[,"win.sum_prop"]),0),
                    #  "%)",
                      " PI's are responsible for",
                      "\n  10%",
                      " of requested funds won"
                      )
 )



axis(side =1, at = seq(from = 0, to = 3000, by = 500))
axis(side =1, at = seq(from = 0, to = 3000, by = 500), labels = paste0(round(100*seq(from = 0, to = 3000, by = 500)/nrow(piEmplid),0),"%" ), line = 1, tick = FALSE, font = 3)

axis(side =2, 
     at = seq(from = 0, to = 1, by = 0.2),
     labels = paste0(seq(from = 0, to = 100, by = 20),"%"),
     las = 2
     )


mtext(side = c(1,2,3), 
      text = c("Count of PI's",
               "Percentage of total won funds requested (cumulative)",
        "The U follows the Pareto principle (80/20 rule) \nin winning funds"),
      font = c(1,1, 2),
      cex = c(1.5,1.5,1.5),
      line = c(4.5,4.25,2.618)
)

```



```{r}

# Save "piEmplid" object

saveRDS(piEmplid, here::here("Robjects", "Clustering PIs", "Percentiles from Pareto cluster child.rds"))

```
