---
output:
  html_document:
    keep_md: true
---

```{r include = FALSE}

# Notes

# This follows "Complex cluster of PIs child.Rmd

```



```{r include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.height=7, fig.width=10)

# adjust graphic parameters

oldPar <- par(cex.main = 3,
    cex.lab = 3,
    cex.axis = 2,
    mar = c(5.1,4.1,4.1,2.1) # default is c(5.1,4.1,4.1,2.1)
    )

```



```{r}

##########
## LOAD ##
##########


source(here::here("Prep scripts", "Prep Script from Win Rates by College.R"))

###############
## LIBRARIES ##
###############

library(viridis)
library(pheatmap)

######################
## CALCULATE PER PI ##
######################

piEmplid <- calculateWinRates(data = prepData, categoryColumn = "PROPOSAL_PI_EMPLID", functionList = list(mean = mean, median = median)) |> (\(x){ x[[1]] })()
# really starting to hate the list nature and why do I want "call" ?

piEmplid$count.total <- apply(piEmplid[,c("win.count","loss.count")], 1, sum, na.rm = TRUE )
piEmplid$sum.total <- apply(piEmplid[,c("win.sum","loss.sum")], 1, sum, na.rm = TRUE )

##########
## PREP ##
##########

piEmplid$emplid <- row.names(piEmplid)
filterThreeCount <- piEmplid$count.total > 3
piMap <- data.frame(college = row.names(piEmplid), abbrv = row.names(piEmplid), color = NA, pch = 19, cex = 0.3 )

#######################
## TEST FOR NORMALCY ##
#######################

normalTest <- apply(piEmplid[filterThreeCount,!colnames(piEmplid) %in% c("emplid", "count_cut","rate_cluster","totals_cluster", "rate_cluster")], 2, function(x) shapiro.test(x)$p.value) |>
  (\(x){any(x > 0.05) })() # FALSE

# unsurprisingly --- let's log transform before z-transform


#####################
## CREATE CLUSTERS ##
#####################

ratePCA <- piEmplid[filterThreeCount,] |>
  (\(x){return(data.frame(row.names = rownames(x),
                          sum.rate = x[,"sum.rate"],
                          count.rate = x[,"count.rate"]                 )
  ) })() |>
  scale() |>
  PCA(graph = FALSE)

rateHCPC <- ratePCA |>
  HCPC(nb.clust = 8, graph = FALSE)


```


```{r}
             
# Manage colors

rateClusterColors <- c("peru", "darkseagreen", "darkcyan", "slateblue", "goldenrod", "indianred", "mediumaquamarine", "firebrick")

clusterRateMapping <- piMap
clusterRateMapping[,"pch"] <- rep(19, nrow(clusterRateMapping))
clusterRateMapping[,"color"] <- rateClusterColors[rateHCPC$data.clust$clust[match(clusterRateMapping[,"college"], row.names(rateHCPC$data.clust))] ]

```



```{r}

# Merge clusters back into piEmplid
piEmplid <- merge(piEmplid, rateHCPC$data.clust[,"clust", drop = FALSE], by.x = "emplid", by.y= "row.names")

row.names(piEmplid) <- piEmplid$emplid
names(piEmplid)[names(piEmplid) == "clust"] <- "rate_cluster" 


# Merge into prepData
prepData <- merge(prepData, piEmplid[,c("emplid","rate_cluster")], by.x = "PROPOSAL_PI_EMPLID", by.y = "emplid", all.x = TRUE)

```



```{r}

# Calculate win rates

rateClusterWins <- calculateWinRates(prepData, categoryColumn = "rate_cluster", functionList = list(mean = mean, median = median)) |>
  (\(x){
    return(x$summary[match(levels(prepData$rate_cluster), row.names(x$summary)),])
  })()

rateClusterWins$color <- rateClusterColors

rateClusterWins$population <- aggregate(emplid ~ rate_cluster,  data = piEmplid, length )[,2]

rateClusterWins$pop_perc <- round(100*(rateClusterWins$population / sum(rateClusterWins$population)),1)


rateClusterWins$sum_perc <- round(100*(rateClusterWins$win.sum/sum(rateClusterWins$win.sum)),1)

rateClusterWins$count_total <- apply(rateClusterWins[,c("win.count","loss.count")], 1, sum)

rateClusterWins$count_perc <- round(100*(apply(rateClusterWins[,c("win.count","loss.count")], 1, sum)/sum(apply(rateClusterWins[,c("win.count","loss.count")], 1, sum))),1)


medianPIbyCluster <- aggregate(win.sum ~ rate_cluster, data = piEmplid, median) # for plot

```


```{r}

# Population heatmaps and table

piNames <- row.names(piEmplid)[piEmplid$count.total > 3]
piFilter <- prepData$PROPOSAL_PI_EMPLID %in% piNames

collegePopulation <- unique(prepData[piFilter, c("PROPOSAL_PI_EMPLID", "college","rate_cluster" )])

# > length(unique(collegePopulation[,1]))
# [1] 1672
# > length((collegePopulation[,1]))
# [1] 1922

# I think I'll just allow the duplication

```

```{r}

# cluster bounds

clusterBounds <- aggregate(cbind(sum.rate,count.rate) ~ rate_cluster, data = piEmplid, function(x){c(max = max(x),min = min(x))} )

```

# SIMPLE CLUSTERS BASED ON WIN RATES

### SECTIONS

  (1) Summary   
  (2) Method    
  (3) Clusters verbally described 
  (4) Clusters visually described
  (5) Table of aggregated values by cluster   
  (6) Hierachical clustering on principal components    
  (7) Clusters displayed on scatter plots   
  (8) Variable relevance per cluster   
  (9) Clusters over time    
  (10) Cluster populations by college and major institutions    

### 1. SUMMARY

This shows principal investigators sorted into eight clusters using only two criteria: the win rates by count of proposal and sum of funds requested.    

### 2. METHOD

First, principal investigators (PI's) with three or fewer proposals are filtered out.  `r format(table(filterThreeCount)[1], big.mark = ",")` principal investigators (`r  round(100*proportions(table(filterThreeCount))[1], 0) `%) are removed and `r format(table(filterThreeCount)[2], big.mark = ",")` principal investigators (`r  round(100*proportions(table(filterThreeCount))[2], 0) `%) are kept.    

Second, two variables are calculated per principal investigator:    

  * Rate of proposals won (count rate)        
  * Rate of funds requested won (sum rate)    

Third, the rates are centered and scaled.

Fourth, principle components are extracted.

Fifth, hierarchical clustering using Euclidean distance and Ward's method is used to produce eight clusters.

The clusters are named, described verbally and visually, and the cluster populations by colleges and major institutions are shown.  

### 3. CLUSTERS VERBALLY DESCRIBED

**CLUSTER 1: ROCK BOTTOM**    
  Principal investigators in Cluster 1 have less than ~20% win rates by both count and sum.
  
  It has a population of `r rateClusterWins[1,"population"]` PIs (`r rateClusterWins[1,"pop_perc"]`%) and accounts for only `r rateClusterWins$sum_perc[1] `% of total requested funds won.

**CLUSTER 2: DOMINANT**     
  Principal investigators in Cluster 2 have less than ~40% win rates by both count and sum.

  It has the largest share of the population and proposals submitted, as well as a large share of funds awarded. 

  It has a population of `r rateClusterWins[2,"population"]` PIs (`r rateClusterWins[2,"pop_perc"]`%) and accounts for `r rateClusterWins$sum_perc[2] `% of total requested funds won.     
  
**CLUSTER 3: BIG WHIFF**    
  Principal investigators in Cluster 3 have less than ~30% win rates by sum and less than ~70% win rate by count.   
  
  They have a respectable win rate by count but lose the larger proposals. 

  Cluster 3 has a population of `r rateClusterWins[3,"population"]` PIs (`r rateClusterWins[3,"pop_perc"]`%) and accounts for  `r rateClusterWins$sum_perc[3] `% of total requested funds won.
  
**CLUSTER 4:  MONEY**    
  Principal investigators in Cluster 4 have less than ~60% win rates by sum and less than ~50% win rates by count.

  It has the largest share of funds requested won.

  Cluster 4 has a population of `r rateClusterWins[4,"population"]` PIs (`r rateClusterWins[4,"pop_perc"]`%) and accounts for  `r rateClusterWins$sum_perc[4] `% of total requested funds won.
  
**CLUSTER 5:  MISSING THE BIG WINS**    
Principal investigators in Cluster 5 have less than ~50% win rates by sum and greater than ~50% win rates by count.

They have an excellent win rate by count but like Cluster 3, they lose the larger proposals.

Cluster 5 has a population of `r rateClusterWins[5,"population"]` PIs (`r rateClusterWins[5,"pop_perc"]`%) and accounts for a large majority (`r rateClusterWins$sum_perc[5] `%) of total requested funds won. 

**CLUSTER 6: BRONZE MEDAL**

Principal investigators in Cluster 6 have greater than ~40% win rates by sum and less than ~65% win rates by count.

Their win rates by sum are in third place behind Clusters 7 and 8.  And compared to Clusters 3 and 5, they win the large proposals.

Cluster 6 has a population of `r rateClusterWins[6,"population"]` PIs (`r rateClusterWins[6,"pop_perc"]`%) and accounts for a large majority (`r rateClusterWins$sum_perc[6] `%) of total requested funds won.

**CLUSTER 7:  SILVER MEDAL**    
Principal investigators in Cluster 7 have greater than ~50% win rates by sum and greater than ~60% win rates by count. They have the second-highest win rates.


Cluster 7 has a population of `r rateClusterWins[7,"population"]` PIs (`r rateClusterWins[7,"pop_perc"]`%) and accounts for a large majority (`r rateClusterWins$sum_perc[7] `%) of total requested funds won.

**CLUSTER 8:  GOLD MEDAL**   
Principal investigators in Cluster 8 have greater than ~80% win rates by sum and greater than ~65% win rates by count. They have the highest win rates.

Cluster 8 has a population of `r rateClusterWins[8,"population"]` PIs (`r rateClusterWins[8,"pop_perc"]`%) and accounts for a large majority (`r rateClusterWins$sum_perc[8] `%) of total requested funds won. 

### 4. CLUSTERS VISUALLY DESCRIBED

```{r fig.height = 4.5, fig.width = 9}

#########################################
## THREE PROPORTION BARS AND A BOXPLOT ##
#########################################


####################
## POPULATION BAR ##
####################

par(bg = "ivory",
    fg = "gray10",
    oma = c(0,0,3,0),
    mar = c(3,3,3,1),
    mfrow= c(1,2))

# dummy graph to create the background

theBat <- barplot(
  rateClusterWins[,"population"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  rateClusterWins[,"population"],
  col = rateClusterWins[,"color"],
  las = 2
)

mtext(text = row.names(rateClusterWins),
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font =2,
      col = rateClusterWins[,"color"])

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Population"
      )

text(x = theBar,
     y = rateClusterWins[,"population"],
     cex = rep(0.9,8),
     pos = c(3,1,3,3,3,3,3,3),
     font = 2,
     col = c(rep("gray10",8),"gray90"),
     labels = paste0(rateClusterWins$pop_perc,"%")
     )


###############
## FUNDS BAR ##
###############

# adjust margins

par(mar = c(3,5,3,1))

# dummy graph to create the background

theBat <- barplot(
  rateClusterWins[,"win.sum"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  rateClusterWins[,"win.sum"],
  col = rateClusterWins[,"color"],
  las = 2
)

mtext(text = row.names(rateClusterWins),
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font = 2,
      col = rateClusterWins[,"color"]
        )

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Funds requested won"
      )

text(x = theBar,
     y = rateClusterWins[,"win.sum"],
     pos = c(3,3,3,1,rep(3,4)),
     cex = rep(0.9,8),
     font = 2,
     col = c(rep("gray10",8),"gray90"),
     labels = paste0(rateClusterWins$sum_perc,"%")
     )

# float the median over THAT !

mtext(text = "Proportions and proposal amounts per cluster",
      side = 3,
      line = 0.309,
      font = 2,
      cex = 2,
      outer = TRUE)

###############
## COUNT BAR ##
###############

# adjust margins

par(mar = c(3,4,3,1),
    oma = c(0,0,0,0))

# par(bg = "ivory",
#    fg = "gray10",
#    oma = c(0,0,0,0),
#    mar = c(3,3,3,1)
#    )

# dummy graph to create the background

theBat <- barplot(
  rateClusterWins[,"count_total"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  rateClusterWins[,"count_total"],
  col = rateClusterWins[,"color"],
  las = 2
)

mtext(text = row.names(rateClusterWins),
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font = 2,
      col = rateClusterWins[,"color"]
        )

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Count of proposals"
      )

text(x = theBar,
     y = rateClusterWins[,"count_total"],
     pos = c(3,1, rep(3,6) ),
     cex = rep(0.9,8),
     font = 2,
     col = c(rep("gray10",8),"gray90"),
     labels = paste0(rateClusterWins$count_perc,"%")
     )


#############
## BOXPLOT ##
#############

par(mar = c(3,4,3,1))

# par(bg = "ivory",
#    fg = "gray10",
#    oma = c(0,0,0,0),
#    mar = c(3,4,4,1))
    
#    ,
#    mfrow= c(1,1))

boxplot(log(PROPOSAL_TOTAL_SPONSOR_BUDGET) ~ rate_cluster,
        data = prepData,
        names = rep("", length(unique(prepData$rate_cluster[!is.na(prepData$rate_cluster)])) ),
        yaxt = "n",
        ylab = ""
        )

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBox <- boxplot(log(PROPOSAL_TOTAL_SPONSOR_BUDGET) ~ rate_cluster,
        data = prepData,
        col = rateClusterWins$color,
        names = rep("",length(unique(prepData$rate_cluster[!is.na(prepData$rate_cluster)])) ),
        ylab = "",
        las = 1
        )

mtext(text = c("Proposal funds requested", "Funds requested (log)" ),
      side = c(3,2),
      font = c(1,1),
      cex = c(1.75,1.5),
      line = c(0.309, 2.427)
      
      )

mtext(text = row.names(rateClusterWins),
      line = 1,
      at = 1:8,
      side = 1,
      cex = 1.5,
      font = 2,
      col = rateClusterWins[,"color"]
        )

```

```{r}

# Simple bars describing the clusters

par(oma = c(0,0,3,0), mfrow = c(2,1))

plotWinRates(rateClusterWins,
             agg = "count",
             line = "count.rate",
             bar_params = list(mar = c(2,7,4,5)),
             bar_args = list(las = 1),
             bar_mtext_args = list(
               text = c("Proposals per cluster",
                        "",
                        "Count of proposals"
                        ),
               line = c(1.618,0,5),
               col = c("gray10", "gray10", "gray10"),
               font = c(1,3,1),
               cex = c(1.75,1.25,1.15),
               side = c(3,3,2)
                        ),
             bar_legend_args = list(x = "topleft")
             )

mtext(text = "Clusters described by count and sum",
      side = 3,
      line = 0.309,
      font = 2,
      cex = 2,
      outer = TRUE)

plotWinRates(rateClusterWins,
             agg = "sum",
             line = "sum.rate",
             bar_params = list(mar = c(2,7,4,5)),
             bar_args = list(las = 1),
             bar_mtext_args = list(
               text = c("Funds requested per cluster",
                        "",
                        "Sum of funds requested"
                        ),
               line = c(1.618,0,5),
               col = c("gray10", "gray10", "gray10"),
               font = c(1,3,1),
               cex = c(1.75,1.25,1.15),
               side = c(3,3,2)
                        ),
             bar_legend_args = list(x = "topleft")
)

```

### 5. TABLE OF AGGREGATED VALUES BY CLUSTER

```{r}

displayColumns <- c(
    "population",
    "pop_perc",
    "count_total", 
    "count_perc",
    "win.sum",
    "sum_perc",
    "win.count",   
    "win.mean",    
    "win.median", 
  "loss.sum",    
  "loss.count",  
  "loss.mean",   
  "loss.median",
  "sum.rate",    
  "count.rate",  
  "color"
)


columnRounding <- c(
    "population" = 0,
    "pop_perc" =1,
    "count_total" =-1, 
    "count_perc" = 0,
    "win.sum" = -6,
    "sum_perc" = 0,
    "win.count" = -1,   
    "win.mean" = -3,    
    "win.median" = -3, 
  "loss.sum" = -6,    
  "loss.count" = -1,  
  "loss.mean" = -2,   
  "loss.median" = -2,
  "sum.rate" = 2,    
  "count.rate" = 2  
)

# Convert result_df to a flextable
ft <- data.frame(cluster = rownames(rateClusterWins), rateClusterWins[,displayColumns] ) |>
  (\(x){ 
      x[, names(columnRounding)] <- Map(round, x[, names(columnRounding)], columnRounding);
    x  # Return the modified data frame
    })() |>
  flextable::flextable() |>
  flextable::theme_vanilla() %>%  # Apply a clean theme
  flextable::set_caption("Aggregated Metrics by Rate Cluster") %>%  # Add a title
  flextable::colformat_num(j = 2:ncol(rateClusterWins), digits = 2) %>%  # Format numeric columns
  flextable::autofit()  # Adjust column widths for better display

# Print the table
ft


```


### 6. HIERARCHICAL CLUSTERING ON PRINCIPAL COMPONENTS


```{r fig.height = 4.5, fig.width = 9}

library(gridExtra)

p1 <- fviz_pca_biplot(ratePCA,
                       geom = "none",
                       title = "Principal components of multiple variables")

p2 <- fviz_cluster(rateHCPC, 
                   geom = "point",
                   main = "PI clusters plotted on component dimensions",
                   show.clust.cent = TRUE)

# Arrange plots side by side
grid.arrange(p1, p2, ncol = 2)  # Change `ncol = 1` for vertical arrangement

```

```{r fig.height = 4.5, fig.width=4.5}

# Dendrogram

if(length(rateHCPC[["desc.var"]][[2]]) <= 12) {
  plot(rateHCPC, choice = "tree", cex = 0.05)
}

```

### 7. CLUSTERS DISPLAYED ON SCATTER PLOTS

```{r}

# Plot by win sum and rate clusters

par(oma = c(0,3,4,0), mfrow =c(1,1))

yBYx(data = piEmplid,  
     x_axis =  "win.sum", 
     y_axis = "sum.rate", 
     detailMapping = clusterRateMapping, 
     log = c("x"), 
     scatter_args = list(xlim = c(0,18.5)),
     scatter_text_args = list(
       labels = "",
       cex = 1.618),
     scatter_mtext_args = list(text = c("", "Sum of funds requested won (log)", "Fraction of funds requested won"), 
                               cex = c(1.75,1.5,1.5),
                               font = c(0,0,0)   ))

mtext(text = "Sum rate by total funds won" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

text(x = log(medianPIbyCluster$win.sum) |> (\(x){ x[is.infinite(x)] <- 0; return(x)})() ,
     y = rateClusterWins$sum.rate,
     labels = 1:(length(rateHCPC[["desc.var"]][[2]])),
     cex = 2,
     font = 2,
     col = rateClusterWins$color,
     xpd = TRUE
) 

```


```{r fig.width=10}

# Tiled sum scatter plots of rate clusters

par(oma = c(5,7,4,1), mfrow = c(4,2))

invisible(
  lapply(levels(piEmplid$rate_cluster), function(x){ 
    
    yBYx(data = piEmplid[piEmplid$rate_cluster == x,],  
         x_axis = "win.sum", 
         y_axis = "sum.rate", 
         detailMapping = clusterRateMapping, 
         log = c("x"), 
         scatter_params = list(mar = c(0,0,0,0)),
         scatter_args = list(cex = 1.618, xlim = c(0,20), ylim = c(0,1)),
         scatter_text_args = list(
           labels = "",
           cex = 1.618),
         scatter_mtext_args = list(text = c("", "", "") )
         #                              cex = c(1.75,1.5,1.5),
         #                              font = c(0,0,0)   )
    )
    
    textFilter <- row.names(rateClusterWins) == x
    #text(x = totalsClusterWins$count.rate[textFilter],
    #    y = totalsClusterWins$sum.rate[textFilter],
    #    labels = row.names(totalsClusterWins)[textFilter],
    #    cex = 2,
    #    font = 2,
    #    col = totalsClusterWins$color[textFilter]
    #    )   
    
    text(x = 0.05,
         y = 0.9,
         labels = row.names(rateClusterWins)[textFilter],
         cex = 3,
         font = 2,
         col = rateClusterWins$color[textFilter],
         xpd = TRUE
    )  
    
    
  })    
)    



# mtext(text = "Clusters based on multiple dimensions" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

mtext(text = "Sum of funds requested won (log)" , side = 1, outer = TRUE, font = 1, cex = 1.5, line = 3) 

mtext(text = "Fraction of funds requested won" , side = 2, outer = TRUE, font = 1, cex = 1.5, line = 3) 

```


```{r}

# Rate scatter plot

par(oma = c(0,3,4,0), mfrow =c(1,1))

yBYx(data = piEmplid,  
     x_axis = "count.rate", 
     y_axis = "sum.rate", 
     detailMapping = clusterRateMapping, 
     log = NA, 
     scatter_text_args = list(
       labels = "",
       cex = 1.618),
     scatter_mtext_args = list(text = c("", "Fraction of proposals won", "Fraction of funds requested won"), 
                               cex = c(1.75,1.5,1.5),
                               font = c(0,0,0)   ))

text(x = rateClusterWins$count.rate,
     y = rateClusterWins$sum.rate,
     labels = row.names(rateClusterWins),
     cex = 2,
     font = 2,
     col = rateClusterWins$color
) 

mtext(text = "Sum rate and count rate" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

```

```{r}

# Tiled rate scatter plots of rate clusters

par(oma = c(5,7,4,1), mfrow = c(4,2))

invisible(
  lapply(levels(piEmplid$rate_cluster), function(x){ 
    
    
    
    yBYx(data = piEmplid[piEmplid$rate_cluster == x,],  
         x_axis = "count.rate", 
         y_axis = "sum.rate", 
         detailMapping = clusterRateMapping, 
         log = NA, 
         scatter_params = list(mar = c(0,0,0,0)),
         scatter_args = list(cex = 1.618, xlim = c(0,1), ylim = c(0,1)),
         scatter_text_args = list(
           labels = "",
           cex = 1.618),
         scatter_mtext_args = list(text = c("", "", "") )
         #                              cex = c(1.75,1.5,1.5),
         #                              font = c(0,0,0)   )
    )
    
    textFilter <- row.names(rateClusterWins) == x
    #text(x = totalsClusterWins$count.rate[textFilter],
    #    y = totalsClusterWins$sum.rate[textFilter],
    #    labels = row.names(totalsClusterWins)[textFilter],
    #    cex = 2,
    #    font = 2,
    #    col = totalsClusterWins$color[textFilter]
    #    )   
    
    text(x = 0.05,
         y = 0.9,
         labels = row.names(rateClusterWins)[textFilter],
         cex = 3,
         font = 2,
         col = rateClusterWins$color[textFilter],
         xpd = TRUE
    )  
    
    
  })    
)    



# mtext(text = "Clusters based on multiple dimensions" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

mtext(text = "Fraction of funds requested won" , side = 2, outer = TRUE, font = 1, cex = 1.5, line = 3) 

mtext(text = "Fraction of proposals won" , side = 1, outer = TRUE, font = 1, cex = 1.5, line = 3) 

```

### 8. VARIABLE RELEVANCE PER CLUSTER

*This section is omitted due to the few variables used in clustering.  Both variables are relevant to all clusters.*

```{r eval=FALSE}

# Variable relevance per cluster

oldPar <- par(oma = c(0,0,4,1),
              bg = "ivory",
              fg = "gray10",
              mfrow = c(4,2), 
              mar = c(3,7,0,0))

invisible(
  lapply(levels(piEmplid$rate_cluster), function(x){ 
 
    # Extract v.test values for the cluster
    vtest_values <- rateHCPC$desc.var$quanti[[x]][, "v.test"]
         
    # Generate a color gradient using viridis, centered around 0
    colors <- viridis(length(vtest_values), option = "G")  # "C" is a blue-to-yellow scale
             
    # Order values for proper coloring
    ordered_idx <- order(vtest_values)  
    sorted_values <- vtest_values[ordered_idx]  
    sorted_colors <- colors[rank(sorted_values, ties.method = "first")]  # Match colors to sorted values
    
    rateHCPC$desc.var$quanti[[x]][,"v.test"] |> 
      (\(x){ 
      barplot(
        rev(x),
        las = 2,
        horiz = TRUE,
        col = sorted_colors)   
      })()
        
    legend("topleft", legend = x, bty = "n", text.font = 2, cex = 2, text.col = rateClusterWins[as.numeric(x),"color"])    
})
)

mtext(side = 3, 
      font = 2, 
      cex = 2, 
      text = "Variable relevance per cluster",
      col = "gray10",
      line = 0.618,
      outer = TRUE)
```

### 9. CLUSTERS OVER TIME

Although this clustering did not include any time variables, the submission patterns by cluster are shown.

```{r fig.height = 4}

# I need to fix the timeDots graphics before I can get this to tile.  I should  probably leave this out.

piYr <- elapsed(data = prepData[!is.na(prepData$rate_cluster), ])


# I haven't improved my timeline graphics yet

par(mfrow = c(1,1), bg = "ivory", fg = "gray10",
    mar = c(0,0,0,0))

invisible(
lapply(levels(prepData$rate_cluster), function(x){ 

  filter <- piYr$each[,"PROPOSAL_PI_EMPLID"] %in% piEmplid$emplid[piEmplid$rate_cluster == x]
  
# par(new = TRUE)  
timeline(data = piYr$each[filter,])
mtext(x, side = 3, line = -2, cex = 2, font = 2, outer = FALSE, col = rateClusterWins$color[as.numeric(x)])
  
  })
)



```

### 10. CLUSTER POPULATIONS BY COLLEGE AND MAJOR INSTITUTIONS

 `r nrow(collegePopulation) - length(unique(collegePopulation[,1]))` (`r  ( (nrow(collegePopulation) - length(unique(collegePopulation[,1]))) /length(unique(collegePopulation[,1])) ) |> (\(x){round(x*100, 1)})()`)%  PI's appear in multiple colleges or institutions.


```{r}

table(collegePopulation[,c("college", "rate_cluster")]) |>
  as.data.frame.matrix() |>
  apply(1, proportions) |>
  t() |>
  pheatmap(
    scale = "column",
  #  Rowv = NA,
  #  Colv = NA,
    main = "Organization by cluster",
  #  margins = c(3,3)
  fontsize_col = 15,
  angle_col = 0
  )


# that's pretty much what I want

# let's publish the table now

collegeTable <- table(collegePopulation[,c("college", "rate_cluster")]) |>
  as.data.frame.matrix() |>
  apply(1, proportions) |>
  t() |>
  as.data.frame() |>
  (\(x){round(100*x,1) })() |>
  (\(x){cbind(Org = rownames(x), x) })() |>  #
  flextable::flextable() |>
  flextable::theme_vanilla() |>  # Apply a clean theme
  flextable::set_caption("Percent of organization population per cluster") |>  # Add a title
  flextable::colformat_num(j = 2:6, digits = 2) |>  # Format numeric columns
  flextable::autofit()  # Adjust column widths for better display

# Print the table
collegeTable

# man that's beautiful


# let's reverse the table -- 

table(collegePopulation[,c("college", "rate_cluster")]) |>
  as.data.frame.matrix() |>
  apply(2, proportions) |>
  t() |>
  pheatmap(
    scale = "column",
  #  Rowv = NA,
  #  Colv = NA,
    main = "Cluster by organization",
  #  margins = c(3,3)
  las = 3
  )


collegeTable <- table(collegePopulation[,c("college", "rate_cluster")]) |>
  as.data.frame.matrix() |>
  apply(2, proportions) |>
  t() |>
  as.data.frame() |>
  (\(x){round(100*x,1) })() |>
  (\(x){cbind(Org = rownames(x), x) })() |>  #
  flextable::flextable() |>
  flextable::theme_vanilla() |>  # Apply a clean theme
  flextable::set_caption("Percent of cluster population per organization") |>  # Add a title
  flextable::colformat_num(j = 2:6, digits = 2) |>  # Format numeric columns
  flextable::autofit()  # Adjust column widths for better display

# Print the table
collegeTable



```



```{r}

# Save "piEmplid" object

saveRDS(piEmplid, here::here("Robjects", "Clustering PIs", "Rate cluster from rate cluster child.rds"))

```

