---
output: 
  powerpoint_presentation:
    reference_doc: Reference_Template_vB0.pptx
---

```{r include = FALSE}

# This is a child ppt presentation that describes simple clustering based on rates

```

```{r include = FALSE}

knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.height=6.75, fig.width=12)

# adjust graphic parameters

oldPar <- par(cex.main = 3,
    cex.lab = 3,
    cex.axis = 2,
    mar = c(5.1,4.1,4.1,2.1) # default is c(5.1,4.1,4.1,2.1)
    )

# par(oldPar) #restore old parameters # restore old parameters after the plot

```

```{r}

##########
## LOAD ##
##########

source(here::here("Prep scripts", "Adjusting prepData and loading things.R"))

###############
## LIBRARIES ##
###############

library(viridis)
library(pheatmap)

######################
## CALCULATE PER PI ##
######################

piEmplid <- calculateWinRates(data = prepData, categoryColumn = "PROPOSAL_PI_EMPLID", functionList = list(mean = mean, median = median)) |> (\(x){ x[[1]] })()
# really starting to hate the list nature and why do I want "call" ?

piEmplid$count.total <- apply(piEmplid[,c("win.count","loss.count")], 1, sum, na.rm = TRUE )
piEmplid$sum.total <- apply(piEmplid[,c("win.sum","loss.sum")], 1, sum, na.rm = TRUE )

##########
## PREP ##
##########

piEmplid$emplid <- row.names(piEmplid)
filterThreeCount <- piEmplid$count.total > 3
piMap <- data.frame(college = row.names(piEmplid), abbrv = row.names(piEmplid), color = NA, pch = 19, cex = 0.3 )

#######################
## TEST FOR NORMALCY ##
#######################

normalTest <- apply(piEmplid[filterThreeCount,!colnames(piEmplid) %in% c("emplid", "count_cut","rate_cluster","totals_cluster", "rate_cluster")], 2, function(x) shapiro.test(x)$p.value) |>
  (\(x){any(x > 0.05) })() # FALSE

# unsurprisingly --- let's log transform before z-transform


#####################
## CREATE CLUSTERS ##
#####################

ratePCA <- piEmplid[filterThreeCount,] |>
  (\(x){return(data.frame(row.names = rownames(x),
                          sum.rate = x[,"sum.rate"],
                          count.rate = x[,"count.rate"]                 )
  ) })() |>
  scale() |>
  PCA(graph = FALSE)

rateHCPC <- ratePCA |>
  HCPC(nb.clust = 8, graph = FALSE)


```


```{r}
             
# Manage colors

rateClusterColors <- c("peru", "darkseagreen", "darkcyan", "slateblue", "goldenrod", "indianred", "mediumaquamarine", "firebrick")

clusterRateMapping <- piMap
clusterRateMapping[,"pch"] <- rep(19, nrow(clusterRateMapping))
clusterRateMapping[,"color"] <- rateClusterColors[rateHCPC$data.clust$clust[match(clusterRateMapping[,"college"], row.names(rateHCPC$data.clust))] ]

```



```{r}

# Merge clusters back into piEmplid
piEmplid <- merge(piEmplid, rateHCPC$data.clust[,"clust", drop = FALSE], by.x = "emplid", by.y= "row.names")

row.names(piEmplid) <- piEmplid$emplid
names(piEmplid)[names(piEmplid) == "clust"] <- "rate_cluster" 


# Merge into prepData
prepData <- merge(prepData, piEmplid[,c("emplid","rate_cluster")], by.x = "PROPOSAL_PI_EMPLID", by.y = "emplid", all.x = TRUE)

```



```{r}

# Calculate win rates

rateClusterWins <- calculateWinRates(prepData, categoryColumn = "rate_cluster", functionList = list(mean = mean, median = median)) |>
  (\(x){
    return(x$summary[match(levels(prepData$rate_cluster), row.names(x$summary)),])
  })()

rateClusterWins$color <- rateClusterColors

rateClusterWins$population <- aggregate(emplid ~ rate_cluster,  data = piEmplid, length )[,2]

rateClusterWins$pop_perc <- round(100*(rateClusterWins$population / sum(rateClusterWins$population)),1)


rateClusterWins$sum_perc <- round(100*(rateClusterWins$win.sum/sum(rateClusterWins$win.sum)),1)

rateClusterWins$count_total <- apply(rateClusterWins[,c("win.count","loss.count")], 1, sum)

rateClusterWins$count_perc <- round(100*(apply(rateClusterWins[,c("win.count","loss.count")], 1, sum)/sum(apply(rateClusterWins[,c("win.count","loss.count")], 1, sum))),1)


medianPIbyCluster <- aggregate(win.sum ~ rate_cluster, data = piEmplid, median) # for plot

```


```{r}

# Population heatmaps and table

piNames <- row.names(piEmplid)[piEmplid$count.total > 3]
piFilter <- prepData$PROPOSAL_PI_EMPLID %in% piNames

collegePopulation <- unique(prepData[piFilter, c("PROPOSAL_PI_EMPLID", "college","rate_cluster" )])

# > length(unique(collegePopulation[,1]))
# [1] 1672
# > length((collegePopulation[,1]))
# [1] 1922

# I think I'll just allow the duplication

```

```{r}

# cluster bounds

clusterBounds <- aggregate(cbind(sum.rate,count.rate) ~ rate_cluster, data = piEmplid, function(x){c(max = max(x),min = min(x))} )

```


# SIMPLE CLUSTERS BASED ON WIN RATES

# SECTIONS

  (1) Summary   
  (2) Method    
  (3) Clusters verbally described 
  (4) Clusters visually described
  (5) Table of aggregated values by cluster   
  (6) Hierachical clustering on principal components    
  (7) Clusters displayed on scatter plots   
  (8) Variable relevance per cluster   
  (9) Clusters over time    
  (10) Cluster populations by college and major institutions    
  
# 1. SUMMARY

This shows principal investigators sorted into eight clusters using only two criteria: the win rates by count of proposal and sum of funds requested.    
  

# 2. METHOD

First, principal investigators (PI's) with three or fewer proposals are filtered out.  `r format(table(filterThreeCount)[1], big.mark = ",")` principal investigators (`r  round(100*proportions(table(filterThreeCount))[1], 0) `%) are removed and `r format(table(filterThreeCount)[2], big.mark = ",")` principal investigators (`r  round(100*proportions(table(filterThreeCount))[2], 0) `%) are kept.    

Second, two variables are calculated per principal investigator:    

  * Rate of proposals won (count rate)        
  * Rate of funds requested won (sum rate)    

Third, the rates are centered and scaled.

Fourth, principle components are extracted.

Fifth, hierarchical clustering using Euclidean distance and Ward's method is used to produce eight clusters.

The clusters are named, described verbally and visually, and the cluster populations by colleges and major institutions are shown.  

# 3. CLUSTERS VERBALLY DESCRIBED

**CLUSTER 1: ROCK BOTTOM**    
  Principal investigators in Cluster 1 have less than ~20% win rates by both count and sum.
  
  It has a population of `r rateClusterWins[1,"population"]` PIs (`r rateClusterWins[1,"pop_perc"]`%) and accounts for only `r rateClusterWins$sum_perc[1] `% of total requested funds won.

**CLUSTER 2: DOMINANT**     
  Principal investigators in Cluster 2 have less than ~40% win rates by both count and sum.

  It has the largest share of the population and proposals submitted, as well as a large share of funds awarded. 

  It has a population of `r rateClusterWins[2,"population"]` PIs (`r rateClusterWins[2,"pop_perc"]`%) and accounts for `r rateClusterWins$sum_perc[2] `% of total requested funds won.     
  
**CLUSTER 3: BIG WHIFF**    
  Principal investigators in Cluster 3 have less than ~30% win rates by sum and less than ~70% win rate by count.   
  
  They have a respectable win rate by count but lose the larger proposals. 

  Cluster 3 has a population of `r rateClusterWins[3,"population"]` PIs (`r rateClusterWins[3,"pop_perc"]`%) and accounts for  `r rateClusterWins$sum_perc[3] `% of total requested funds won.
  
**CLUSTER 4:  MONEY**    
  Principal investigators in Cluster 4 have less than ~60% win rates by sum and less than ~50% win rates by count.

  It has the largest share of funds requested won.

  Cluster 4 has a population of `r rateClusterWins[4,"population"]` PIs (`r rateClusterWins[4,"pop_perc"]`%) and accounts for  `r rateClusterWins$sum_perc[4] `% of total requested funds won.
  
**CLUSTER 5:  MISSING THE BIG WINS**    
Principal investigators in Cluster 5 have less than ~50% win rates by sum and greater than ~50% win rates by count.

They have an excellent win rate by count but like Cluster 3, they lose the larger proposals.

Cluster 5 has a population of `r rateClusterWins[5,"population"]` PIs (`r rateClusterWins[5,"pop_perc"]`%) and accounts for a large majority (`r rateClusterWins$sum_perc[5] `%) of total requested funds won. 

**CLUSTER 6: BRONZE MEDAL**

Principal investigators in Cluster 6 have greater than ~40% win rates by sum and less than ~65% win rates by count.

Their win rates by sum are in third place behind Clusters 7 and 8.  And compared to Clusters 3 and 5, they win the large proposals.

Cluster 6 has a population of `r rateClusterWins[6,"population"]` PIs (`r rateClusterWins[6,"pop_perc"]`%) and accounts for a large majority (`r rateClusterWins$sum_perc[6] `%) of total requested funds won.

**CLUSTER 7:  SILVER MEDAL**    
Principal investigators in Cluster 7 have greater than ~50% win rates by sum and greater than ~60% win rates by count. They have the second-highest win rates.


Cluster 7 has a population of `r rateClusterWins[7,"population"]` PIs (`r rateClusterWins[7,"pop_perc"]`%) and accounts for a large majority (`r rateClusterWins$sum_perc[7] `%) of total requested funds won.

**CLUSTER 8:  GOLD MEDAL**   
Principal investigators in Cluster 8 have greater than ~80% win rates by sum and greater than ~65% win rates by count. They have the highest win rates.

Cluster 8 has a population of `r rateClusterWins[8,"population"]` PIs (`r rateClusterWins[8,"pop_perc"]`%) and accounts for a large majority (`r rateClusterWins$sum_perc[8] `%) of total requested funds won.


# 4. CLUSTERS VISUALLY DESCRIBED

```{r}

# fig.height = 4.5, fig.width = 9

#########################################
## THREE PROPORTION BARS AND A BOXPLOT ##
#########################################


####################
## POPULATION BAR ##
####################

par(bg = "ivory",
    fg = "gray10",
    oma = c(0,0,3,0),
    mar = c(3,3,3,1),
    mfrow= c(2,2))

# dummy graph to create the background

theBat <- barplot(
  rateClusterWins[,"population"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  rateClusterWins[,"population"],
  col = rateClusterWins[,"color"],
  las = 2
)

mtext(text = row.names(rateClusterWins),
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font =2,
      col = rateClusterWins[,"color"])

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Population"
      )

text(x = theBar,
     y = rateClusterWins[,"population"],
     cex = rep(0.9,8),
     pos = c(3,1,3,3,3,3,3,3),
     font = 2,
     col = c(rep("gray10",8),"gray90"),
     labels = paste0(rateClusterWins$pop_perc,"%")
     )


###############
## FUNDS BAR ##
###############

# adjust margins

par(mar = c(3,5,3,1))

# dummy graph to create the background

theBat <- barplot(
  rateClusterWins[,"win.sum"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  rateClusterWins[,"win.sum"],
  col = rateClusterWins[,"color"],
  las = 2
)

mtext(text = row.names(rateClusterWins),
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font = 2,
      col = rateClusterWins[,"color"]
        )

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Funds requested won"
      )

text(x = theBar,
     y = rateClusterWins[,"win.sum"],
     pos = c(3,3,3,1,rep(3,4)),
     cex = rep(0.9,8),
     font = 2,
     col = c(rep("gray10",8),"gray90"),
     labels = paste0(rateClusterWins$sum_perc,"%")
     )

# float the median over THAT !

mtext(text = "Proportions and proposal amounts per cluster",
      side = 3,
      line = 0.309,
      font = 2,
      cex = 2,
      outer = TRUE)

###############
## COUNT BAR ##
###############

# adjust margins

par(mar = c(3,4,3,1),
    oma = c(0,0,0,0))

# par(bg = "ivory",
#    fg = "gray10",
#    oma = c(0,0,0,0),
#    mar = c(3,3,3,1)
#    )

# dummy graph to create the background

theBat <- barplot(
  rateClusterWins[,"count_total"],
  xlab = "",
  ylab = "",
  yaxt = "n",
  xaxt = "n"
)

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBar <- barplot(
  rateClusterWins[,"count_total"],
  col = rateClusterWins[,"color"],
  las = 2
)

mtext(text = row.names(rateClusterWins),
      line = 1,
      at = theBar,
      side = 1,
      cex = 1.5,
      font = 2,
      col = rateClusterWins[,"color"]
        )

mtext(side = 3,
      cex = 1.75,
      line = 0.309,
      text = "Count of proposals"
      )

text(x = theBar,
     y = rateClusterWins[,"count_total"],
     pos = c(3,1, rep(3,6) ),
     cex = rep(0.9,8),
     font = 2,
     col = c(rep("gray10",8),"gray90"),
     labels = paste0(rateClusterWins$count_perc,"%")
     )


#############
## BOXPLOT ##
#############

par(mar = c(3,4,3,1))

# par(bg = "ivory",
#    fg = "gray10",
#    oma = c(0,0,0,0),
#    mar = c(3,4,4,1))
    
#    ,
#    mfrow= c(1,1))

boxplot(log(PROPOSAL_TOTAL_SPONSOR_BUDGET) ~ rate_cluster,
        data = prepData,
        names = rep("", length(unique(prepData$rate_cluster[!is.na(prepData$rate_cluster)])) ),
        yaxt = "n",
        ylab = ""
        )

# now define the proper par("usr") values
    default_rect_args <- list(
      xleft = par("usr")[1], 
      ybottom = par("usr")[3], 
      xright = par("usr")[2], 
      ytop = par("usr")[4],
      col = "gray90",
      border = NA
    )
  
    rect_args = list()
      
    # draw background color
    rect_args <- modifyList(default_rect_args, rect_args)
    do.call("rect", rect_args)

grid(nx = NA, ny = NULL, col = "gray99", lwd = 2)

par(new = TRUE)

theBox <- boxplot(log(PROPOSAL_TOTAL_SPONSOR_BUDGET) ~ rate_cluster,
        data = prepData,
        col = rateClusterWins$color,
        names = rep("",length(unique(prepData$rate_cluster[!is.na(prepData$rate_cluster)])) ),
        ylab = "",
        las = 1
        )

mtext(text = c("Proposal funds requested", "Funds requested (log)" ),
      side = c(3,2),
      font = c(1,1),
      cex = c(1.75,1.5),
      line = c(0.309, 2.427)
      
      )

mtext(text = row.names(rateClusterWins),
      line = 1,
      at = 1:8,
      side = 1,
      cex = 1.5,
      font = 2,
      col = rateClusterWins[,"color"]
        )

```

# 4. CLUSTERS DESCRIBED BY COUNT AND SUM    

```{r}

# Simple bars describing the clusters

par(oma = c(0,0,3,0), mfrow = c(2,1))

plotWinRates(rateClusterWins,
             agg = "count",
             line = "count.rate",
             bar_params = list(mar = c(2,7,4,5)),
             bar_args = list(las = 1),
             bar_mtext_args = list(
               text = c("Proposals per cluster",
                        "",
                        "Count of proposals"
                        ),
               line = c(1.618,0,5),
               col = c("gray10", "gray10", "gray10"),
               font = c(1,3,1),
               cex = c(1.75,1.25,1.15),
               side = c(3,3,2)
                        ),
             bar_legend_args = list(x = "topleft")
             )

mtext(text = "Clusters described by count and sum",
      side = 3,
      line = 0.309,
      font = 2,
      cex = 2,
      outer = TRUE)

plotWinRates(rateClusterWins,
             agg = "sum",
             line = "sum.rate",
             bar_params = list(mar = c(2,7,4,5)),
             bar_args = list(las = 1),
             bar_mtext_args = list(
               text = c("Funds requested per cluster",
                        "",
                        "Sum of funds requested"
                        ),
               line = c(1.618,0,5),
               col = c("gray10", "gray10", "gray10"),
               font = c(1,3,1),
               cex = c(1.75,1.25,1.15),
               side = c(3,3,2)
                        ),
             bar_legend_args = list(x = "topleft")
)

```

# RATE SCATTER PLOT

```{r}

# Rate scatter plot

par(oma = c(0,3,4,0), mfrow =c(1,1))

yBYx(data = piEmplid,  
     x_axis = "count.rate", 
     y_axis = "sum.rate", 
     detailMapping = clusterRateMapping, 
     log = NA, 
     scatter_text_args = list(
       labels = "",
       cex = 1.618),
     scatter_mtext_args = list(text = c("", "Fraction of proposals won", "Fraction of funds requested won"), 
                               cex = c(1.75,1.5,1.5),
                               font = c(0,0,0)   ))

text(x = rateClusterWins$count.rate,
     y = rateClusterWins$sum.rate,
     labels = row.names(rateClusterWins),
     cex = 2,
     font = 2,
     col = rateClusterWins$color
) 

mtext(text = "Sum rate and count rate" , outer = TRUE, font = 2, cex = 2, line = 0.309) 

```

